---
title: "SRF PPL & IUP Standardization"
author: "Walker Grimshaw & Phil Cork"
date: "2022-12-20"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## Introduction

This notebook transforms raw data scraped from state PPLs and IUPs and provides a single, harmonized output. 

Details for necessary transformations, crucial context, and discussion on decisions and assumptions can be found in each state's Background Information PDFs, available as a data download through the SRF Dashboard.


Process Overview

The overall process flow for each state is:
1. clean names
2. remove rows and columns that don't include project data
3. process numeric columns
4. process text columns
5. add state name
6. standardize column names
7. keep only cleaned and standardized column names

Once this process is done for each state, the cleaned dataframes are combined and formatted as input for `dw-dashboard-data-cleaner.R`
which prepares the data for external presentation in the SRF Dashboard.


Outline code

Within this documents outline, parenthetical codes denote where in process each state is:
- None: Not currently in need of updates
- A: Applicant data needs to be added
- ADD: Add to Data Dictionary, awaiting adding documentation page
- DD: Added to Data Dictionary, awaiting review from policy team
- NS: Not Started
- Q: Question currently blocking progress, awaiting resolution from policy team
- WIP: Work in progress



## Process States

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen=999999999)

library(tidyverse)
library(data.table)
library(janitor)
library(aws.s3)
library(googledrive)

```


## Alabama (Q in SS) (A) 

New Columns: 
- No mention of *Interest Rates or Loan Terms*
- Applicant data includes `applied for project amount` which can be interpreted as requested amount

```{r}

### APPLICANT ### 

# (364, 11)
al_applicant <- fread("data/year1/csv/1-Alabama_ARPA_SRF_Combined_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

al_app_clean <- al_applicant %>%
  # remove columns
  select(-date_pre_app_or_supp_received, -project_award_date_when_grant_is_signed,
         -unfunded_mat_ch_portion) %>%
  # process numeric columns
  mutate(
  requested_amount = as.numeric(str_replace_all(applied_for_project_amount,"[^0-9.]", "")),
  funding_amount = as.numeric(str_replace_all(funded_portion,"[^0-9.]", "")),
  principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness_and_or_grant,"[^0-9.]", "")),
  ) %>%
  # process text columns
  mutate(
    borrower = str_squish(applicant_name),
    project_description = str_squish(project_description),
    cities_served = str_squish(county),
    fundable = case_when(
      project_approved_yes_or_no == "Yes" ~ "Fundable",
      TRUE ~ "Applicant"),
    state = "Alabama"
  ) %>%
  select(borrower, cities_served, project_description, requested_amount, funding_amount,
         principal_forgiveness_amount, state, fundable)


### FUNDABLE ### 

# (7,14)
al_base <- fread("data/year1/csv/1-Alabama_Base-PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# (19,14)
al_bil <- fread("data/year1/csv/1-Alabama_BIL-PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# (3,14)
al_lead <- fread("data/year1/csv/1-Alabama_Lead-PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (29,14)
al_comb <- bind_rows(al_base, al_bil, al_lead) 

al_clean <- al_comb %>%
  # remove columns
  select(-gpr_component_cost, -gpr_type, -gpr_project, -estimated_construction_start_date) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population_served,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(assistance_amount,"[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(additional_subsidization_principal_forgiveness,
                                                                   "[^0-9.]", "")),
         ) %>%
  # process text columns
  mutate(cities_served = str_squish(county_served),
         borrower = str_replace_all(applicant_name, "\\*", ""),
         borrower = str_squish(borrower),
         state_score = str_replace_all(priority_point_rank,"[^0-9.]", ""),
         disadvantaged = case_when(
           disadvantaged_criteria == "Y" ~ "Yes",
           TRUE ~ "No"),
         pwsid = str_squish(pwsid_number),
         pwsid = str_replace(pwsid, "AL000341", "AL0000341"),
         project_name = str_squish(project_name),
         project_description = str_squish(project_description),
         project_description = str_replace(project_description, "Project Description: ", ""),
         # Lead projects when LEAD funds specified
         project_type = case_when(
           grepl("LEAD", appropriation_fund_being_used, ignore.case=TRUE) ~ "Lead",
           TRUE ~ "Other"),
         # ALL IUP projects fundable
         fundable = "Fundable",
         state = "Alabama"
  ) %>%
  select(state_score, borrower, pwsid, project_name, project_description, funding_amount, principal_forgiveness_amount, 
         # add requested amount in once the applicant data is added to al_comb
         #requested_amount, 
         population, cities_served, disadvantaged, project_type, fundable, state)
  
```


## Alaska (NS)
- Waiting on final data

```{r}

```


## Arkansas (Q)

- Applicant data is preprocessed and ready to be merged in at next update.
- TODO: Confirm `project_cost` should be used for `funding_amount` in both Applicant and Fundable.
- Mention of *Interest Rates or Loan Terms* in the fundable list


```{r}
# the manually merged data from MDI includes fundable projects with columns from the comprehensive list
# this is a combination of charts 3 and 4 from the IUP
# (15,21)
ar_merge <- fread("data/year1/csv/4-Arkansas_Merged.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# the comprehensive PPL includes applicant projects, (747,13)
ar_comp <- fread("data/year1/csv/4-Arkansas.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

ar_app <- ar_comp %>%
  ## get rid of rows that the rank is not a number (not a project, total columns, etc)
  mutate(state_rank = as.numeric(no)) %>%
  filter(!is.na(state_rank)) %>%
  ## process numeric columns
  mutate(
         ## getting rid of $ and comma and replacing blanks with zero         
         population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(project_cost,"[^0-9.]", "")),
         ) %>%
  mutate(
        # fix one-off NAs for pwsid
        pws_id = case_when(
          pws == "FORT SMITH WATER  UTILITIES" ~ "507",
          TRUE ~ pws_id
        ),
        pwsid = case_when(
          # pws_id is missing leading zeros, so standardize length to 9 characters based on length
           nchar(pws_id) == 1 ~ paste0("AR000000", pws_id), 
           nchar(pws_id) == 2 ~ paste0("AR00000", pws_id),
           nchar(pws_id) == 3 ~ paste0("AR0000", pws_id),
           nchar(pws_id) == 4 ~ paste0("AR000", pws_id)),
         # project types are denoted by X's in their respective column
         project_type = case_when(
           lsl == "X" ~ "Lead",
           ec == "X" ~ "Emerging Contaminants",
           TRUE ~ "Other"),
        # disadv column is either YES, NO, or NA by default
         disadvantaged = case_when(
           disadv_antaged_y_n == "YES" ~ "Yes",
           disadv_antaged_y_n == "NO" ~ "No",
           TRUE ~ as.character(NA)),
         borrower = str_squish(pws),
         state_score = gsub(",", "", total_points),
         # all projects in ppl are applicants, see Merged for fundable projects
         fundable = "Applicant",
         state = "Arkansas"
         ) %>%
  ## keep relevant columns
  select(borrower, pwsid, project_description, project_type,
         state_rank, state_score, funding_amount, 
         disadvantaged, population, fundable, state)


## Process Fundable
ar_fund <- ar_merge %>% 
  clean_names() %>%
  select(-b_c_date_actual_estimated, -term_in_years, -interest_rate,
         -green_project_reserve_amt_estimate, -gpr_category_estimate,
         -mhi, -small_system_y_n) %>%
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(no,"[^0-9.]", "")),
         population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(project_cost,"[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(additional_subsidy,"[^0-9.]", ""))
  ) %>%
  # process text columns
  mutate(borrower = str_squish(pws),
         pwsid = case_when(
           nchar(pws_id) == 2 ~ paste0("AR00000", pws_id),
           nchar(pws_id) == 3 ~ paste0("AR0000", pws_id),
           nchar(pws_id) == 4 ~ paste0("AR000", pws_id)),
         project_description = str_squish(project_description),
         state_score = str_replace_all(total_points,"[^0-9.]", ""),
         project_type = case_when(lsl == "X" ~ "Lead",
                                  ec == "X" ~ "Emerging Contaminants",
                                  TRUE ~ "Other"),
         disadvantaged = str_to_sentence(disadvantaged_community),
         # all projects in ppl are applicants, see Merged for fundable projects
         fundable = "Fundable",
         state = "Arkansas",
         ) %>%
  select(state_rank, state_score, borrower, pwsid, project_description, 
         funding_amount, principal_forgiveness_amount, population, disadvantaged, project_type, fundable, state)

# expect (760,12)
ar_clean <- bind_rows(ar_app, ar_fund)

```


## Arizona (NS)
- Waiting on final data

```{r}

```


## California (DD)

- Applicant data has been added in. 
- TODO: Possible we might extract some Disadvantaged info from COMP that isn't currently in FUND

New Columns:
- Does not have `requested_amount`, but does have `estimated_project_cost` that could be included
- EC mentions `requested_amount` but it is used for funding amount
- Applicant data treats `estimated_requested_project_costs` as project_costs to be consistent with the funded project data
- No mention of *Interest Rates or Loan Terms*

```{r}

### FUNDABLE ###

# (16,12) -> (15,9)
# intended to be funded projects that are not EC
ca_fund <- fread("data/year1/csv/5-California_Fundable_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # get rid of non-project columns
  filter(!grepl("Total", project_number)) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         # no longer using service connections, but can be added back in if expanded
         # service_connections = as.numeric(str_replace_all(service_connections,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(estimated_dwsrf_loan,"[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(estimated_pf_grant_amount,"[^0-9.]", "")),
         project_cost = as.numeric(str_replace_all(estimated_project_costs, "[^0-9.]", ""))
  ) %>%
  # process text columns
  # pwsid is CA + first seven characters of project number
  mutate(pwsid = paste0("CA", as.character(map(strsplit(project_number, split = "-"), 1))),
         borrower = str_squish(applicant),
         project_name = str_squish(project_number),
         project_description = str_squish(project_title_description),
         project_type = "Other",
         # it is possible disadvantaged is denoted in the Class column, but the IUP does not define it explicitly
         # disadvantaged = ,
         state = "California",
         fundable = "Fundable"
         ) %>%
  select(borrower, pwsid, project_name, project_description, funding_amount, principal_forgiveness_amount, 
         project_type, project_cost, population, state, fundable)


# (11,9) -> (11,10)
# EC projects that have funding data
ca_ec <- fread("data/year1/csv/5-California_EC.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(requested_funding,"[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(estimated_maximum_pf_grant_amount,"[^0-9.]", "")),
         ) %>%
  # process text columns
  mutate(pwsid = paste0("CA", as.character(map(strsplit(project_number, split = "-"), 1))),
         borrower = str_squish(applicant),
         project_name = str_squish(project_number),
         project_description = str_squish(project_title_description),
         project_type = "Emerging Contaminants",
         disadvantaged = case_when(
           grepl("Not", degree_of_disadvantaged) ~ "No",
           TRUE ~ "Yes"
         ),
         state = "California",
         fundable = "Fundable") %>%
  select(borrower, pwsid, project_name, project_description, funding_amount, principal_forgiveness_amount, 
         project_type, population, disadvantaged, state, fundable)


# -> (26,11)
ca_merge <- bind_rows(ca_fund, ca_ec)

### APPLICANT ### 

# (345, 11) -> (313, 8) when dropping non-project rows and likely to be funded projects and consolidating columns
ca_app <- fread("data/year1/csv/5-California_Comprehensive_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # drop non-project rows
  filter(!is.na(applicant) & applicant != "Applicant" & !is.na(project_number2)) %>%
  # drop fundable projects already processed
  filter(!project_number2 %in% ca_merge$project_name) %>%
  # process numeric columns
  mutate(
    project_cost = as.numeric(str_replace_all(estimated_requested_project_costs,"[^0-9.]", "")),
    population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
  ) %>%
  # process text columns
  mutate(
    borrower = str_squish(applicant),
    project_name = str_squish(project_number2),
    project_description = str_squish(project_title_description),
    # classify different categorical variables of DAC status
    degree_of_disadvantaged = str_squish(degree_of_disadvantaged),
    disadvantaged = case_when(
      degree_of_disadvantaged == "DAC" | degree_of_disadvantaged == "DAC" | degree_of_disadvantaged == "SDAC" | degree_of_disadvantaged == "Large Disadvantaged" | degree_of_disadvantaged == "Large Severely Disadvantaged" ~ "Yes",
      degree_of_disadvantaged == "non-DAC" | degree_of_disadvantaged == "Not Disadvantaged" ~ "No",
      degree_of_disadvantaged == "Pending" ~ "No Information",
      TRUE ~ "No Information"
    ),
    project_type = case_when(
      grepl("Lead", project_title_description, ignore.case=TRUE) ~ "Lead",
      TRUE ~ "Other"
    ),
    fundable = "Applicant",
    state = "California"
  ) %>%
  select(borrower, project_name, project_description, project_type, project_cost, population, 
         disadvantaged, state, fundable)


# (339, 12)
ca_clean <- bind_rows(ca_merge, ca_app)

```


## Colorado (DD) (A)

- TODO Applicant data
- Mentions *Interest Rates or Loan Terms*

```{r}

## Appendix B

# (165,11)
co_b <- fread("data/year1/csv/7-Colorado_AppendixB.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()


# currently assumes that all funding should be added into a single funding column because, "Specific PF amounts are not available in this IUP, probably because p36 states, "Eligibility for BIL PF will be determined when a pre-qualification application is submitted." Also, on p6 they mention that "Amounts available will vary, and at times, may not be available."
co_aggregate <- co_b %>%
  # process numeric columns for aggregating
  mutate(funding_amount = as.numeric(str_replace_all(approved_loan_amount,"[^0-9.]", "")),
         ) %>%
  # total different sources of funding by project ID
  group_by(project_number) %>%
  summarize(total_funding = sum(funding_amount, na.rm=TRUE))
  
co_b_clean <- co_b %>%
  left_join(co_aggregate) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(pop,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(total_funding,"[^0-9.]", "")),
         project_cost = as.numeric(str_replace_all(estimated_project_cost,"[^0-9.]", "")),
    ) %>%
  # process text columns
  mutate(pwsid= str_squish(pws_id_number),
         project_name = str_squish(project_number),
         borrower = str_squish(facility),
         cities_served = str_squish(county),
         state_score = str_replace_all(pts,"[^0-9.]", ""),
         disadvantaged = case_when(
           dac == "Y" ~ "Yes",
           TRUE ~ "No"
         ),
         project_type = "Other",
         project_description = str_squish(project_description),
         state = "Colorado",
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant")) %>%
  select(borrower, pwsid, project_name, project_description, state_score, project_cost,
         funding_amount, population, cities_served, disadvantaged, project_type, state, fundable) %>%
  distinct()

  
# There are currently some rows with NA project_descriptions that need to be removed where they are duplicates, but not all NA project descriptions should be dropped as some are simply missing from the data.
co_antijoin <- co_b_clean %>%
  filter(is.na(project_description) & project_name %in% c("132321D-Q", "140401D-M", "140421D-I", "140521D-Q", 
                                                           "140771D-I", "140951D-I", "142361D-Q", "230340D")) 

# (-> 134,13) # should be same number as when aggregated by project number and is
co_b_clean <- co_b_clean %>%
  anti_join(co_antijoin) %>%
  # then, get rid of one project that is duplicated because of cut/cropped project descriptions from scraping
  mutate(
    project_description = case_when(
      project_name == "132321D-Q" ~ "New Water Treatment Facilities; Improvement/Expansion of Water Treatment Facilities; Connection to a New or Existing Water Treatment Facility; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters; Source Water Protection Plan; Green Infrastructure",
      project_name == "140831D-Q" ~ "Improvement/Expansion of Water Treatment Facilities; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Rights",
      project_name == "141951D-I" ~ "Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Meters; Green Infrastructure",
      project_name == "141970D" ~ "Improvement/Expansion of Water Treatment Facilities; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters; Source Water Protection Plan",
      project_name == "142130D" ~ "Improvement/Expansion of Water Treatment Facilities; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters",
      project_name == "170361D-Q" ~ "New Water Treatment Facilities; Improvement/Expansion of Water Treatment Facilities; Connection to a New or Existing Water Treatment Facility; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters; Source Water Protection Plan",
      project_name == "230450D" ~ "Connection to a New or Existing Water Treatment Facility; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Meters",
      project_name == "142220D" ~ "Improvement/Expansion of Water Treatment Facilities; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters; Source Water Protection Plan; Green Infrastructure",
      project_name == "142341D-Q" ~ "New Water Treatment Facilities; Improvement/Expansion of Water Treatment Facilities; Consolidation of Water Treatment Facilities; Connection to a New or Existing Water Treatment Facility; Construction or Rehabilitation of Distribution and/or Transmission Lines; Water Storage Facilities; Water Supply Facilities; Water Meters; Water Rights; Source Water Protection Plan; Green Infrastructure",
      TRUE ~ project_description)
    ) %>%
  distinct()
  
  

## Appendix B.1

# (104,13)
co_b1 <- fread("data/year1/csv/7-Colorado_AppendixB1.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (104,13)
co_b1_clean <- co_b1 %>%
  # drop unused columns
  select(-term_yrs, -loan_type, -interest_rate) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(pop,"[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(estimated_project_cost,"[^0-9.]","")),
          # Assumes estimated_project_cost should be project_cost, not funding_amount,
         # because while the projects are fundable, the IUP does not make it clear which ones will be funded,
         # only that these projects expressed that they would be replacing lead service lines with their project funding
         funding_amount = as.numeric(NA)
         ) %>%
  # process text columns
  mutate(project_name = str_squish(project_number),
         state_score = str_squish(pts),
         borrower = str_squish(facility),
         cities_served = str_squish(county),
         pwsid = str_squish(pws_id_number),
         project_description = str_squish(project_description),
         # Assumes 'BIL' in 'dac' means they are disadvantaged, see p36 of the IUP
         disadvantaged = "Yes",
         fundable = "Fundable",
         state = "Colorado",
         project_type = "Lead") %>%
  # select relevant states
  select(state_score, borrower, pwsid, project_name, project_description, project_cost,
         funding_amount, project_type, cities_served, population, disadvantaged, state, fundable)

## Appendix B.2

# (26,13)
co_b2 <- fread("data/year1/csv/7-Colorado_AppendixB2.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (26,13)
co_b2_clean <- co_b2 %>%
  # drop unused columns
  select(-term_yrs, -loan_type, -interest_rate) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(pop,"[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(estimated_project_cost,"[^0-9.]","")),
         # Assumes estimated_project_cost should be project_cost, not funding_amount,
         # because while the projects are fundable, the IUP does not make it clear which ones will be funded,
         # only that these projects expressed that they would be replacing lead service lines with their project funding
         funding_amount = as.numeric(NA),
         ) %>%
  # process text columns
  mutate(project_name = str_squish(project_number),
         state_score = str_squish(pts),
         borrower = str_squish(facility),
         cities_served = str_squish(county),
         pwsid = str_squish(pws_id_number),
         project_description = str_squish(project_description),
         # Assumes 'BIL' in 'dac' means they are disadvantaged, see p36 of the IUP
         disadvantaged = "Yes",
         fundable = "Fundable",
         state = "Colorado",
         project_type = "Emerging Contaminants") %>%
  # select relevant states
  select(state_score, borrower, pwsid, project_name, project_description, project_cost,
         funding_amount, project_type, cities_served, population, disadvantaged, state, fundable)
  

# -> (264,13)
co_clean <- bind_rows(co_b_clean, co_b1_clean, co_b2_clean)


```



## Connecticut (NS)
- To be turned into a data dictionary by Water Team

```{r}

```


## Delaware

- Applicant data is processed and ready to add
- Mentions *Interest Rates or Loan Terms*

```{r}

## This code was used to test the degree of overlap between ppl and fund, kept for posterity
# de_ppl_test <- fread("data/year1/csv/8-Delaware_PPL.csv",
#                       colClasses = "character", na.strings = "") %>%
#   clean_names() %>%
#   mutate(rank = as.numeric(str_replace_all(rank,"[^0-9.]","")),
#          amount = as.numeric(str_replace_all(amount,"[^0-9.]","")),
#          ) %>%
#   select(rank, amount)
# 
# ## FUNDED PROJECTS
# 
# de_fund_test <- fread("data/year1/csv/8-Delaware_PPL_fund.csv",
#                       colClasses = "character", na.strings = "") %>%
#   clean_names() %>%
#   mutate(rank = as.numeric(str_replace_all(rank,"[^0-9.]","")),
#          amount = as.numeric(str_replace_all(amount,"[^0-9.]","")),
#          fundable = "Fundable"
#          )
# de_test <- merge(de_fund_test, de_ppl_test, all=TRUE)


## APPLICANT LIST
# only three projects on this list aren't already on the fundable list, so 
# we just need to identify those projects and then subset to them
de_ppl <- fread("data/year1/csv/8-Delaware_PPL.csv",
                       colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  filter(
    (rank == "43" & water_system_borrower == "Artesian") |
    (rank == "44" & water_system_borrower == "Artesian") | 
    (rank == "39" & water_system_borrower == "Artesian")  
  ) %>%
  mutate(fundable = "Applicant") %>%
  rename(project_name = comprehensive_project_name)

## FUNDED LIST
de_fund <- fread("data/year1/csv/8-Delaware_PPL_fund.csv",
                       colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable = "Fundable") %>%
  rename(project_name = fundable_project_name)

# combine comprehensive and fundable lists
de_all <- bind_rows(de_fund, de_ppl)

de_clean <- de_all %>%
  ## rowwise operations for sums
  rowwise() %>%
  ## make relevant columns numbers
  mutate(state_rank = as.numeric(str_replace_all(rank,"[^0-9.]","")),
         population = as.numeric(str_replace_all(population_served,"[^0-9.]","")),
         ## sum of columns, getting rid of everything besides numbers and periods
         ## replace NA with 0 by using sum even if only summing one number
         funding_amount = sum(as.numeric(str_replace_all(amount,"[^0-9.]", "")),
                              na.rm = T),
         ## difference, getting rid of everything besides numbers and .
         principal_forgiveness_amount = sum(as.numeric(str_replace_all(anticipated_subsidy_amount,"[^0-9.]","")),
                                            na.rm = T),
         ) %>%
  ungroup() %>%
  ## new columns or text columns that need functions
  mutate(project_type = case_when(grepl("lead|lsl", project_name, ignore.case = T) ~
                                    "Lead",
                                  grepl("pfas", project_name, ignore.case = T) ~
                                    "Emerging Contaminants",
                                  TRUE ~ "Other"),
         state_score = str_replace_all(total_points,"[^0-9.]",""),
         fundable = "Fundable",
         disadvantaged = case_when(!is.na(dac_a_e_w_u) ~ "Yes",
                                   TRUE ~ "No"),
         pwsid_number = str_replace(pwsid_number, "Delmar DE0000567 Holy Oak               DE0000568", "DE0000567"),
         borrower = str_squish(water_system_borrower),
         pwsid = str_squish(pwsid_number),
         project_name = str_squish(project_name),
         project_description = str_squish(project_description),
         state = "Delaware"
         ) %>%
  ## keep relevant columns
  select(borrower, pwsid, project_name, project_description, project_type,
         state_rank, state_score, funding_amount, principal_forgiveness_amount,
         disadvantaged, population, fundable, state)
```


## District of Columbia

- Don't currently have Applicant data, may need to track down, if possible
- No mention of `Project Cost` or `Requested Amount`
- No mention of *Interest Rates or Loan Terms*

```{r}

# 13,3
dc_ppl <- fread("data/year1/csv/DistrictofColumbia_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (13,5)
dc_clean <- dc_ppl %>%
  mutate(
  project_description = str_squish(project),
  funding_amount = as.numeric(str_replace_all(grant_funding,"[^0-9.]","")),
  state = "District of Columbia",
  project_type = "Other",
  fundable = "Fundable"
  ) %>%
  select(project_description, funding_amount, project_type, state, fundable)

```


## Florida (DD - A)

- TODO: Applicant data - awaiting answer on whether it is the "waiting" list within the iup or not.
- No mention of *Interest Rates or Loan Terms*

```{r}

fl_fund <- fread("data/year1/csv/9-Florida_ppl_fundable.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # format numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(authorized_loan_amount,"[^0-9.]","")),
         principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness_amt,"[^0-9.]","")),
         ) %>%
  # formant text columns
  mutate(borrower = str_squish(applicant),
         project_name = str_squish(project_number),
         project_description = str_squish(project_description),
         state_score = str_replace_all(priority_score,"[^0-9.]",""),
         project_type = "Other",
         state = "Florida",
         fundable = "Fundable") %>%
  select(borrower, project_name, project_description, state_score, funding_amount, 
         principal_forgiveness_amount, project_type, state, fundable)
  

```


## Georgia

- Applicant data is preprocessed and ready to be merged in at next update.
- Includes `project_cost`
- Mentions *Interest Rates or Loan Terms*

```{r}

## BASE IUP
# (60,14)
ga_raw <- fread("data/year1/csv/10-Georgia_Comp_clean.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# (60,14) -> (60,8)
ga_base <- ga_raw %>%
  # remove columns
  select(-est_notice_to_proceed, -est_construction_start, -est_construction_completion, -
           est_interst_rate, -est_terms) %>%
  # process numeric columns
  mutate(
         population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(total_disburs, "[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(total_project_cost, "[^0-9.]","")),
  ) %>%
  # process text columns
  mutate(borrower = str_squish(community),
         project_description = str_squish(project_description),
         state_score = str_replace_all(project_score,"[^0-9.]",""),
         project_type = case_when(
           grepl("Lead", project_description, ignore.case=TRUE) ~ "Lead",
           grepl("PFAS", project_description, ignore.case=TRUE) ~ "Emerging Contaminants",
           TRUE ~ "Other"),
         # Affordability Score in Attachment 1: 2022 Comprehensive List in PPL is 29 or above.
         disadvantaged = case_when(
          affordability_score >= 29 ~ "Yes",
           TRUE ~ "No"
         ),
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         state = "Georgia"
  ) %>% 
  select(state_score, borrower, project_description, funding_amount, project_cost,
         population, disadvantaged, fundable, project_type, state)



## SUPPLEMENTAL IUP

ga_supp_1 <- fread("data/year1/csv/10-Georgia_Supplemental_IUP_A1.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

ga_supp_2 <- fread("data/year1/csv/10-Georgia_Supplemental_IUP_A2.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(project, total_disburs) %>%
  # manually standardize names for matching
  mutate(project = str_replace(project, "Clayton County Water Authority", "Clayton County Water Authority (CCWA)"),
         total_disburs = str_squish(total_disburs)) %>%
  rename(community = project)

ga_supp <- ga_supp_1 %>%
  left_join(ga_supp_2) %>%
  # remove columns
  select(-est_notice_to_proceed, -est_construction_start, -est_construction_completion, -
           est_interst_rate, -est_terms) %>%
  # process numeric columns
  mutate(
         population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(total_disburs, "[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(total_project_cost, "[^0-9.]","")),
  ) %>%
  # process text columns
  mutate(borrower = str_squish(community),
         project_description = str_squish(project_description),
         state_score = str_replace_all(project_score,"[^0-9.]",""),
         project_type = case_when(
           grepl("Lead", project_description, ignore.case=TRUE) ~ "Lead",
           grepl("PFAS", project_description, ignore.case=TRUE) ~ "Emerging Contaminants",
           TRUE ~ "Other"),
         # Affordability Score in Attachment 1: 2022 Comprehensive List in PPL is 29 or above.
         disadvantaged = case_when(
          affordability_score >= 29 ~ "Yes",
           TRUE ~ "No"
         ),
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         state = "Georgia"
  ) %>% 
  select(state_score, borrower, project_description, funding_amount, project_cost,
         population, disadvantaged, fundable, project_type, state)
  


## LEAD IUP
ga_lead_1 <- fread("data/year1/csv/10-Georgia_Lead_IUP_A1.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

ga_lead_2 <- fread("data/year1/csv/10-Georgia_Lead_IUP_A2.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(project, total_disburs) %>%
  # manually standardize names for matching
  mutate(
    project = str_squish(str_replace(project, "Bartow County Water", "Bartow County Water Department")),
    project = str_squish(str_replace(project, "Lee County Water Authority", "Lee County Utility Authority")),
    project = str_squish(str_replace(project, "City of Monroe", "Monroe County")),
    project = str_squish(str_replace(project, "Cobb County Water Authority", "Cobb County Water System")),
    project = str_squish(str_replace(project, "Jackson County Water Authority", "Jackson County Water and Sewerage Authority")),
    total_disburs = str_squish(total_disburs)) %>%
  rename(community = project)

ga_lead <- ga_lead_1 %>%
  # remove asterisks for joining second table by community name
  mutate(community = str_squish(str_remove_all(community, "\\*"))) %>%
  full_join(ga_lead_2) %>%
  # process numeric columns
  mutate(
         population = as.numeric(str_replace_all(x2020_pop,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(total_disburs, "[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(total_project_cost, "[^0-9.]","")),
  ) %>%
  # process text columns
  mutate(borrower = str_squish(community),
         project_description = str_squish(project_description),
         state_score = str_replace_all(affordability_score,"[^0-9.]",""),
         project_type = "Lead",
         # Affordability Score in Attachment 1: 2022 Comprehensive List in PPL is 29 or above.
         disadvantaged = case_when(
          as.numeric(state_score) >= 29 ~ "Yes",
           TRUE ~ "No"
         ),
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         state = "Georgia"
  ) %>% 
  select(state_score, borrower, project_description, funding_amount, project_cost,
         population, disadvantaged, fundable, project_type, state)


## EC IUP
# (2,10)
ga_ec_1 <- fread("data/year1/csv/10-Georgia_EC_IUP_A1.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# (2,10) -> (2,2)
ga_ec_2 <- fread("data/year1/csv/10-Georgia_EC_IUP_A2.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(project, total_disburs) %>%
  # prep for merging with appendix 1 table
  rename(community = project) %>%
  mutate(community = str_squish(str_remove_all(community, "\\*")))

# process and merge GA EC tables
# -> (2,10)
ga_ec <- ga_ec_1 %>%
  left_join(ga_ec_2) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(x2019_pop,"[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(total_project_cost,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(total_disburs,"[^0-9.]","")),
         ) %>%
  # process text columns
  mutate(borrower = str_squish(community),
         project_description = str_squish(project_description),
         project_type = "Emerging Contaminants",
         state_score = str_replace_all(affordability_score,"[^0-9.]",""),
         # Affordability Score in Attachment 1: 2022 Comprehensive List in PPL is 29 or above.
         disadvantaged = case_when(
          as.numeric(state_score) >= 29 ~ "Yes",
           TRUE ~ "No"
         ),
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         state = "Georgia"
  )%>%
  select(state_score, borrower, project_description, funding_amount, project_cost,
         population, disadvantaged, fundable, project_type, state)


# join four tables together
# NOTE: for funding_amount we would sum together values from base and supplemental, but there are no projects in common on both lists, so summing is not necessary, we can just bind the rows together.
# -> (323,10)
ga_clean <- bind_rows(ga_base, ga_supp, ga_lead, ga_ec)

```


## Hawaii

- Applicant data is processed and ready to be added in future update.
- Does not include a `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (33,8)
# note this if from the july 22 IUP, but it is the same in the March 23 amended IUP
# https://health.hawaii.gov/sdwb/files/2023/03/SFY2023.IUP_Amendment1_2023.03.pdf
hi_fund <- fread("data/year1/csv/11-Hawaii_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (33,10)
hi_clean <- hi_fund %>%
  ## process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(priority_ranking,"[^0-9.]","")),
         population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         requested_amount = as.numeric(str_replace_all(estimated_eligible_amount,"[^0-9.]", "")),
         # when project is in funded list, treat estimated_eliglbe_amount as funding_amount as well as requested_amount
         funding_amount = case_when(
           state_rank %in% c(1,2,6,11,13,16,20,21,29,33) ~ requested_amount,
           TRUE ~ 0
         ),
  ## process text columns       
         # manually add disadvantaged based on subscript from IUP that couldn't be scraped
         # denoted by "E" footnote subscript in fundable list, table 7 in IUP
         # projects that are listed are either DAC or not, but full applicant list is No Information because they don't denote disadvantaged in the full table in Appendix A
          disadvantaged = case_when(state_rank %in% c(1,2,11,13,20,21) ~ "Yes",
                                    state_rank %in% c(6, 16, 29, 33) ~ "No",
                                    TRUE ~ as.character(NA)),
          project_name = str_squish(project_name),
          borrower = str_squish(borrower),
          # manually mark projects as funded from amended IUP
          fundable = case_when(funding_amount > 0 ~ "Fundable",
                               TRUE ~ "Applicant"),
          project_type = "Other",
          state = "Hawaii"
         ) %>%
  ## keep relevant columns
  select(borrower, project_name,
         state_rank, requested_amount, funding_amount, project_type,
         disadvantaged, population, fundable, state)

# amendments - in the march 2023 IUP, a number of changes are highlighted in yellow
hi_clean <- hi_clean %>%
  mutate(
    # manually modify funding amounts that are highlighted yellow in amendment table
    funding_amount = case_when(
      state_rank == 1 ~ 10000000,
      state_rank == 11 ~ 1572484.20,
      state_rank == 12 ~ 225964,
      state_rank == 20 ~ 1770889.20,
      state_rank == 21 ~ 165867.30,
      state_rank == 29 ~ 459731.84,
      # remove three projects scratched through
      state_rank %in% c(2, 6, 33) ~ 0,
      TRUE ~ funding_amount),
    # manually modifying yellow highlighted project descriptions
    project_name = case_when(
      state_rank == 21 | state_rank == 29 ~ "Backup Generator and SCADA Upgrades",
      TRUE ~ project_name),
    fundable = case_when(
      # set three scratched through projects as not funded
      state_rank %in% c(2, 6, 33) ~ "Applicant",
      TRUE ~ fundable)
    )
```


## Idaho
- Applicant data is preprocessed and ready to be merged in at next update.
- `project_cost` included, no mention of `requested_amount``
- id_fund mentions *Interest Rates or Loan Terms*

```{r}

# (81, 9)
id_ppl <- fread("data/year1/csv/12-Idaho_PPL.csv", 
                      colClasses = "character", na.strings = "") %>% 
          clean_names()

# (13,11) -> (13,4)
# reduce fundable projects to only their rank (a UID in this case), 
# principal forgiveness amount, and loan amount
id_fnd <- fread("data/year1/csv/12-Idaho_Fundable.csv",
                      colClasses = "character", na.strings = "") %>%
        clean_names() %>%
        mutate(rank = str_replace_all(rank,"[^0-9.]", ""),
               loan_amt = as.character(map(strsplit(loan_amt_est_loan_date, split = "  "), 1)),
               proposed_funding_terms = str_squish(proposed_funding_terms),
               # see NOTE for disadv definition
               disadvantaged = case_when(
                  grepl("30 years", proposed_funding_terms) ~ "Yes",
                  grepl("principal forgiveness", proposed_funding_terms, ignore.case = TRUE) ~ "Yes",
                  grepl("PF", proposed_funding_terms) ~ "Yes",
                  TRUE ~ "No"),
               # extract PF value and format as numeric
               principal_forgiveness_amount = as.character(map(strsplit(proposed_funding_terms, split = "with"), 2)),
               principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness_amount, "[^0-9.]", "")),
               principal_forgiveness_amount = replace_na(principal_forgiveness_amount, 0)
               ) %>%
        select(rank, loan_amt, principal_forgiveness_amount, disadvantaged)
        
# (81,12)
id_merged <- merge(id_ppl, id_fnd, by=c("rank"), all.x=TRUE)

# -> (81,13)
id_clean <- id_merged %>%
  # process numeric columns
  mutate(state_rank = as.numeric(rank),
         population = as.numeric(str_replace_all(pop_served, "[^0-9.]", "")),
         project_cost = as.numeric(str_replace_all(project_cost, "[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(loan_amt, "[^0-9.]", "")),
         # ASSUMPTION: 0 different from NA, removing these steps until suggested otherwise
         # funding_amount = replace_na(funding_amount, 0),
         # principal_forgiveness_amount = replace_na(principal_forgiveness_amount, 0)
  ) %>%
  # process text columns
  mutate(project = str_squish(project),
         # remove excess spaces from scrape
         regional_office = str_squish(regional_office),
         state_score = str_replace_all(rating_points, "[^0-9.]", ""),
         project_description = str_squish(project_description),
         # custom touch-up squishing doesn't fix
         project_description = str_replace(project_description, "C onstruction", "Construction"),
         # if from fundable table, else Applicant
         fundable = case_when(
           !is.na(funding_amount) ~ "Fundable",
           TRUE ~ "Applicant"),
         state="Idaho",
         # only project with Lead is the "All System fund"
         project_type = case_when(
           grepl("Lead", project) ~ "Lead",
           TRUE ~ "Other"),
         pwsid = str_replace(system_number, "Unknown", as.character(NA)),
         pwsid = str_replace(pwsid, "All", as.character(NA))
         ) %>%
  rename(borrower = project) %>%
  select(state_rank, state_score, borrower, pwsid, project_description, population, project_cost,
         disadvantaged, funding_amount, fundable, principal_forgiveness_amount, project_type, state)

```


## Illinois

- Applicant data is processed
- No mention of `project_cost` or `requested_amount``
- No mention of *Interest Rates or Loan Terms*

```{r}
## Read in the manually separated versions of the PPL and lead data 
## from the excel spreadsheets sent by the state

# (70,11)
il_ppl_f <- fread("data/year1/csv/13-Illinois_PPL_Fundable.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Fundable",
         project_type="Other",
         # replace "N/E" with 0 for principal forgiveness since N/E is different from the NAs in other docs
         disadvantaged_community_principal_forgiveness = str_replace(disadvantaged_community_principal_forgiveness, "N/E", "0")
           )

# (19,11)
il_ppl_a <- fread("data/year1/csv/13-Illinois_PPL_Applicant.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Applicant",
         project_type="Other")

# concat files of the same structure
# determine disadvantaged based on defined PF column, to use in merging with lead below
# (89,11)
il_ppl <- bind_rows(il_ppl_a, il_ppl_f) %>%
    mutate(disadvantaged = case_when(
           as.numeric(str_replace_all(disadvantaged_community_principal_forgiveness, "[^0-9.]", "")) > 0 ~ "Yes",
           TRUE ~ "No"))

# create (85,2) list of communities and DAC status
dacs <- il_ppl %>%
  select(loan_applicant, disadvantaged) %>%
  distinct()

# Preprocess Lead Projects
# (18,10)
il_lead_f <- fread("data/year1/csv/13-Illinois_Lead_Fundable.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Fundable",
         project_type="Lead")

# (1,10)
il_lead_a <- fread("data/year1/csv/13-Illinois_Lead_Applicant.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Applicant",
         project_type="Lead")

# merge lead projects together
# (19,10)
il_lead <- bind_rows(il_lead_a, il_lead_f)

# merge lead projects with dac list
il_lead <- merge(il_lead, dacs, all.x=TRUE, by="loan_applicant") %>%
  mutate(disadvantaged = replace_na(disadvantaged, "No Information") )


# (108, 11)
il_merge <- bind_rows(il_ppl, il_lead)

# -> (108,11)
il_clean <- il_merge %>%
  # drop columns
  select(-l17_number, -estimated_construction_start) %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(estimated_loan_amount, "[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(disadvantaged_community_principal_forgiveness, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(service_population, "[^0-9.]", "")),
         ) %>%
  # process text columns
  mutate(borrower = str_squish(loan_applicant),
         borrower = str_to_title(borrower, locale="en"),
         state_score = str_replace_all(loan_priority_score, "[^0-9.]", ""),
         project_description = str_squish(project_description),
         project_description = str_to_sentence(project_description),
         state = "Illinois"
  ) %>%
  # rename columns
  rename(pwsid = facility_no) %>%
  # keep relevant columns
  select(state_score, borrower, pwsid, project_description, population, disadvantaged, funding_amount, principal_forgiveness_amount, project_type, fundable, state)

```


  ## Indiana (DD)

- Assumes `requested_amount` to be included

```{r}

## PPL
# (73,17)
in_ppl <- fread("data/year1/csv/14-Indiana_Q4Final_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

## Lead
# (9,17)
in_lead_ppl <- fread("data/year1/csv/14-Indiana_Q4Final_LeadPPL.csv",
                 colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (82,17)
in_comb <- bind_rows(in_ppl, in_lead_ppl)


# -> (82,12)  
in_clean <- in_comb %>%  
  # drop non-project rows
  filter(participant != "NA") %>%
  # process numeric columns
  mutate(
    population = as.numeric(str_replace_all(population_served, "[^0-9.]", "")),
    requested_amount = as.numeric(str_replace_all(requested_funds, "[^0-9.]", "")),
  
  ) %>%
  # process text column
  mutate(
    borrower = str_squish(participant),
    pwsid = case_when(
      pwsid_no_s == "TBD" ~ "No Information",
      TRUE ~ paste0("IN", pwsid_no_s)),
    project_name = str_squish(srf_project_no),
    project_description = str_squish(project_description),
    state_score = str_replace_all(ppl_score, "[^0-9.]", ""),
    state_rank = str_replace_all(ppl_rank, "[^0-9.]", ""),
    disadvantaged = str_squish(disadvantaged_community),
    project_type = case_when(
      emerging_contaminants == "Yes" ~ "Emerging Contaminants",
      as.numeric(str_replace_all(lead_service_line_replacement_cost, "[^0-9.]", "")) > 0 ~ "Lead",
      TRUE ~ "Other"
    ),
    state = "Indiana",
    fundable = as.character(NA),
  ) %>%
  select(borrower, pwsid, state_rank, state_score, project_name, project_description,
         project_type, requested_amount, population, disadvantaged, project_type, state, fundable)


```


## Iowa (NS)

- We currently have three quarters of data, need to combine, see if projects are listed duplicately?
- Convert excel files to CSV and transform color coding to column variables?
- Need to confirm that this is only applicant data considering there are some marks for loans being signed?

```{r}

```


## Kansas (Q in DD)

- assumes `requested_amount` should be included
- need to untangle the variations of loan_forgiveness variables
- can disadvantaged be assumed from the `loan_forgiveness_disadvantaged_community` in base and lead?
- should PF be defined as sum of the `loan_forigveness` variables?
- is receiving `loan_forgiveness` indicative of being a fundable project?

```{r}

## Base
# -> (98,9)
ks_base <- fread("data/year1/csv/16-Kansas_PPL-base.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(-v3, -v5, -v9, -v10, -v12, -v14, -v15, -v17) %>%
  mutate(
    project_type = "Other"
  )

## Lead
# -> (16,8)
ks_lead <- fread("data/year1/csv/16-Kansas_PPL-lead.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(-v3, -v5, -v7, -v8, -v10, -v11, -v13, -v14, -v16) %>%
  mutate(
    project_type = "Lead"
  ) 

# Emerging Contaminants
# -> (9,7)
ks_ec <- fread("data/year1/csv/16-Kansas_PPL-ec.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(-v3, -v5, -v7, -v8, -v9, -v10, -v12, -v13, -v14, -v16, -v18) %>%
  mutate(
    project_type = "Emerging Contaminants"
  ) %>%
  rename(loan_request = x,
         population_served = served,
         loan_forgiveness = forgiveness)

## Combine
# -> (123,12)
ks_comb <- bind_rows(ks_base, ks_lead, ks_ec)


## Clean
ks_clean <- ks_comb %>%
  # process numeric columns
  mutate(
    requested_amount = as.numeric(str_replace_all(loan_request, "[^0-9.]", "")),
    population = as.numeric(str_replace_all(population_served, "[^0-9.]", ""))
  ) %>%
  # process text columns
  mutate(
    borrower = str_squish(municipality_name),
    project_name = str_squish(project_number),
    state_score = str_squish(rating),
    project_description = str_squish(project_description),
    state = "Kansas",
    # assume all applicant until confirmed otherwise
    fundable = "Applicant"
  ) %>%
  select(borrower, state_score, project_name, project_description, requested_amount,
         population, project_type, state, fundable)

  
```


## Kentucky (NS)
- To be transformed into a data dictionary by Water Team

```{r}

```


## Louisiana

- Applicant data is processed
- No mention of `requested_amount` or `loan_amount`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (40,9)
la_ppl <- fread("data/year1/csv/18-Lousiana_Comprehensive_PPL-clean.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Applicant")

# (8,11)
la_fnd <- fread("data/year1/csv/18-Lousiana_Fundable_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(fundable="Fundable")

# (48,11)
la_comb <- bind_rows(la_ppl, la_fnd)

# -> (46,12)
la_clean <- la_comb %>%
  # drop columns and total rows
  select(-est_date_to_close_loan) %>%
  filter(system_name !="Total") %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(est_loan_amount, "[^0-9.]", "")),
         state_rank = as.numeric(str_replace_all(rank, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(population, "[^0-9.]", "")),
         disadvantaged_subsidy = as.numeric(str_replace_all(disadvantaged_subsidy, "[^0-9.]", "")),
         # replace 0s specifically for adding columns together
         disadvantaged_subsidy = replace_na(disadvantaged_subsidy, 0),
         additional_subsidy_amount = as.numeric(str_replace_all(additonal_subsidy_amount, "[^0-9.]", "")),
         additional_subsidy_amount = replace_na(additional_subsidy_amount, 0),
         principal_forgiveness_amount = disadvantaged_subsidy + additional_subsidy_amount
  ) %>%
  # process text columns
  mutate(borrower = str_squish(system_name),
         pwsid = paste0("LA", pwsid),
         state_score = str_replace_all(points, "[^0-9.]", ""),
         project_description = str_squish(project_description),
         disadvantaged = case_when(
           disadvantaged_subsidy > 0 ~ "Yes",
           TRUE ~ "No"),
         state = "Louisiana",
         project_type = "Other"
  ) %>%
  select(state_rank, state_score, borrower, pwsid, project_description, funding_amount, principal_forgiveness_amount, population, disadvantaged, project_type, fundable, state)
  
```


## Maine (Q in DD)(A)

- TODO: Applicant data - Q in DD about where to interpret it
- TODO: Confirm addition of `amount_requested` from EC/Comp as `funding_amount`
- No mention of *Interest Rates or Loan Terms*

```{r}

# process three projects in separate list for emerging contaminants
me_ec <- fread("data/year1/csv/19-Maine_PL_EC.csv",
                colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # rename columns that don't need to be modified
  rename(borrower = water_system,
         project_name = srf_project_number,
         state_score = total_priority_points) %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(amount_requested, "[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness, "[^0-9.]", ""))
           ) %>%
  # process text columns
  mutate(project_type = "Emerging Contaminants",
         state = "Maine", 
         fundable = "Fundable"
         ) %>%
  select(borrower, project_name, project_description, state_score, funding_amount, 
         principal_forgiveness_amount, project_type, state, fundable)
  

# process comprehensive list
me_comp <- fread("data/year1/csv/19-Maine_PPL_Comprehensive.csv",
                colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # drop columns
  select(-pf_percent, -percent_annual_water_bill_of_mhi, -project_type) %>%
  # process numeric columns
         # subtract ARPA funding from funding amount after setting both to numeric
  mutate(funding_amount = as.numeric(str_replace_all(amount_requested, "[^0-9.]", "")),
         arpa_mjrp = replace_na(arpa_mjrp, "0"),
         arpa_mjrp = as.numeric(str_replace_all(arpa_mjrp, "[^0-9.]", "")),
         funding_amount = funding_amount - arpa_mjrp,
         principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(population_served, "[^0-9.]", ""))
  ) %>%
  # process text columns
  mutate(borrower = str_squish(water_system),
         pwsid = str_squish(pwsid),
         # NOTE: project_description was manually extracted from borrower given inconsistent
         # splitting and spacing throughout the projects.
         project_description = str_squish(project_description),
         project_name = str_squish(srf_project_number),
         state_score = str_squish(total_priority_points),
         # disadvantaged = case_when PF > 10% funding amount
         disadvantaged = case_when(
           principal_forgiveness_amount > funding_amount * .1 ~ "Yes",
           TRUE ~ "No"),
         fundable = "Fundable",
         state = "Maine",
         project_type = "Other",
         ) %>%
  select(borrower, pwsid, state_score, project_name,  project_description, funding_amount, 
         principal_forgiveness_amount,  disadvantaged, project_type, population, state, fundable)

##TODO: row bind EC and clean
me_clean <- bind_rows(me_comp, me_ec)
```


## Maryland (A)

- TODO: Applicant data - https://mde.maryland.gov/programs/water/WQFA/Documents/FFY22%20SFY24%20Final%20DW%20PPL.pdf
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# Due to the format of the Maryland PPL table format, the PPL and Project Funding List are manually combined
# before further cleaning and standardization
# (24,19)
md_raw <- fread("data/year1/csv/20-Maryland_PPL_manual.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (24,14)
md_clean <- md_raw %>%
  #drop columns
  select(-comments, -v16, -system_size_ownership_small_large_public_private, -const_start) %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(dwsrf_total_funding,"[^0-9.]","")),
         dwsrf_base_loan_forgive = as.numeric(str_replace_all(dwsrf_base_loan_forgive,"[^0-9.]","")),
         #replace NA for adding columns
         dwsrf_base_loan_forgive = replace_na(dwsrf_base_loan_forgive, 0),
         bil_loan = as.numeric(str_replace_all(bil_loan,"[^0-9.]","")),
         bil_loan_forgive = as.numeric(str_replace_all(bil_loan_forgive,"[^0-9.]","")),
         bil_loan_forgive = replace_na(bil_loan_forgive, 0),
         principal_forgiveness_amount = bil_loan_forgive + dwsrf_base_loan_forgive,
         state_rank = as.numeric(str_replace_all(priority_rank,"[^0-9.]","")),
         population = as.numeric(str_replace_all(population,"[^0-9.]",""))
  ) %>%
  # process text columns
  mutate(project_name = as.character(map(strsplit(project_name_mde_project_number, split = "\\(DW"), 1)),
         project_name = str_to_sentence(project_name),
         project_name = str_squish(project_name),
         project_description = str_squish(project_description),
         project_description = str_to_sentence(project_description),
         cities_served = str_to_title(county),
         state_score = str_replace_all(scoring_points,"[^0-9.]",""),
         borrower = str_to_title(applicant),
         disadvantaged = str_to_sentence(dac_community),
         pwsid = str_squish(pwsid),
         # manually fix pwsid typos
         pwsid = str_replace(pwsid, "MD006002", "MD0060002"),
         pwsid = str_replace(pwsid, "MD023006", "MD0230006"),
         pwsid = str_replace(pwsid, "MD210010", "MD0210010"),
         fundable = "Fundable",
         project_type = "Other",
         state = "Maryland"
  ) %>%
  select(state_rank, state_score, borrower, pwsid, project_name, project_description, funding_amount,
         principal_forgiveness_amount, cities_served, population, disadvantaged, fundable, project_type, state)

## APPLICANT
# this dataset is 42 projects, the first 24 of which appear to be on the fundable list above. Others will be listed as Not Funded.


```


## Massachusetts (WIP)

- Applicant data is processed
- Mentions `project_cost`, included
- No mention of *Interest Rates or Loan Terms*

```{r}

## ALL PROJECTS
# (120,7) -> (113,7)
ma_comp_raw <- fread("data/year1/csv/21-Massachusetts_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  filter(!is.na(project))

# process data for all non-funding columns first
ma_comp <- ma_comp_raw %>%
  # process numeric columns
  mutate(project_cost = as.numeric(str_replace_all(project_cost,"[^0-9.]", "")),
         population = as.numeric(str_replace_all(pop,"[^0-9.]", "")),
                                   ) %>%
  # process text columns
  mutate(project_name = str_squish(srf_id),
         # get rid of * in some ranks
         state_score = str_squish(str_replace_all(rating,"[^0-9.]", "")),
         pwsid = str_squish(pwsid),
         # remove all caps and drop special characters
         borrower = str_to_title(str_replace_all(applicant, "\\s*\\([^\\)]+\\)", "")),
         borrower = str_squish(borrower),
         project_type = case_when(grepl("(lr)", applicant, ignore.case = T) ~ "Lead",
                                  grepl("(ec)", applicant, ignore.case = T) ~ "Emerging Contaminants",
                                  TRUE ~ "Other"),
         project_description = str_squish(project),
         # DAC defined by receiving PF as denoted in the applicant column
         disadvantaged = case_when(grepl("(pf)", applicant, ignore.case = T) ~ "Yes",
                         TRUE ~ "No"),
         state = "Massachusetts"
         ) %>%
  # select columns
  select(borrower, pwsid, project_name, project_type, project_cost, project_description,
         population, state_score, disadvantaged, state, srf_id)


## FUNDED PROJECTS
# Table 1, listing funded projects in 4 sections, p5-8
# (74,6) -> (68,6)
ma_fund_raw <- fread("data/year1/csv/21-Massachusetts_Fundable.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  ## get rid of title rows
  filter(!is.na(project))


## MERGE ALL PROJECTS
ma_fund <- ma_fund_raw %>%
  # get funding amount, set all projects as funded, keep only new columns and srf_id for matching
  mutate(funding_amount = as.numeric(str_replace_all(iup_cost_2022,"[^0-9.]", ""))) %>%
  mutate(fundable = "Fundable") %>%
  select(srf_id, funding_amount, fundable)

# join funded data to all projects
ma_clean <- left_join(ma_comp, ma_fund) %>%
  # fill created NAs and set funding status
  mutate(funding_amount = replace_na(funding_amount, 0),
         fundable = case_when(
           fundable == "Fundable" ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         # "For projects listed under Planning Projects section of Funding List that include (LR) in Applicant, the amount under 2022IUP Cost is considered Principal Forgiveness." -> 
         # Manually identify projects under the Planning Projects subheading and set their funding_amount as principal_forgiveness_amount or set to 0
         principal_forgiveness_amount = case_when(
           srf_id %in% c("7187", "7082", "7181", "7159", "7174") ~ funding_amount,
           TRUE ~ 0
         )) %>%
  # drop column used for matching
  select(-srf_id)
         


```


## Michigan

- Applicant data processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (112,24)
mi_raw <- fread("data/year1/csv/22-Michigan_PPL.csv",
                      colClasses = "character", na.strings = "")
# -> (110,12)
mi_clean <- mi_raw %>%
  ## clean names
  clean_names() %>%
  ## get rid of rows without projects
  filter(!is.na(project_number)) %>%
  ## rowwise operations for sums
  rowwise() %>%
  ## make relevant columns numbers
  mutate(state_rank = as.numeric(gsub(",", "", rank)),
         population = as.numeric(gsub(",", "", population)),
                  state_score = gsub(",", "",total_points),
         ## sum of columns, getting rid of $ and comma and replacing blanks with zero
         funding_amount = sum(as.numeric(gsub("\\$|,", "", dwsrf_loan)),
                              as.numeric(gsub("\\$|,", "", dwsrf_pf)),
                              as.numeric(gsub("\\$|,", "", bil_dwsrf_supplemental_loan)),
                              as.numeric(gsub("\\$|,", "", bil_dwsrf_supplemental_pf)),
                              as.numeric(gsub("\\$|,", "", bil_dwsrf_emerging_contaminants_pf)),
                              as.numeric(gsub("\\$|,", "", bil_dwsrf_lslr_loan)),
                              as.numeric(gsub("\\$|,", "", bil_dwsrf_lslr_pf)),
                              na.rm = T),
         principal_forgiveness_amount = sum(as.numeric(str_replace_all(total_pf_grant,"[^0-9.]", "")),
                         -as.numeric(str_replace_all(arp_grant, "[^0-9.]", "")),
                         na.rm = T)) %>%
           # sum(as.numeric(str_replace_all(total_pf_grant,"[^0-9 -]", "")),
           #               -as.numeric(str_replace_all(arp_grant,"[^0-9 -]", "")),
           #               na.rm = T)) %>%
  ungroup() %>%
  ## standardize names
  rename(borrower = applicant_name,
         cities_served = project_location) %>%
  ## create project type column
  mutate(project_type = case_when(lslr_costs != "" ~ "Lead",
                                  pfas_costs != "" ~ "Emerging Contaminants"),
         project_type = replace_na(project_type, "Other"),
         ## fundable column
         fundable = case_when(state_rank <= 70 ~ "Fundable",
                                     state_rank > 70 ~ "Applicant"),
         disadvantaged = replace_na(disadvantaged, "No"),
         state = "Michigan"
         ) %>%
  ## keep relevant columns
  select(borrower, cities_served, project_description, project_type,
         state_rank, state_score, funding_amount, principal_forgiveness_amount,
         disadvantaged, population, fundable, state)
```


## Minnesota

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (142, 16)
mn_raw <- fread("data/year1/csv/23-Minnesota_IUP_manual.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# -> (142, 11)
mn_clean <- mn_raw %>%
  # remove extra columns
  select(-est_const_start, v16) %>%
  # process numeric columns
  mutate(
    state_rank = as.numeric(str_replace_all(mdh_2023_ppl_rank,"[^0-9.]","")),
    population = as.numeric(str_replace_all(population,"[^0-9.]","")),
    # transform string to numeric and replace NAs specifically for adding columns
    est_lslr_pf = as.numeric(str_replace_all(estimated_dwrf_lead_service_line_replacement_principal_forgiveness_not_final_1,"[^0-9.]","")),
    est_lslr_pf = replace_na(est_lslr_pf, 0),
    est_ec_pf = as.numeric(str_replace_all(estimated_dwrf_emerging_contaminant_principal_forgiveness_not_final_2,"[^0-9.]","")),
    est_ec_pf = replace_na(est_ec_pf, 0),
    est_dac_pf = as.numeric(str_replace_all(estimated_dwrf_disadvantgd_community_principal_forgiveness_not_final_3,"[^0-9.]","")),
    est_dac_pf = replace_na(est_dac_pf, 0),
    # Sum of Estimated DWRF Lead, Estimated DWRF Emerging Contaminant, 
    # Estimated DWRF Disadvantaged Community
    principal_forgiveness_amount = est_lslr_pf + est_ec_pf + est_dac_pf,
    # transform to numeric to save rows where numbers are present,
    # but fill in 0s from adding above with "No Information" for the final database
    # per State Sources Data Dictionary, funding amount = estimated loan + PFA
    funding_amount = as.numeric(str_replace_all(estimated_dwrf_loan,"[^0-9.]","")) + principal_forgiveness_amount,
  ) %>%
  # process text columns
  mutate(
    borrower = str_squish(system_name),
    project_description = str_squish(project_description),
    state_score = str_replace_all(mdh_2023_ppl_points,"[^0-9.]",""),
    pwsid = paste0("MN", as.character(map(strsplit(project_number, split = "-"), 1))),
    project_type = case_when(
      status == "Part A5,LSL" ~ "Lead",
      status == "Carryover,LSL" ~ "Lead",
      status == "Part A6,EC" ~ "Emerging Contaminants",
      TRUE ~ "Other"),
    # Only projects with a dollar amount indicated for Estimated DWRF Loan and/or Estimated Principal Forgiveness are expected to receive SRF awards at this time and only these projects are included in the dashboard analysis.
    fundable = case_when(
      !is.na(funding_amount) | principal_forgiveness_amount > 0 ~ "Fundable",
      TRUE ~ "Applicant"),
    # Projects with a dollar amount or eligible in any of the estimated principal forgiveness columns or in the estimated WIF grant column are considered DACs.
    disadvantaged = case_when(
      # if not NA or two characters (the space and dash in otherwise empty cells in the raw data), mark as DAC for each column
      !is.na(estimated_dwrf_lead_service_line_replacement_principal_forgiveness_not_final_1) & nchar(  estimated_dwrf_lead_service_line_replacement_principal_forgiveness_not_final_1) > 2 ~ "Yes",
      !is.na(estimated_dwrf_emerging_contaminant_principal_forgiveness_not_final_2) & nchar(estimated_dwrf_emerging_contaminant_principal_forgiveness_not_final_2) > 2 ~ "Yes",
      !is.na(estimated_dwrf_disadvantgd_community_principal_forgiveness_not_final_3) & nchar(estimated_dwrf_disadvantgd_community_principal_forgiveness_not_final_3) > 2 ~ "Yes",
      !is.na(estimated_wif_grant_not_final_4) & nchar(estimated_wif_grant_not_final_4) > 2 ~ "Yes",
      TRUE ~ "No"
    ),
    state = "Minnesota"
  ) %>%
  # subset columns
  select(state_rank, state_score, borrower, pwsid, project_description, population, disadvantaged,
         funding_amount, principal_forgiveness_amount, fundable, project_type, state)


```


## Mississippi (DD)

- Applicant data processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# (48,9)
ms_raw <- fread("data/year1/csv/24-Mississippi_PPL.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# -> (42,8)
ms_clean <- ms_raw %>%
  # drop category rows and funding line row
  filter(!grepl("Category", project) & project_description != "NA") %>%
  # format numeric columns
  mutate(
         population = as.numeric(str_replace_all(service_area_population,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(loan_amount_requested,"[^0-9.]","")),
         principal_forgiveness_amount = as.numeric(str_replace_all(eligible_pf_amount,"[^0-9.]","")),
         # use this to separate funding/applicant projects, but don't keep in standardized data
         state_cumulative = as.numeric(str_replace_all(statewide_cum, "[^0-9.]",""))) %>%
  # format text columns
  mutate(borrower = str_squish(project),
         project_description = str_squish(project_description),
         state_score = str_replace_all(priority_points,"[^0-9.]",""),
         state = "Mississippi",
         project_type = "Other",
         fundable = case_when(
           # funding line set at 42,500,000 in PPL
           state_cumulative < 42500000 ~ "Fundable",
           TRUE ~ "Applicant")) %>%
  select(borrower, project_description, state_score, funding_amount, principal_forgiveness_amount,
         state, project_type, fundable)
  


```


## Missouri

- Applicant data is processed
- No mention `project_cost`, mentions `requested_amount`
- No mention of *Interest Rates or Loan Terms*

```{r}
# this is the initially scraped IUP from 10/11/22,
# however, it is structurally the same as the list in the amended IUP dated 7/11/23, so we keep it and amend individual values below
# (43,12) -> (43,13)
mo_ppl <- fread("data/year1/csv/25-Missouri_PPL_Cleaned.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  mutate(
    # Projects included on the General Funding List are considered Funded. Projects included on the General Contingency List, General Planning List are considered Not Funded.
    fundable = case_when(
      # list projects in planning list
      project_number %in% c("DW291386-01", "DW291170-02", "DW291387-01", "DW291393-01", "DW291388-01", 
                            "DW291389-01", "DW291390-01", "DW291391-01", "DW291392-01", "DW291314-04") ~ "Applicant",
      TRUE ~ "Fundable")
    )

# read in lead tables from the amended IUP dated 7/11/23
mo_lead_fundable <- fread("data/year1/csv/25-Missouri_LSLFundable.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  mutate(
    fundable = "Fundable",
  )

mo_lead_contingency <- fread("data/year1/csv/25-Missouri_LSLContingency.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  mutate(
    fundable = "Applicant",
  )

# (272,14)
mo_comb <- bind_rows(mo_ppl, mo_lead_fundable, mo_lead_contigency)


### Amendments
# The following changes bring the document in line with the most recent figures in the 7/11/23 IUPP
# We also need to decide what to do with projects that are not on this amended list that were previously included.
# mo_amended <- mo_ppl %>%
#   mutate(
#     iup_amount_requested = case_when(
#       project_number == "DW291142-03" ~ "$951,500",
#       project_number == "DW291384-01" ~ "$628,400",
#       TRUE ~ iup_amount_requested),
#     loan_amount = case_when(
#       project_number == "DW291370-01" ~ "$1,640,505",
#       project_number == "DW291142-02" ~ "$593,700",
#       project_number == "DW291363-02" ~ "$892,238",
#       project_number == "DW291142-03" ~ "$951,500",
#       project_number == "DW291384-01" ~ "$628,400",
#       project_number == "DW291301-02" ~ "$1,164,940",
#       TRUE ~ loan_amount),
#     additional_subsidization_amount = case_when(
#       project_number == "DW291370-01" ~ "$3,000,000",
#       project_number == "DW291142-02" ~ "$1,781,100",
#       project_number == "DW291363-02" ~ "$2,676,712",
#       project_number == "DW291378-01" ~ "$1,954,995",
#       project_number == "DW291301-02" ~ "$5,361,000",
#       TRUE ~ additional_subsidization_amount)
#   )

# -> (272,10)
mo_clean <- mo_comb %>%
  # drop columns
  select(-carryover) %>%
  # process numeric columns
  mutate(
         population = as.numeric(str_replace_all(service_area_population,"[^0-9.]","")),
         requested_amount = as.numeric(str_replace_all(iup_amount_requested,"[^0-9.]","")),
         loan_amount = as.numeric(str_replace_all(loan_amount,"[^0-9.]","")),
         additional_subsidization_amount = as.numeric(
           str_replace_all(additional_subsidization_amount,"[^0-9.]","")),
         # replace NA with 0 for column adding
         additional_subsidization_amount = replace_na(additional_subsidization_amount, 0),
         loan_amount = replace_na(loan_amount, 0),
         funding_amount = loan_amount + additional_subsidization_amount
           ) %>%
  # process text columns
  mutate(borrower = str_squish(applicant),
         project_name = str_squish(project_number),
         state_score = str_replace_all(priority_points,"[^0-9.]",""),
         project_description = str_squish(description_needs_category),
         disadvantaged = case_when(
           disadvantaged == "D" ~ "Yes",
           TRUE ~ "No"),
         project_type = case_when(
           grepl("EC",additional_subsidization_funding_source) ~ "Emerging Contaminants",
           grepl("LSLI", description_needs_category) ~ "Lead",
           TRUE ~ "Other"
         ),
         state = "Missouri"
  ) %>%
  rename(principal_forgiveness_amount = additional_subsidization_amount) %>%
  select(state_score, borrower, project_description, population, funding_amount,
         principal_forgiveness_amount, disadvantaged, fundable, state, project_type)
  
  
```


## Montana (A)

- TODO: Applicant data needs to be scraped from Appendix 2 of this doc - https://deq.mt.gov/files/Water/TFAB/DWSRF/IUP-PPL/2023_DWSRF_%20IUP_FINAL.pdf -- Data can't be scraped because it is a non-machine readable screenshot. Will need to manually create spreadsheet.
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

## FUNDABLE
# (37,4)
# Table 1. DWSRF Projects Anticipated to Receive Funding SFY 2023 from 
# https://deq.mt.gov/files/Water/TFAB/DWSRF/IUP-PPL/2023_DWSRF_%20IUP_FINAL.pdf
mt_raw <- fread("data/year1/csv/26-Montana_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() 

# -> (36,9)
mt_clean <- mt_raw %>%
  # drop rows
  filter(!is.na(priority_rank)) %>%
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(priority_rank,"[^0-9.]","")),
         # split project_information and transform to numbers
         population = as.character(map(strsplit(project_information, split = "\\."), 1)),
         population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         funding_amount = as.character(map(strsplit(srf_cost, split = " "), 1)),
         funding_amount = as.numeric(str_replace_all(funding_amount,"[^0-9.]","")),
         # extract whether funding includes prinicipal forgiveness or not
         funding_P = str_squish(as.character(map(strsplit(srf_cost, split = " "), 2))),
         # assume that principal forgiveness is 75% of loan, up to 750k per page 5 of IUP
         principal_forgiveness_amount = case_when(
           # if PF included, store PF as 75% of loan amount
           funding_P == "P" ~ round(funding_amount * 0.75,0),
           funding_P != "P" ~ 0,
           TRUE ~ 0),
         # if PF is greater than 750k, set to max of 750k
         principal_forgiveness_amount = case_when(
           principal_forgiveness_amount > 750000 ~ 750000,
           TRUE ~ principal_forgiveness_amount)
         ) %>%
  # process text columns
  mutate(project_description = as.character(map(strsplit(project_information, split = "\\."), 2)),
         project_description = str_squish(project_description),
         borrower = str_squish(project),
         disadvantaged = case_when(
           funding_P == "P" ~ "Yes",
           TRUE ~ "No"),
         # all projects are fundable based on title name
         fundable = "Fundable",
         # No Lead or EC projects found
         project_type = "Other",
         state = "Montana") %>%
  select(state_rank, borrower, project_description, funding_amount, 
         principal_forgiveness_amount, population, project_type, fundable, state)

## APPLICANT
# Appendix 2: DWSRF COMPREHENSIVE PROJECT LISTSFY 2023 from
# https://deq.mt.gov/files/Water/TFAB/DWSRF/IUP-PPL/2023_DWSRF_%20IUP_FINAL.pdf

  
```


## Nebraska (A)

- TODO: Applicant data needs to be processed - scrape appendix B2 https://drive.google.com/file/d/19Pb8cJRKm-fO2s_Zt2iKilELdTDhmpIX/view
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# read in base, all non-lead or ec projects, (14,9)
ne_base <- fread("data/year1/csv/27-Nebraska_ppl_base.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type="Other")

# read in bil, all non-lead or ec projects, (13,9)
ne_bil <- fread("data/year1/csv/27-Nebraska_ppl_bil.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type="Other")

# read in ec projects, (7,9)
ne_ec <- fread("data/year1/csv/27-Nebraska_ec.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type="Emerging Contaminants")

# read in lead projects (19,9)
ne_lsl <- fread("data/year1/csv/27-Nebraska_lslr.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type="Lead")

# merge data, (53,9)
ne_merged <- bind_rows(ne_base, ne_bil, ne_ec, ne_lsl)

## APPLICANT
# Appendix B2 

# -> (53,9)
ne_clean <- ne_merged %>%
  # drop columns
  select(-forgiveness_percent) %>%
  # format numeric columns
  mutate(
         population = as.numeric(str_replace_all(population, "[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(project_est_cost, "[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(forgiveness_amount, "[^0-9.]", "")),
  ) %>%
  # format text columns
  mutate(fundable= "Fundable",
         state= "Nebraska",
         state_score = str_squish(priority_points),
         disadvantaged = case_when(
           principal_forgiveness_amount > 0 ~ "Yes",
           TRUE ~ "No")
         ) %>%
  rename(borrower = community,
         pwsid = pws_number) %>%
  select(borrower, pwsid, state_score, project_description, funding_amount, principal_forgiveness_amount, project_type,
         disadvantaged, fundable, state)

```


## New Hampshire

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (190,32)
nh_raw <- fread("data/year1/csv/29-NewHampshire_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (181,13)
nh_clean <- nh_raw %>%
  # drop columns
  select(-a_water_quality_score, -b_quantity_deficiencies_score, 
         -c_treatment_score, -d_storage_score, -e_distribution_score, 
         -f_affordability_score, -g_cap_dev_score, -h_green_score, 
         -i_resiliency_score,
         -j_consolidation_interconnecti_on_score, -k_critical_infrastructur_e_score, 
         -l_project_readiness_score,
         -water_rate, -mhi, -affordability_index,
         -requested_loan_amount, -arpa_disadvantaged_co_op_or_epa_wiin_grant,
         -x2022_arpa_grant) %>%
  
  # drop empty rows
  filter(!is.na(rank)) %>%
  # drop "n/a" rows which are only ARPA Funding
  filter(rank != "n/a") %>%
  
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(rank, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(population, "[^0-9.]", "")),
         
         # transform all funding columns for summing
         # only NAs are the "Bypass..." applicant projects,
         # these will be filled with 0s, but are technically NA, 
         # but this will only matter when we expand the project to work with all Applicant projects 
         # and can be replaced with NA downstream for easier processing
         emerging_contaminants = replace_na(emerging_contaminants, "0"),
         emerging_contaminants = as.numeric(str_replace_all(emerging_contaminants, "[^0-9.]", "")),
         lead_service_line_lsl_loan_amount = replace_na(lead_service_line_lsl_loan_amount, "0"),
         lead_service_line_lsl_loan_amount = as.numeric(
           str_replace_all(lead_service_line_lsl_loan_amount, "[^0-9.]", "")),
         base_supplemental_loan_amount = replace_na(base_supplemental_loan_amount, "0"),
         base_supplemental_loan_amount = as.numeric(
           str_replace_all(base_supplemental_loan_amount, "[^0-9.]", "")),
         lsl_forgiveness = replace_na(lsl_forgiveness, "0"),
         lsl_forgiveness = as.numeric(str_replace_all(lsl_forgiveness, "[^0-9.]", "")),
         base_supplemental_forgiveness = replace_na(base_supplemental_forgiveness, "0"),
         base_supplemental_forgiveness = as.numeric(
           str_replace_all(base_supplemental_forgiveness, "[^0-9.]", "")),
         ## define funding amount as sum of loan, assumes it includes forgiveness_amount
         funding_amount = emerging_contaminants + lead_service_line_lsl_loan_amount 
         + base_supplemental_loan_amount,
         ## define PF as sum of forgiveness columns
         principal_forgiveness_amount = lsl_forgiveness + base_supplemental_forgiveness
           ) %>%
  
  # process text columns
  mutate(pwsid = case_when(
          nchar(pws_number) == 7 ~paste0("NH", pws_number),
          nchar(pws_number) == 6 ~ paste0("NH0", pws_number),
          nchar(pws_number) == 5 ~ paste0("NH00", pws_number),
          pws_number == "1392200/139222 0/1392230" ~ "NH1392200/NH1392220/NH1392230"),
         # deal with multiple pwsid entries
         # pwsid = case_when(
         # grepl("/", pwsid) ~ as.character(map(strsplit(pwsid, split = "/"), 1)),
         # TRUE ~ pwsid),
         state_score = str_replace_all(total_score, "[^0-9.]", ""),
         borrower = str_to_title(str_squish(applicant)),
         cities_served = str_to_title(str_squish(system_town)),
         project_name = str_to_title(str_squish(project_name)),
         disadvantaged = case_when(
           financially_disadvantaged == "YES" | environmentall_y_disadvantaged == "YES" ~ "Yes",
           TRUE ~ "No"),
         # fundable indicated by highlighting in document, manually add projects by rank
         fundable = case_when(
           state_rank %in% c(1,2,3,4,6,7,8,9,10,11,12,15,16,17,
                             19,20,21,22,23,24,25,26,27,28,29,30) ~ "Fundable",
           TRUE ~ "Applicant"),
         project_type = case_when(
           emerging_contaminants > 0 ~ "Emerging Contaminants",
           lead_service_line_lsl_loan_amount > 0 ~ "Lead",
           TRUE ~ "Other"),
         state = "New Hampshire"
         ) %>%
  select(borrower, pwsid, cities_served, project_name, project_type, state_rank, state_score,
         funding_amount, principal_forgiveness_amount, disadvantaged, population, fundable, state)
  
  
  
```


## New Jersey (DD)

- Applicant data processed
- Assumes `project_cost` should be included
- No mention of *Interest Rates or Loan Terms*

```{r}

# (676, 19)
nj_raw <- fread("data/year1/csv/30-NewJersey_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (676,10)
nj_clean <- nj_raw %>%
  select(-cat_a, -cat_b, -cat_c_a, -cat_c_b, -cat_c_c, -cat_c_d, -cat_d, -cat_e) %>%
  # process numeric columns
  mutate(
    population = as.numeric(str_replace_all(population, "[^0-9.]", "")),
    project_cost = as.numeric(str_replace_all(estimated_cost, "[^0-9.]", "")),
  ) %>%
  # process text columns
  mutate(
    borrower = str_squish(project_sponsor),
    project_name = str_squish(project_number),
    pwsid = str_squish(pwsid),
    project_description = str_squish(project_name),
    state_rank = str_squish(rank),
    state_score = str_squish(rank_points),
    project_type = case_when(
      grepl("EC", bil_eligibility) ~ "Emerging Contaminants",
      grepl("LSL", bil_eligibility) ~ "Lead",
      TRUE ~ "Other"),
    state = "New Jersey",
    fundable = "No Information"
  ) %>%
  select(borrower, pwsid, state_rank, state_score, project_name, project_description, project_cost, 
         population, state, fundable)

```


## New Mexico (A)

- TODO: Applicant data needs to be processed - need to find it, not in a separate list in the document in our data dictionary
- mentions `amount_requested`, but used for funding amount
- No mention of *Interest Rates or Loan Terms*

```{r}
# (6,13)
nm_raw <- fread("data/year1/csv/31-NewMexico_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (6,12)
nm_clean <- nm_raw %>%
  # drop columns
  select(-mandatory_cap_grant_subsidy, -interest_rate, -designation) %>%
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(ranking, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(population, "[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(amount_requested, "[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(subsidy_amount_eligible_to_project, "[^0-9.]", "")),
  ) %>%
  # process text columns
  # use the start of the pwsid to separate name and pwsid, then reappend the state abbreviation
  mutate(borrower = as.character(map(strsplit(water_system_name_and_number, split = "NM"), 1)),
         borrower = str_squish(borrower),
         state_score = str_replace_all(score, "[^0-9.]", ""),
         pwsid = paste0("NM", as.character(map(strsplit(water_system_name_and_number, split =  "NM"), 2))),
         cities_served = str_to_title(county),
         project_description = str_to_sentence(str_squish(project_description)),
         disadvantaged = case_when(
           grepl("Disadvantaged", disadvantaged_status, ignore.case=TRUE) ~ "Yes",
           TRUE ~ "No"),
         fundable = "Fundable",
         project_type = "Other",
         state = "New Mexico"
         ) %>%
  select(state_rank, state_score, borrower, pwsid, project_description, funding_amount, principal_forgiveness_amount, disadvantaged, population, cities_served, project_type, fundable, state)

```


## Nevada

- Applicant data is processed
- No mention of `project_cost`, includes `requested_amount``
- No mention of *Interest Rates or Loan Terms*

```{r}
# (172, 16)
# Nevada Drinking Water State Revolving Fund Priority List Effective June 2022
# https://ndep.nv.gov/uploads/water-financing-srf-drinkingwater-docs/DW_modified_FINAL_PL-Effective-June-2022.pdf
nv_ppl <- fread("data/year1/csv/28-Nevada_PPL.csv",
                colClass="character", na.strings="") %>%
  clean_names()

# -> (169, 11) - drop is due to the removal of "Total" rows and non fundable projects
nv_ppl <- nv_ppl %>%
  # remove non project rows
  filter(!grepl("Total", priority_number)) %>%
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(priority_number, "[^0-9.]", "")),
         population = as.numeric(str_replace_all(project_population, "[^0-9.]", "")),
         requested_amount = as.numeric(str_replace_all(estimated_loan_amount, "[^0-9.]", ""))) %>%
  # process text columns
  mutate(disadvantaged = case_when(
    is.na(d2) ~ "No",
    TRUE ~ "Yes"
  ),
         borrower = str_squish(entity),
         pwsid = str_squish(state_id),
         state_score = str_replace_all(revised_points, "[^0-9.]", ""),
         project_description = str_squish(project_description)
  ) %>%
  select(borrower, pwsid, project_description, state_rank, state_score, requested_amount, 
         population, disadvantaged)


# (28,15)
### Nevada Attachment C: DWSRF 2022 Fundable List DWSRF 2nd Amended Intended Use Plan SFY 2022
nv_fund <- fread("data/year1/csv/28-Nevada_Fundinglist.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  # process numeric columns
  mutate(
         state_rank = as.numeric(str_replace_all(pl_rank, "[^0-9.]", "")),
         # replace the listing of 93,94, and 95 with only 95 to attach funding_amount to a single project
         # then manually set all state ranks to the PPL list to force correct matches
         state_rank = case_when(
           state_rank == 939495 ~ 95,
           state_rank == 57 ~ 60,
           state_rank == 61 ~ 62,
           state_rank == 97 ~ 100,
           state_rank == 122 ~ 126,
           state_rank == 129 & pws_id == "NV0000226" ~ 132,
           state_rank == 163 ~ 165,
           state_rank == 169 ~ 52,
           TRUE ~ state_rank),
         funding_amount = as.numeric(str_replace_all(total_assistance, "[^0-9.]", "")),
         # transform the components of PF into numeric and replace NAs for summing
         # TODO: Confirm whether disadvantaged_subsidy should also be included here
         base_disadvantaged_subsidy = as.numeric(str_replace_all(base_disadvantaged_subsidy, "[^0-9.]", "")),
         base_disadvantaged_subsidy = replace_na(base_disadvantaged_subsidy, 0),
         base_additional_subsidy = as.numeric(str_replace_all(base_additional_subsidy, "[^0-9.]", "")),
         base_additional_subsidy = replace_na(base_additional_subsidy, 0),
         disadvantaged_subsidy = as.numeric(str_replace_all(disadvantaged_subsidy, "[^0-9.]", "")),
         disadvantaged_subsidy = replace_na(disadvantaged_subsidy, 0),
         principal_forgiveness_amount = base_disadvantaged_subsidy + base_additional_subsidy + disadvantaged_subsidy
         ) %>%
  # process text columns
  mutate(
         fundable = "Fundable",
         ) %>%
  # keep standardized columns to match onto the PPL List
  select(state_rank, funding_amount, principal_forgiveness_amount, fundable)


### Join PPL and funding list
nv_clean <- merge(nv_ppl, nv_fund, all=TRUE, by="state_rank") %>%
  # assign variables consistent across both
  mutate(state = "Nevada",
         project_type = "Other",
         # manually update the two projects that are funded in project ranked 95, fill rest in as applicant
         fundable = case_when(
           state_rank == 93 ~ "Fundable",
           state_rank == 94 ~ "Fundable",
           is.na(fundable) ~ "Applicant",
           TRUE ~ fundable
         ),
         #manually fix one missing project description
         project_description = case_when(
           borrower == "Beatty Water & Sanitation District - DW" ~ "Water System Rehab",
           TRUE ~ project_description
         ))



```


## New York (DD)

- Walker's notes say that when Score == H, it denotes disadvantaged in the base PPL. Should projects with a numeric Score be marked as not disadvantaged or "No Information"? Below assumes No Info since that is the more conservative interpretation.

```{r}

## Base
# project annual list, p33 of FFY 2023 Final IUP
# (558,10) -> (558,12)
ny_base <- fread("data/year1/csv/32-NewYork_base_ppl.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  mutate(project_type = "Other",
         disadvantaged = case_when(
           score == "H" ~ "Yes",
           TRUE ~ "No"))

## BIL
# (120,10) -> (120,11)
ny_bil <- fread("data/year1/csv/32-NewYork_bil_ppl.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  select(-v1) %>%
  mutate(project_type = "Other",
         disadvantaged = "Yes")

## Lead - Amendment 4
# (109,11) -> (109,12)
ny_lead <- fread("data/year1/csv/32-NewYork_lead.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  select(-v1) %>%
  mutate(project_type = "Lead",
         disadvantaged = case_when(
           dac == "DAC" ~ "Yes",
           TRUE ~ "No"
         ))

## Emerging Contaminants - Amendment 2
# (37,10) -> (37,11)
ny_ec <- fread("data/year1/csv/32-NewYork_ec.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  select(-pfas) %>%
  mutate(project_type = "Emerging Contaminants",
         disadvantaged = case_when(
           dac == "DAC" ~ "Yes",
           TRUE ~ "No")
         )

## Combine & Clean
# -> (824, 13)
ny_comb <- bind_rows(ny_base, ny_bil, ny_lead, ny_ec)

# -> (824,11)
ny_clean <- ny_comb %>%
  select(-dac, -cumulative_total, -code) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(pop, "[^0-9.]", "")),
         project_cost = as.numeric(str_replace_all(project_cost, "[^0-9.]", "")),
         ) %>%
  # process text columns
  mutate(project_name = str_squish(project_number),
         cities_served = str_squish(county),
         borrower = paste(str_squish(system_name), "/", str_squish(borrower)),
         state_score = str_squish(score),
         project_description = str_squish(description),
         state = "New York",
         fundable = "Applicant") %>%
  select(borrower, state_score, project_name, project_description, project_cost,
         population, cities_served, disadvantaged, project_type, state, fundable)


```


## North Carolina (Q in SS)

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# (221, 15)
nc_raw <- fread("data/year1/csv/33-NorthCarolina_PPL_Comprehensive.csv",
                colClass="character", na.strings="") %>%
  clean_names()

# -> (221,11)
nc_clean <- nc_raw %>%
  # format numeric columns
  mutate(population = as.numeric(str_replace_all(service_populati_on, "[^0-9.]", "")),
         # pre-format columns that will add up to funding amount and PF
         base_dwsrf_loans = as.numeric(str_replace_all(base_dwsrf_loans, "[^0-9.]", "")),
         bil_suppl_dwsrf_loans = as.numeric(str_replace_all(bil_suppl_dwsrf_loans, "[^0-9.]", "")),
         bil_suppl_dwsrf_principal_forgivene_ss = as.numeric(str_replace_all(bil_suppl_dwsrf_principal_forgivene_ss, "[^0-9.]", "")),
         base_dwsrf_principal_forgivene_ss = as.numeric(str_replace_all(base_dwsrf_principal_forgivene_ss, "[^0-9.]", "")),
         # combine for funding amount and PF
         # does dwsrf loans already include PF? - does not appear so, as there are rows with only PF and DWSRF loans are 0, 
         # so they will need to be added together as well
         principal_forgiveness_amount = bil_suppl_dwsrf_principal_forgivene_ss + base_dwsrf_principal_forgivene_ss,
         funding_amount = base_dwsrf_loans + bil_suppl_dwsrf_loans + principal_forgiveness_amount
         )%>%
  # format text columns
  mutate(borrower = str_squish(applicant_name),
         project_name = str_squish(project_name),
         state_score = str_replace_all(priorit_y_points, "[^0-9.]", ""),
         pwsid = case_when(
           pwsid == 'x' ~ as.character(NA),
           TRUE ~ pwsid),
         cities_served = str_squish(county),
         project_type = "Other",
         state = "North Carolina") %>%
  mutate(fundable = case_when(
    !is.na(funding_amount) ~ "Fundable",
    TRUE ~ "Applicant")) %>%
  select(borrower, pwsid, project_name, state_score, funding_amount, principal_forgiveness_amount,
         cities_served, project_type, population, state, fundable)

```


## North Dakota

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

## Read in tables for PF funding from the tables in page 10-13 of PPL

# (4,6) -> (4,7)
nd_gen <- fread("data/year1/csv/34-NorthDakota_GeneralSupplementalFunding.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  mutate(project_type = "Other")

# (2,6) -> (2,7)
nd_ec <- fread("data/year1/csv/34-NorthDakota_EmergingContaminantsFunding.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  mutate(project_type = "Emerging Contaminants")

# (45,6) -> (45,7)
nd_lead <- fread("data/year1/csv/34-NorthDakota_LeadFunding.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  mutate(project_type = "Lead")


## COMBINE GEN, EC, AND LEAD FOR PF INFO
# -> (51, 8)
nd_funding <- bind_rows(nd_gen, nd_ec, nd_lead) %>%
  mutate(principal_forgiveness = as.numeric(str_replace_all(additional_subsidy,"[^0-9.]","")),
         project_name = str_squish(tracking_no)) %>%
  # combine PF funding across the tables on p 10-13 where projects appear twice
  group_by(project_name) %>%
  summarize(principal_forgiveness_amount = sum(principal_forgiveness))
  


## APPENDIX B 
# (262,14) -> (262,11)
nd_ppl <- fread("data/year1/csv/34-NorthDakota_PPL.csv",
                colClass="character", na.strings="") %>%
  clean_names() %>%
  # drop columns not needed for analysis
  select(-construction_start_date, -est_loan_term, -v14)


#funded projects by rank determined by highlighting in table
funded <- c(1,4,5,6,8,9,10,12,13,22,25,35,38,39,40,50,60,75,76,78,88,93,108,109,110,111,122,124,125,126,127,
            148,151,152,156,166,169,178,199,200,201,202,215,216,217,218,219,220,221,245,250)

# (262,12)
nd_clean <- nd_ppl %>%
  # transform numeric columns
  mutate(state_rank = as.numeric(priority_ranking_supplemental),
         population = as.numeric(str_replace_all(present_population,"[^0-9.]","")),
         # transform to numeric and multiply by 1000
         funding_amount_gen_1000 =  as.numeric(str_replace_all(project_cost_1_000,"[^0-9.]","")) * 1000,
         funding_amount_ec_1000 =  as.numeric(str_replace_all(project_cost_emerging_contaminants_1_000,"[^0-9.]","")) * 1000,
         funding_amount_lead_1000 =  as.numeric(str_replace_all(project_cost_lead_1_000,"[^0-9.]","")) * 1000,
         # replace NAs with 0 for adding together
         funding_amount_gen_1000 =   replace_na(funding_amount_gen_1000, 0),
         funding_amount_ec_1000 =   replace_na(funding_amount_ec_1000, 0),
         funding_amount_lead_1000 = replace_na(funding_amount_lead_1000, 0),
         # add together to make total funding amount
         funding_amount = funding_amount_gen_1000 + funding_amount_ec_1000 + funding_amount_lead_1000
  ) %>%
  # process text columns
  mutate(project_description = str_squish(project_description),
         project_name = str_squish(tracking_no),
         fundable = case_when(
           state_rank %in% funded ~ "Fundable",
           TRUE ~ "Applicant"),
         # extract numeric portion of pwsid and append state abbreviation 
         pwsid = paste0("ND", as.character(map(strsplit(tracking_no, split = "-"), 1))),
         project_type = case_when(
           project_cost_emerging_contaminants_1_000 != "-" ~ "Emerging Contaminants",
           project_cost_lead_1_000 != "-" ~ "Lead",
           TRUE ~ "Other"),
         state = "North Dakota"
         ) %>%
  # rename columns
  rename(disadvantaged = disadvantaged_community,
         borrower = system_name) %>%
  left_join(nd_funding) %>%
  # replace NAs introduced by adding principal_forgiveness_amount
  mutate(principal_forgiveness_amount = replace_na(principal_forgiveness_amount, 0)) %>%
  # keep columns for analysis
  select(borrower, project_name, pwsid, project_type, project_description,
         state_rank, funding_amount, principal_forgiveness_amount, disadvantaged, population, fundable, state)



```


## Ohio (Q in DD)

- From the IUP, it looks like 35-Ohio_PPL_Base includes all projects and the other tables simply add data
- what should we call `estimated_loan_amount` considering we are treating this like Applicant data because clearly they are not all going to get funded? is it `project_cost`? Shouldn't be `funding_amount` if we are keeping that for only fundable projects
- what should be done with `estimated_principal_forgiveness`? Similar questions as above, but also note that three of the six projects are listed as BYPASS, meaning that would need to be dropped for it to be a numeric column and we'd lose half of the values from that table

```{r}

## Base
# -> (440,9)
oh_base <- fread("data/year1/csv/35-Ohio_PPL_Base.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(-v1) %>%
  # reset all column names and drop column name row
  rename(entity = v2, project = v3, pwsid = v4, population = v5, county = v6, 
         estimated_loan_amount = v7, loan_type = v8, estimated_award_date = v9, rate = v10) %>%
  filter(entity != "Entity")
  


## HAB Regional PF
#
oh_pf <- fread("data/year1/csv/35-Ohio_PPL_RegionalPF.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  filter(!is.na(estimated_principal_forgiveness)) %>%
  select(entity, estimated_loan_amount, estimated_principal_forgiveness)


# merge PF into the base table
oh_base <- oh_base %>%
  left_join(oh_pf) %>%
  # manually fix one instance where est. PF is assigned to a single project where two projects are listed in base,
  # assume that it goes to the larger portion of funds for construction
  mutate(estimated_principal_forgiveness = case_when(
    entity == "Walnut Creek Water Company" & estimated_loan_amount == "$3,950,000" ~ "$2,070,000",
    TRUE ~ estimated_principal_forgiveness
  ))


## Lead
#
oh_lead <- fread("data/year1/csv/35-Ohio_LSLR.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(entity, project, estimated_loan_amount) %>%
  mutate(project_type = "Lead")
  

## Emerging Contaminants
# 
oh_ec <- fread("data/year1/csv/35-Ohio_PPL_HAB_PFAS.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  select(entity, project, estimated_loan_amount) %>%
  mutate(project_type = "Emerging Contaminants")


# combined lead and ec for single left join to bring in project_type
oh_lead_ec <- bind_rows(oh_lead, oh_ec)

oh_comb <- oh_base %>%
  left_join(oh_lead_ec, by=c("entity", "estimated_loan_amount"))

# -> (440,9)
oh_clean_2 <- oh_comb %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         project_cost = as.numeric(str_replace_all(estimated_loan_amount,"[^0-9.]","")),
         ) %>%
  # process text columns
  mutate(borrower = str_squish(entity),
         project_description = str_squish(project.x),
         pwsid = str_squish(pwsid),
         cities_served = str_squish(county),
         state = "Ohio",
         fundable = "Applicant",
         project_type = case_when(
           # fill in a few gaps from the joins
           is.na(project_type) & grepl("HAB", rate) ~ "Emerging Contaminants",
           grepl("Lead", project_description) | grepl("LSL", project_description) | grepl("LSL", rate) ~ "Lead",
           is.na(project_type) ~ "Other",
           TRUE ~ project_type),
         disadvantaged = case_when(
           grepl("DIS", rate) ~ "Yes",
           TRUE ~ "No")
         ) %>%
  select(borrower, pwsid, project_description, project_cost,
         cities_served, population, disadvantaged, state, project_type, fundable) %>%
  # remove duplicates from potential one:many relationship from the left_join for project_type
  distinct()

```


## Oklahoma

- Applicant data processed
- Includes `requested_amount`, does not include `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}


## base / bil
# (69,12) -> (69,14)
ok_b <- fread("data/year1/csv/36-Oklahoma_PPL-AppB.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type = "Other") %>%
  rename(disadvantaged = severly_disadvantged_disadvantged_or_no)

ok_b$fundable <- ""
ok_b$fundable[1:46] <- "Fundable"
ok_b$fundable[47:69] <- "Applicant"

# for fundable projects, loan amount is funding amount
ok_b <- ok_b %>%
  mutate(funding_amount = case_when(
    fundable == "Fundable" ~ as.numeric(str_replace_all(loan_amount,"[^0-9.]","")),
    TRUE ~ 0
  ))

## lead
# (42,11) -> (42,13)
ok_e <- fread("data/year1/csv/36-Oklahoma_PPL-AppE.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type = "Lead") %>%
  rename(disadvantaged = disadvantaged_y_or_n)

# only first 24 projects are fundable 
ok_e$fundable <- ""
ok_e$fundable[1:24] <- "Fundable"
ok_e$fundable[25:42] <- "Applicant"

# for fundable projects, loan amount is funding amount
ok_e <- ok_e %>%
  mutate(funding_amount = case_when(
    fundable == "Fundable" ~ as.numeric(str_replace_all(loan_amount,"[^0-9.]","")),
    TRUE ~ 0
  ))


## emerging contaminants
# (3,11) -> (3,13)
ok_f <- fread("data/year1/csv/36-Oklahoma_PPL-AppF.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  mutate(project_type = "Emerging Contaminants",
         fundable = "Fundable",
         # all projects on EC list are funded
         funding_amount = as.numeric(str_replace_all(loan_amount,"[^0-9.]",""))) %>%
  rename(disadvantaged = disadvantaged_y_or_n)


# (114, 12)
ok_comb <- bind_rows(ok_b, ok_e, ok_f)

# -> (114,10)
ok_clean <- ok_comb %>%
  # drop columns
  select(-base, -cumulative_amount, -anticipated_binding_commitment_date, -anticipated_construction_date) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         # for all projects on all lists, loan amount is requested amount, even if it is also funding_amount for funded projects
         requested_amount = as.numeric(str_replace_all(loan_amount,"[^0-9.]","")),) %>%
  # process text columns
  mutate(state_score = case_when(
    priority_points == "Being Ranked" ~ "No Information",
    TRUE ~ str_squish(priority_points)),
    borrower = str_squish(system),
    project_description = str_squish(project_description),
    project_name = str_squish(project_number),
    # disadvantaged can either be S/D/Y for some degree of disadvantaged or N for No
    disadvantaged = case_when(
      disadvantaged == "N" ~ "No",
      TRUE ~ "Yes"),
    state = "Oklahoma",
         ) %>%
  select(borrower, state_score, project_name, project_description, requested_amount, funding_amount,
         disadvantaged, population, project_type, state, fundable)


```


## Oregon (NS)

- Data exists in Excel spreadsheet as well as in PDF tables that appear to have not been scraped, though it is unclear precisely which document is which or how to interpret
- Imported purely as a CSV, data will be difficult to parse when combined in the same columns, will likely require manual edits for a file that can be processed

```{r}

```


## Pennsylvania

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (62,23)
pa_raw <- fread("data/year1/csv/38-Pennsylvania_manual_fundable_PPL.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (62,14)
pa_clean <- pa_raw %>%
  mutate(state_rank = as.numeric(gsub(",", "", projrank)),
         state_score = gsub(",", "",pv_rating),
         population = as.numeric(gsub(",", "", population)),
         funding_amount = as.numeric(str_replace_all(assistance_amount,"[^0-9.]", "")),
         # ASSUMPTION: NA different from 0, not adding 0 unless needed for adding columns
         # funding_amount = replace_na(funding_amount, 0),
         principal_forgiveness_amount = as.numeric(str_replace_all(principal_forgiveness,"[^0-9.]", "")),
         ) %>%
  ## split applicant name to borrower and project name
  ## get rid of wifta- then split at - if there is one
  mutate(applicant = str_to_title(applicant),
         applicant = gsub("wifta-|wifta -|wifta|wifta partial -", "", applicant, ignore.case = T),
         ## remove parenthesis and anything in between
         applicant = str_replace_all(applicant, "\\s*\\([^\\)]+\\)", ""),
         ## borrower is text before dash
         borrower = str_squish(str_extract(applicant, "[^-]+")),
         ## project name is text after dash
         project_name = str_squish(str_extract(applicant, "[^-]+$")),
         ## append PA to beginning of pwsids
         pwsid = paste0("PA", pwsid),
         ## recategorize project type column
         project_type = case_when(grepl("lsl", project_type, ignore.case = T) ~ "Lead",
                                  grepl("lead service", project_type, ignore.case = T) ~ "Lead",
                                  grepl("lead service", proj_description, ignore.case = T) ~ "Lead",
                                  grepl("pfas", proj_description, ignore.case = T) ~ "Emerging Contaminants",
                                  grepl("pfas", applicant, ignore.case = T) ~ "Emerging Contaminants",
                                  grepl("lsl", project_type, ignore.case = T) &
                                    grepl("pfas", proj_description, ignore.case = T) ~ "Lead,
                                                                                        Emerging Contaminants",
                                  TRUE ~ "Other"),
         project_description = str_squish(proj_description),
         ## fundable column may need to be done manually because no column to join on
         fundable = case_when(fundable == "yes" ~ "Fundable",
                              TRUE ~ "Applicant"),
         cities_served = str_to_title(city),
         # only three EC projects have a disadvantaged distinction
         disadvantaged = case_when(
           borrower == "Susquehanna Area Reg. Airport Authority" ~ "No",
           project_name == "Pa American Water Company Frackville" ~ "Yes",
           project_name == "Saegertown Borough" ~ "No",
           TRUE ~ as.character(NA)
         ),
         state = "Pennsylvania"
         ) %>%
  ## standardize names
  ## keep relevant columns
  select(borrower, pwsid, cities_served, project_name, project_description, project_type,
         state_rank, state_score, funding_amount, principal_forgiveness_amount, disadvantaged, population, fundable, state)
```


## Rhode Island (WIP)

- Applicant data is processed, but not ready to add
- Mentions `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

the comprehensive list provides data on each borrower such as population and PWSID
as well as project descriptions and scores for each project, even where there are multiple projects for the same borrower

the funding list, on the other hand, is only at the borrower level and only includes
the loan amount, the borrower's name (which often does not match the comprehensive list),
and has been manually updated to include project_type based on which table the projects were listed
the one:many relationship in the comprehensive list and the different units of analysis creates a
degree of uncertainty in which projects are funded or not.

Easy example: City of Newport is listed as receiving one loan for Lead and 
have only one Lead project which matches between the requested amount and the loan amount

More Difficult example: Providence Water is receiving two loans. 25M for Lead and 21M for General.
While they only have one Lead project, it is for $184M.
They have 3 General projects with different Project Descriptions that are all requesting more than $25M,
so it is not clear which project is being funded and which data should be included in the dashboard.

```{r}

ri_comp <- fread("data/year1/csv/39-RhodeIsland_PPL_Comprehensive.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  # drop columns
  select(-a, -b, -c, -d, -e, -f, -g, -source_fund, -new, -est_start_date) %>%
  # process numeric columns
  mutate(
         funding_amount = as.numeric(str_replace_all(funds_requested,"[^0-9.]", "")),
         population = as.numeric(str_replace_all(pop_served,"[^0-9.]", "")),

         ) %>%
  # process text columns
  mutate(borrower = str_squish(system_name),
         borrower = str_remove(borrower, "\\*"),
         state_score = str_replace_all(scores_total,"[^0-9.]", ""),
         pwsid = paste0("RI", pws_id),
         project_description = str_squish(project_description),
         state = "Rhode Island"
         )

ri_fund <- fread("data/year1/csv/39-RhodeIsland_PPL_Fundable_Manual.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  mutate(funding_amount = as.numeric(str_replace_all(loan_amount,"[^0-9.]", "")),
         fundable = "Fundable",
         borrower = str_remove(borrower, "\\*")) %>%
  select(-loan_amount) %>%
  left_join(ri_comp, by="borrower")

```


## South Carolina (WIP) (A)

- TODO: Applicant data needs to be processed - may not be available
- TODO: Add in supplemental, EC, and lead tables (to be scraped) - see data dictionary
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# (20,11)
sc_raw <- fread("data/year1/csv/37-SouthCarolina_PPL.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# -> (18,11)
sc_clean <- sc_raw %>%
  # remove non-project rows
  filter(sponsor_and_project_name != "") %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(estimated_srf_loan_amount,"[^0-9.]","")),
         funding_amount = replace_na(funding_amount, 0),
         principal_forgiveness_amount = as.numeric(str_replace_all(estimated_principal_forgiveness_assistance2,"[^0-9.]","")),
         principal_forgiveness_amount = replace_na(principal_forgiveness_amount, 0),
         # funding amount should include PF
         funding_amount = funding_amount + principal_forgiveness_amount,
         population = as.numeric(str_replace_all(population_affected_by_project,"[^0-9.]","")),
         ) %>%
  # process text columns
  mutate(project_description = str_squish(project_description),
         # take the first portion of the water system id number and attach SC
         pwsid = paste0("SC", as.character(map(strsplit(srf_project_number, split = "-"), 1))),
         borrower = str_squish(as.character(map(strsplit(sponsor_and_project_name, split = " - "), 1))),
         project_name = str_squish(as.character(map(strsplit(sponsor_and_project_name, split = " - "), 2))),
         state_score = str_replace_all(total_points,"[^0-9.]",""),
         state = "South Carolina",
         project_type = case_when(
           grepl("lead", project_description, ignore.case=TRUE) ~ "Lead",
           grepl("lsl", project_description, ignore.case=TRUE) ~ "Lead",
           TRUE ~ "Other"),
         fundable = "Fundable"
         ) %>%
  select(borrower, pwsid, state_score, project_name, project_description, funding_amount, principal_forgiveness_amount,
         population, project_type, state, fundable)


```


## South Dakota

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- Mentions *Interest Rates or Loan Terms*

```{r}

# (23,8)
sd_fnd <- fread("data/year1/csv/41-SouthDakota_Fundable.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# (136,8)
sd_ppl <- fread("data/year1/csv/41-SouthDakota_PPL2023.csv",
                      colClasses = "character", na.strings = "") %>% clean_names()

# (136,14)
# merge fundable and applicant projects together as they are listed twice,
# using priority points and project number as unique ID
sd_merged <- merge(sd_ppl, sd_fnd, by=c("priority_points", "project_number"), all=TRUE)

# -> (136,10)
sd_clean <- sd_merged %>%
  # drop columns
  select(-expected_loan_rate_term, -loan_recipient, -funding_date, -expected_funding_source) %>%
  # process numeric columns
  mutate(funding_amount = as.numeric(str_replace_all(assistance_amount,"[^0-9.]","")),
         population = as.numeric(str_replace_all(pop_served,"[^0-9.]",""))
         ) %>%
  # process numeric columns with NAs
  rowwise() %>%
  mutate(principal_forgiveness_amount = sum(as.numeric(str_replace_all(principal_forgiveness,"[^0-9.]", "")),
                                            na.rm=TRUE),
        ) %>%
  ungroup() %>%
  # process text columns
  mutate(project_description = str_squish(project_description),
         borrower = str_squish(community_public_water_system),
         state_score = str_replace_all(priority_points,"[^0-9.]",""),
         disadvantaged = case_when(
           grepl("Yes", disadvantaged) ~ "Yes",
           TRUE ~ "No"),
         project_type = case_when(
           grepl("4", fund_project_eligibility) ~ "Lead",
           grepl("5", fund_project_eligibility) ~ "Emerging Contaminants",
           TRUE ~ "Other"
         ),
         fundable = case_when(
           funding_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"
         ),
         state = "South Dakota"
        ) %>% 
  # keep relevant columns
  select(borrower, project_description, project_type, state_score, funding_amount, 
         principal_forgiveness_amount, population, disadvantaged, fundable, state) 
```


## Tennessee (NS)
- Waiting on final data

```{r}

```


## Texas

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# (267,14) -> (265,9)
# 
tx_ppl <- fread("data/year1/csv/43-Texas_PPL.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  # drop & rename columns for easier merge
  select(-owner_type, -green_type, -related_pif_number_s, -gpr, -requested_phase_s) %>%
  rename(project_cost = total_project_cost) %>%
  # fix inconsistent formatting for merge
  mutate(project_description = str_squish(project_description),
         project_description = str_replace(project_description, "12- mile", "12-mile")) %>%
  # get rid of total columns
  filter(!grepl("Total", rank, ignore.case=TRUE))
  

# (38,13) -> (38,10)
# Appendix K: Initial Invited Project List
tx_k <- fread("data/year1/csv/43-Texas_Fundable_K.csv",
               colClasses= "character", na.strings="") %>% 
  clean_names() %>%
  select(-eligible_phase_s, -green_type, -related_pif_number_s, -gpr) %>%
  mutate(project_description = str_squish(project_description),
         # fix specific typos to prevent merge issues
         pws_id = str_replace(pws_id, "1330135", "TX1330135"),
         fundable = "Fundable")

# (268,10)
tx_merge <- merge(tx_ppl, tx_k, by=c("pif_number", "rank", "points", "entity", "pws_id", "population",
                                       "project_description", "project_cost", "disadv_percent"), all=TRUE)

# -> (268,11)
tx_clean <- tx_merge %>%
  # remove extra columns
  select(-pif_number) %>%
  # process numeric columns
  mutate(state_rank = as.numeric(str_replace_all(rank,"[^0-9.]","")),
         population = as.numeric(str_replace_all(population,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(project_cost,"[^0-9.]",""))
  ) %>%
  # process text columns
  mutate(borrower = str_squish(entity),
         pwsid = str_squish(pws_id),
         project_description = str_squish(project_description),
         state_score = str_replace_all(points,"[^0-9.]",""),
         disadvantaged = case_when(
          is.na(disadv_percent) ~ "No",
          TRUE ~ "Yes"),
         project_type = case_when(
           grepl("lead", project_description, ignore.case=TRUE) ~ "Lead",
           grepl("emerging contaminant", project_description, ignore.case=TRUE) ~ "Emerging Contaminants",
           TRUE ~ "Other"),
         fundable = case_when(
           fundable == "Fundable" ~ "Fundable",
           TRUE ~ "Applicant"),
         state = "Texas"
  ) %>%
  select(state_rank, state_score, borrower, pwsid, project_description, funding_amount, population, disadvantaged, fundable, project_type, state)
```


## Utah (Q in DD)

- TODO: Applicant data is processed, but needs to be interpreted and assigned the variable
- `project_total` is mentioned, similar to `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

ut_ppl <- fread("data/year1/csv/44-Utah_Comprehensive.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

ut_clean <- ut_ppl %>%
  # drop columns
  select(-green_project, -green_amount, -equivalency_project,
         -project_segments_sour, -project_segments_treat, -project_segments_stor, -project_segments_dist) %>%
  # format numeric columns
  mutate(
    population = as.numeric(str_replace_all(pop,"[^0-9.]","")),
    # TODO: determine if ARPA needs to be subtracted from it
    funding_amount = as.numeric(str_replace_all(srf_assistance,"[^0-9.]","")),
    principal_forgiveness = as.numeric(str_replace_all(principal_forgiveness,"[^0-9.]","")),
    project_cost = as.numeric(str_replace_all(project_total,"[^0-9.]","")),
  ) %>%
  # format non-numeric columns
  mutate(
    pwsid = str_squish(pwsid),
    state_score = str_squish(priority_points),
    borrower = str_squish(system_name),
    cities_served = str_squish(county),
    project_description = str_squish(project_title),
    disadvantaged = case_when(disadvantaged == "Y" ~ "Yes",
                              TRUE ~ "No"),
    project_type = case_when(
      grepl("lead", project_description, ignore.case=TRUE) ~ "Lead",
      TRUE ~ "Other"),
    state = "Utah",
    # TODO: Add fundable / applicant depending on interpretation
  ) %>%
  select(state_score, borrower, pwsid, project_description, funding_amount, principal_forgiveness,
         project_type, cities_served, disadvantaged, population, state)
  

```


## Virginia (NS)

```{r}

```


## Vermont

- Applicant data is processed
- No mention of `requested_amount` or `project_cost`
- No mention of *Interest Rates or Loan Terms*

```{r}

# PPL - Fundable and Applicants
# Document: DWSRF FFY2022 Priority Funding List - General, IUP for FFY22, PPLs are on pp. 45-48. 
# (99,12)
vt_ppl <- fread("data/year1/csv/45-Vermont_PPL.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# drop the last three items which are ineligible projects
vt_ppl <- vt_ppl[1:96,]

# set fundable variable based on where the line is in the document
vt_ppl$fundable <- ""
vt_ppl$fundable[1:46] <- "Fundable"
vt_ppl$fundable[47:96] <- "Applicant"

# (96,13) -> (93,11)
vt_ppl <- vt_ppl %>%
  # drop rows if WSID is empty (sub-heading columns)
  filter(!is.na(wsid)) %>%
  select(-years, -admin_percent) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(user_popln,"[^0-9.]","")),
         loan_acct_gb = as.numeric(str_replace_all(loan_acct_gb,"[^0-9.]","")),
         gb_disadv_subsidy = as.numeric(str_replace_all(gb_disadv_subsidy,"[^0-9.]","")),
         # replace nas for adding columns
         gb_disadv_subsidy = replace_na(gb_disadv_subsidy, 0),
         loan_acct_gs = as.numeric(str_replace_all(loan_acct_gs,"[^0-9.]","")),
         gs_disadv_subsidy = as.numeric(str_replace_all(gs_disadv_subsidy,"[^0-9.]","")),
         gs_disadv_subsidy = replace_na(gs_disadv_subsidy, 0),
         funding_amount = as.numeric(str_replace_all(x2022_loan_amount,"[^0-9.]","")),
         principal_forgiveness_amount = gb_disadv_subsidy + gs_disadv_subsidy
  ) %>%
  # process text columns
  mutate(project_description = str_squish(project),
         borrower = str_squish(water_system_borrower),
         state_score = str_replace_all(score,"[^0-9.]",""),
         # fill in leading zeros to standardize length
         pwsid = case_when(
           nchar(wsid) == 4 ~ paste0("VT000", wsid),
           nchar(wsid) == 5 ~ paste0("VT00", wsid)),
         # all ppl projects are non ec and non lead
         project_type = "Other",
         # for fundable projects, if PF > 0, Disadvantaged 
         # because it is just the sum of GB_disadv_subsidy and gs_disadv_subsidy
         # Applicants assume unknown since PF will be 0 by default of not receiving subsidy
         state="Vermont"
  ) %>%
  # subset columns
  select(state_score, borrower, pwsid, project_description, population,
         funding_amount, principal_forgiveness_amount, project_type, fundable, project_type, state)


## Import Lead tables independently to reconcile column head differences
## Document: DWSRF FFY22 Priority List - Lead service line removal and service line inventories, IUP for FFY22, pp. 49-50. 
# (3,7)
vt_lead_t1 <- fread("data/year1/csv/45-Vermont_Lead_Table1.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()

# (70,7)
vt_lead_t2 <- fread("data/year1/csv/45-Vermont_Lead_Table2.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names() %>%
  rename(line_loan_amount = loan4,
         loan_forgiveness = loan_forgiveness2,
         water_system_borrower = water_system)

# merge together lead tables to process columns
# (73, 7)
vt_lead <- bind_rows(vt_lead_t1, vt_lead_t2) 

# -> (73,10)
vt_lead <- vt_lead %>%
  select(-lsi_score3, -loan_portion_to_be_repaid) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(user_popln,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(line_loan_amount,"[^0-9.]","")),
         principal_forgiveness_amount = as.numeric(str_replace_all(loan_forgiveness,"[^0-9.]","")),
  ) %>%
  # process text columns
  mutate(borrower = str_squish(water_system_borrower),
         state_score = str_replace_all(application_score,"[^0-9.]",""),
         pwsid = case_when(
           nchar(wsid) == 4 ~ paste0("VT000", wsid),
           nchar(wsid) == 5 ~ paste0("VT00", wsid)),
         project_type = "Lead",
         fundable = "Fundable",
         disadvantaged = case_when(
           principal_forgiveness_amount > 0 ~ "Yes",
           TRUE ~ "No"
         ),
         state = "Vermont"
  ) %>%
  select(state_score, borrower, pwsid, funding_amount, principal_forgiveness_amount, population,
         disadvantaged, project_type, fundable, state)


# Emerging Contaminants
# (17,7) -> (8,7)
vt_ec <- fread("data/year1/csv/45-Vermont_EC.csv",
                      colClasses = "character", na.strings = "") %>% 
  clean_names()
# keep only the top section of the table that are not internal account transfers
vt_ec <- vt_ec[2:9,]

# -> (8,10)
vt_ec <- vt_ec %>%
  select(-subtotals) %>%
  mutate(population = as.numeric(str_replace_all(user_popln,"[^0-9.]","")),
         funding_amount = as.numeric(str_replace_all(line_items,"[^0-9.]","")),
         principal_forgiveness_amount = as.numeric(str_replace_all(line_items,"[^0-9.]","")),
  ) %>%
  mutate(borrower = str_squish(water_system),
         project_description = str_squish(project),
         pwsid = case_when(
           nchar(wsid) == 4 ~ paste0("VT000", wsid),
           nchar(wsid) == 5 ~ paste0("VT00", wsid)),
         state_score = str_replace_all(plist_app_score,"[^0-9.]",""),
         project_type = "Emerging Contaminants",
         fundable = "Fundable",
         disadvantaged = case_when(
            principal_forgiveness_amount > 0 ~ "Yes",
            TRUE ~ "No"
         ),
         state = "Vermont") %>%
  select(state_score, borrower, pwsid, funding_amount, principal_forgiveness_amount,
         project_description, disadvantaged, population, project_type, fundable, state)


# combine (174,11)
vt_clean <- bind_rows(vt_ppl, vt_lead, vt_ec)


```


## Washington (A)

- TODO: Applicant data needs to be processed - may not be published/available unless we count the "rejected/declined" projects in Appendix C, but this doesn't seem to fit the other definitions of Applicant projects in other states
- No mention of `project_cost`
- `loan_request_amount` present, but used for `funding_amount`
- No mention of *Interest Rates or Loan Terms*

```{r}
# import manually merged file (31,12)
wa_merged <- fread("data/year1/csv/46-Washington_PPL_Merged.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (29,12)
wa_clean <- wa_merged %>%
  # drop total rows and extra columns
  filter(!grepl("total", health_application, ignore.case=TRUE)) %>%
  select(-health_application, -comments) %>%
  # process numeric columns
  mutate(population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         funding_amount = as.numeric(str_replace_all(loan_request_amount, "[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(subsidy_award, "[^0-9.]", "")),
  ) %>%
  # process text columns
  # append state abbreviation and leading numbers to water system id
  mutate(pwsid = paste0("WA53", water_system_id),
         borrower = str_squish(applicant_name),
         state_score = str_replace_all(final_scor_e,"[^0-9.]", ""),
         project_name = str_squish(project),
         cities_served = str_squish(county),
         project_description = str_squish(project_description),
         fundable = "Fundable",
         project_type = case_when(
           grepl("lead", project_description, ignore.case=TRUE) ~ "Lead",
           grepl("PFAS", project_description, ignore.case=TRUE) ~ "Emerging Contaminants",
           TRUE ~ "Other"),
         disadvantaged = case_when(
           principal_forgiveness_amount > 0 ~ "Yes",
           principal_forgiveness_amount == 0 ~ "No"),
         state = "Washington"
  ) %>%
  select(state_score, borrower, pwsid, project_name, project_description, cities_served, population, 
         funding_amount, principal_forgiveness_amount, fundable, disadvantaged, project_type, state)
  
```


## West Virginia

- Applicant data processed
- Mentions `requested_amount` as `dwtrf_funding_requested` and `project_cost` as `total_cost`
- Mentions *Interest Rates or Loan Terms*

```{r}
# (163,36)
# this manually merged updated file includes 
wv_raw <- fread("data/year1/csv/48-West_Virginia_Merged_Updated.csv",
                      colClasses = "character", na.strings = "") %>%
  clean_names()

# -> (157,15)
wv_clean <- wv_raw %>%
  ## get rid of rows that the rank is not a number (not a project)
  filter(!is.na(ranking) & county != "NA" & ranking !="Total") %>% 
  mutate(state_rank = as.numeric(ranking)) %>%
  ## rowwise operations for sums
  rowwise() %>%
  ## make relevant columns numbers 
  mutate(state_rank = as.numeric(str_replace_all(ranking,"[^0-9.]", "")),
         population = as.numeric(str_replace_all(new_popu_lation,"[^0-9.]", "")),
         principal_forgiveness_amount =
           # Sum of DWTRF Principal Forgiveness from Base Grant, 
           # DWTRF Principal Forgiveness from Supplemental Grant, 
           # DWTRF Principal Forgiveness from Emerging Cont. Grant, and 
           # DWTRF Principal Forgiveness from LSLR Grant in Funding List
           sum(as.numeric(str_replace_all(dwtrf_principal_forgiveness_from_base_grant,"[^0-9.]","")),
               as.numeric(str_replace_all(dwtrf_principal_forgiveness_from_supplemental_grant,"[^0-9.]","")),
               as.numeric(str_replace_all(dwtrf_principal_forgiveness_from_emerging_cont_grant,"[^0-9.]","")),
               as.numeric(str_replace_all(dwtrf_principal_forgiveness_from_lslr_grant,"[^0-9.]","")),
               na.rm = T),
         funding_amount = 
           # Sum of all PF sources and
           # DWTRF Assistance from Base Grant, 
           # DWTRF Assistance from Supplemental Grant, 
           # DWTRF Assistance from LSLR Grant
           sum(principal_forgiveness_amount,
               as.numeric((str_replace_all(dwtrf_assisitance_from_base_grant,"[^0-9.]",""))),
               as.numeric((str_replace_all(dwtrf_assisitance_from_supplemental_grant,"[^0-9.]",""))),
               as.numeric((str_replace_all(dwtrf_assisitance_from_lslr_grant,"[^0-9.]",""))),
                                         na.rm=T)
         ) %>%
  ungroup() %>%
  # process numeric columns that don't require aggregating
  mutate(
    project_cost = as.numeric(str_replace_all(total_cost,"[^0-9.]","")),
    requested_amount = as.numeric(str_replace_all(dwtrf_funding_requested,"[^0-9.]",""))
  ) %>%
  ## split applicant name to borrower and project name
  ## get rid of wifta- then split at - if there is one
  mutate(project_type = case_when(grepl("lead", dwtrf_eligible_funding_type, ignore.case = T) ~
                                    "Lead",
                                  grepl("pfas", dwtrf_eligible_funding_type, ignore.case = T) ~
                                    "Emerging Contaminants",
                                  TRUE ~ "Other"),
         fundable = case_when(
           funding_amount > 0 | principal_forgiveness_amount > 0 ~ "Fundable",
           TRUE ~ "Applicant"),
         state_score = str_replace_all(points,"[^0-9.]", ""),
         project_description = str_squish(project_description_x),
         project_name = str_squish(project_name),
         borrower = str_squish(system),
         disadvantaged = str_squish(disadvan_taged),
         cities_served = str_squish(county),
         state = "West Virginia"
         ) %>%
  ## keep relevant columns
  select(borrower, project_name, project_description, project_type,
         state_rank, state_score, funding_amount, principal_forgiveness_amount,
         project_cost, requested_amount,
         disadvantaged, population, cities_served, fundable, state)
```


## Wisconsin

- Applicant data is processed
- Mentions `requested_amount`
- No mention of *Interest Rates or Loan Terms*

```{r}
# (54,17) -> (46,19)
wi_fund_raw <- fread("data/year1/csv/49-Wisconsin_PPL.csv",
                     colClasses = "character", na.strings = "") %>%
    clean_names() %>%
  ## remove non project rows
  filter(!is.na(project_points)) %>%
  # define disadvantaged for the non-EC projects as total points > 59 per data dictionary
  mutate(disadvantaged = case_when(
    as.numeric(str_replace_all(total_pf_points,"[^0-9.]", "")) > 59 ~ "Yes",
                         TRUE ~ "No"),
         # determine project type by keyword search for non-EC projects
         project_type = case_when(grepl("lsl", project_description, ignore.case = T) ~ "Lead",
                                  grepl("pfas", project_description, ignore.case=T) ~ "Emerging Contaminants",
                                  TRUE ~ "Other"),)

# (4,17) -> (4,19)
wi_ec_fund <- fread("data/year1/csv/49-Wisconsin_EC_Funding_List.csv",
                    colClasses = "character", na.strings = "") %>%
  clean_names() %>%
  # define disadvantaged as financial need points > 29 for only EC projects
  mutate(disadvantaged = case_when(as.numeric(str_replace_all(financial_need_points,"[^0-9.]", "")) > 29 ~ "Yes",
                         TRUE ~ "No"),
         project_type = "Emerging Contaminants")

# (50,19)
wi_fund_raw <- bind_rows(wi_fund_raw, wi_ec_fund)

# -> (50,11)
wi_clean <- wi_fund_raw %>%
  ## make relevant columns numbers
  mutate(## sum of columns, getting rid of $ and comma and replacing blanks with zero
         funding_amount = as.numeric(str_replace_all(requested_project_costs,"[^0-9.]", "")),
         principal_forgiveness_amount = as.numeric(str_replace_all(pf_estimate,"[^0-9.]", "")),
         ## make principal forgiveness NAs zeros
         principal_forgiveness_amount = replace_na(principal_forgiveness_amount, 0),
         population = as.numeric(str_replace_all(population,"[^0-9.]", "")),
         requested_amount = as.numeric(str_replace_all(requested_project_costs,"[^0-9.]", "")),
         ) %>%
  ## non-numeric columns
  mutate(fundable = case_when(!is.na(funding_amount) ~ "Fundable",
                              TRUE ~ "Applicant"),
         ## get rid of numbers from borrower name
         borrower = str_to_title(str_replace_all(municipality, "[0-9.\\#]", "")),
         state_score = str_replace_all(priority_score,"[^0-9.]",""),
         cities_served = borrower,
         state = "Wisconsin"
         ) %>%
  ## keep relevant columns
  select(borrower, cities_served, project_description, project_type,
         state_score, funding_amount, principal_forgiveness_amount,
         disadvantaged, population, fundable, state)
```


## Wyoming (NS)
- Waiting on final data

```{r}

```


# Post-Process

## Combine States

```{r}
combined_clean <- bind_rows(
           al_clean, 
           ar_clean,
           dc_clean,
           de_clean, 
           ga_clean, 
           hi_clean, 
           id_clean,
           il_clean, 
           la_clean, 
           ma_clean, 
           md_clean,
           me_clean,
           mi_clean, 
           mn_clean, 
           #mo_clean,
           #mt_clean,
           nd_clean, 
           ne_clean, 
           #nh_clean,
           nm_clean,
           nv_clean,
           pa_clean, 
           sd_clean, 
           tx_clean, 
           vt_clean,
           #wa_clean, 
           wi_clean, 
           wv_clean)

```

## Process Combined States

This format is then pre-processed by the Dashboard cleaning script for it's final format as it appears in the application.

```{r}

## Cleaning for table on web, currently leaves in Applicant projects for 1.1 release.
web_combined_clean <- combined_clean %>%
                          mutate(across(everything(), as.character),
                                 # TEMPORARY: fix all fundable nomenclature at once at the end.
                                 # TODO for 1.1.1 - change throughout code for consistency
                                 fundable = case_when(
                                   fundable == "Fundable" ~ "Funded",
                                   fundable == "Applicant" ~ "Not Funded",
                                   TRUE ~ as.character(NA)
                                 )) %>%
                          select(state, cities_served, borrower,
                                 pwsid, project_name, project_description,
                                 project_type, funding_amount, principal_forgiveness_amount, project_cost, requested_amount,
                                 population, disadvantaged, state_rank, state_score, fundable) %>%
                          rename(State = state,
                                Borrower = borrower, 
                                `City Served` = cities_served, 
                                `Project Description` = project_description,
                                `Project Type` = project_type,
                                `State Rank` = state_rank,
                                `State Score` = state_score,
                                `Project Cost` = project_cost,
                                `Requested Amount` = requested_amount,
                                `Funding Amount` = funding_amount, 
                                `Principal Forgiveness` = principal_forgiveness_amount,
                                `Meets State Disadvantaged Criteria` = disadvantaged,
                                 Population = population,
                                 PWSID = pwsid, 
                                `Project Name` = project_name,
                                `Funding Status` = fundable)

```

## Writing Data

## Write to AWS

```{r}

data <- web_combined_clean
filename <- "web_ppl_combined_clean.csv"
aws_filepath <- "clean_data/srf_project_priority_lists/web_ppl_combined_clean_v1-1.csv"

### To put in an AWS bucket, you need these credentials
## Contact gabe@policyinnovation.org for questions/access
## !!! NOTE !!! NEVER WRITE CREDENTIALS IN CODE, ENTER IN CONSOLE AS ENVIRONMENT VARIABLES !!!! NOTE !!!! #### 

# Sys.setenv("AWS_ACCESS_KEY_ID" = "YOUR ACCESS KEY",
#            "AWS_SECRET_ACCESS_KEY" = "YOUR SECRET KEY",
#            "AWS_DEFAULT_REGION" = "us-east-1")

# Writing to temp 
fwrite(data, file.path(tempdir(), filename))

#Putting in Bucket
put_object(
  file = file.path(tempdir(), filename),
  object = aws_filepath,
  bucket = "water-team-data",
  acl = "public-read"
)

```