This notebook transforms the project-level data that powers the SRF Funding Tracker into state-level aggregate data and store it in AWS.

The generated features include status columns for each column in the project-level data - whether it is Complete, Partial, or Missing, as well as summarized statistics for the count of projects, the type of project, whether the applicants are state-defined DACs, and whether they are expecting funding or not. Percentages and other statistics are then generated from these results. See the full data dictionary [here](https://docs.google.com/document/d/1tdFCCTpw3GKs9-kmNnRkm7tjK-yktozTeO3fA9ejuPg/edit?usp=sharing).

The resulting dataset can answer a number of high-level questions for each state in a given state fiscal year, like "what percentage of DACs are expecting funding?" or the inverse - "how many projects expecting funding are supporting DACs?". The dataset is used extensively in `year-over-year.Rmd` to create the data visualizations used in the SRF Funding Tracker state landing pages.

```{r}

source("resources.R")
run_code_from_file("aws.txt")

```

```{r}
# import project-level data
all_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names()
```

## Create Column Status Features

```{r}


create_status_cols <- function(df, col_names) {
  
  # create the returned dataframe of only state and SFY
  output_df <- df %>%
    select(state, state_fiscal_year) %>%
    distinct()
  
  # for each column passed through, make a new column name and test coverage
  for (col_name in col_names) {
  
  new_col <- paste0(col_name, "_status")
  col_sym <- sym(col_name)

  new_df <- df %>%
    group_by(state, state_fiscal_year) %>%
    mutate(
      # save new column name as one of these features based on whether it's NA or includes 'No Information'
      !!new_col := case_when(
        all(is.na(!!col_sym)) ~ 'Missing',
        any(!!col_sym == "No Information") ~ 'Partial',
        TRUE ~ 'Complete'
      )
    ) %>% 
  select(state, state_fiscal_year, !!new_col) %>%
  distinct()
  
  output_df <- output_df %>%
    left_join(new_df, by=c("state", "state_fiscal_year"))
  
 }
  
  return(output_df)

}


feature_coverage <- create_status_cols(all_projects, col_names=c("community_served", "borrower", "pwsid", "project_id", "project_name", "project_type",
                                                     "project_cost", "requested_amount", "funding_amount", "principal_forgiveness", "population",
                                                     "project_description", "disadvantaged", "project_rank", "project_score", "expecting_funding" ))
  
```

## Create Numeric Columns

```{r}

# NOTE: replacing NAs for math when calculating percents, 
# but when this merges in, we need to go back and add "Missing" when it was actually NA in the project data
# for project_type calculations, NA represents when we didn't have this data
project_features <- all_projects %>%
  group_by(state, state_fiscal_year) %>%
  summarize(projects = n(),
            
            general_projects = sum(project_type == "General"),
            ec_projects = sum(project_type == "Emerging Contaminants"),
            lead_projects = sum(project_type == "Lead"),
            
            # number of projects expecting funding and percent of all projects expecting funding
            expecting_funding_projects = sum(expecting_funding == "Yes"),
            expecting_funding_projects = replace_na(expecting_funding_projects, 0),
            expecting_funding_percent = round(expecting_funding_projects / projects, 2)*100,
            
            ef_general_projects = sum(expecting_funding == "Yes" & project_type == "General"),
            ef_ec_projects = sum(expecting_funding == "Yes" & project_type == "Emerging Contaminants"),
            ef_lead_projects = sum(expecting_funding == "Yes" & project_type == "Lead"),

            # number of disadvantaged projects and percent of all projects that are DACs
            disadvantaged_projects = sum(disadvantaged == "Yes"),
            disadvantaged_projects = replace_na(disadvantaged_projects, 0),
            disadvantaged_percent = round(disadvantaged_projects / projects, 2)*100,
            
            dac_general_projects = sum(disadvantaged == "Yes" & project_type == "General"),
            dac_ec_projects = sum(disadvantaged == "Yes" & project_type == "Emerging Contaminants"),
            dac_lead_projects = sum(disadvantaged == "Yes" & project_type == "Lead"),
            
            # projects by type & dacs & expecting funding
            dac_general_ef_projects = sum(disadvantaged == "Yes" & project_type == "General" & expecting_funding == "Yes"),
            dac_lead_ef_projects = sum(disadvantaged == "Yes" & project_type == "Lead" & expecting_funding == "Yes"),
            dac_ec_ef_projects = sum(disadvantaged == "Yes" & project_type == "Emerging Contaminants" & expecting_funding == "Yes"),
            
            # number of projects that are both DACs and expecting funding
            disadvantaged_expecting_funding_projects = sum(expecting_funding == "Yes" & disadvantaged == "Yes"),
            disadvantaged_expecting_funding_projects = replace_na(disadvantaged_expecting_funding_projects, 0),
            
            # inverse: projects that are DACs and NOT expecting funding
            disadvantaged_not_expecting_funding_projects = sum(expecting_funding == "No" & disadvantaged == "Yes"),
            
            # percent of all projects that are DACs expecting funding
            disadvantaged_expecting_funding_percent_all = round(disadvantaged_expecting_funding_projects / projects, 2)*100,
            
            # percent of projects expecting funding that are DACs
            disadvantaged_expecting_funding_percent_ef = round(disadvantaged_expecting_funding_projects / expecting_funding_projects, 2)*100,
            
            # percent of DACs that are expecting funding
            disadvantaged_expecting_funding_percent_dac = round(disadvantaged_expecting_funding_projects / disadvantaged_projects, 2)*100,
            
            # number of projects serving small communities
            sc_projects = sum(convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000),
            sc_ef_projects = sum(convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 & expecting_funding == "Yes"),
            sc_dac_projects = sum(convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 & disadvantaged == "Yes"),
            sc_ef_dac_projects = sum(convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 & expecting_funding == "Yes" & disadvantaged == "Yes"),
            # inverse of above for plotting purposes
            non_sc_projects = projects - sc_projects,
            dac_non_sc = disadvantaged_projects - sc_dac_projects,
            ef_non_sc = expecting_funding_projects - sc_ef_projects
            )

```

```{r}
project_cost_aggregation <- all_projects %>%
  filter(!is.na(project_cost) & project_cost != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost)))

pc_by_type <- all_projects %>%
  filter(!is.na(project_cost) & project_cost != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost))) %>%
  pivot_wider(names_from=project_type, values_from=total_project_cost) %>%
  rename(ec_total_project_cost = `Emerging Contaminants`,
         lead_total_project_cost = `Lead`,
         general_total_project_cost = `General`) %>%
  select(state, state_fiscal_year, ec_total_project_cost, lead_total_project_cost, general_total_project_cost)

project_cost_aggregation <- project_cost_aggregation %>%
  left_join(pc_by_type, by=c("state", "state_fiscal_year"))
```

```{r}
funding_amount_aggregation <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount)))

fa_by_type <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount))) %>%
  pivot_wider(names_from=project_type, values_from=total_funding_amount) %>%
  rename(ec_total_funding_amount = `Emerging Contaminants`,
         lead_total_funding_amount = `Lead`,
         general_total_funding_amount = `General`) %>%
  select(state, state_fiscal_year, ec_total_funding_amount, lead_total_funding_amount, general_total_funding_amount)

funding_amount_aggregation <- funding_amount_aggregation %>%
  left_join(fa_by_type, by=c("state", "state_fiscal_year"))
```

```{r}
requested_amount_aggregation <- all_projects %>%
  filter(!is.na(requested_amount) & requested_amount != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_requested_amount = sum(as.numeric(requested_amount)))

ra_by_type <- all_projects %>%
  filter(!is.na(requested_amount) & requested_amount != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_requested_amount = sum(as.numeric(requested_amount))) %>%
  pivot_wider(names_from=project_type, values_from=total_requested_amount) %>%
  rename(ec_total_requested_amount = `Emerging Contaminants`,
         lead_total_requested_amount = `Lead`,
         general_total_requested_amount = `General`) %>%
  select(state, state_fiscal_year, ec_total_requested_amount, lead_total_requested_amount, general_total_requested_amount)

requested_amount_aggregation <- requested_amount_aggregation %>%
  left_join(ra_by_type, by=c("state", "state_fiscal_year"))
```

```{r}
median_wo_zero <- function(col_name) {
  non_zero <- col_name[col_name != 0]  # Filter out 0s
  median(non_zero)  # Calculate median on the filtered data
}

principal_forgiveness_aggregation <- all_projects %>%
  filter(!is.na(principal_forgiveness) & principal_forgiveness != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_principal_forgiveness = sum(as.numeric(principal_forgiveness)),
            projects_receiving_pf = sum(as.numeric(principal_forgiveness) > 0),
            dacs_receiving_pf = sum(disadvantaged == "Yes" & as.numeric(principal_forgiveness) > 0),
            dacs_general_pf = sum(disadvantaged == "Yes" & project_type == "General" & as.numeric(principal_forgiveness) > 0),
            dacs_lead_pf = sum(disadvantaged == "Yes" & project_type == "Lead" & as.numeric(principal_forgiveness) > 0),
            dacs_ec_pf = sum(disadvantaged == "Yes" & project_type == "Emerging Contaminants" & as.numeric(principal_forgiveness) > 0),
            median_pf = median_wo_zero(as.numeric(principal_forgiveness))
            )
```

```{r}

state_stats <- feature_coverage %>%
  left_join(project_features, by=c("state", "state_fiscal_year")) %>%
  left_join(funding_amount_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(principal_forgiveness_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(project_cost_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(requested_amount_aggregation, by=c("state", "state_fiscal_year"))
  

```

```{r}

 # Writing to temp
 write.csv(state_stats, file.path(tempdir(), "dwsrf-funding-tracker-state-stats.csv"), row.names = FALSE)

 #Putting in Bucket
 put_object(
   file = file.path(tempdir(), "dwsrf-funding-tracker-state-stats.csv"),
   object = "clean_data/srf_project_priority_lists/dwsrf-funding-tracker-state-stats.csv",
   bucket = "water-team-data",
   acl = "public-read"
 )
```
