This notebook transforms the project-level data that powers the SRF Funding Tracker into state-level aggregate data and store it in AWS.

The generated features include status columns for each column in the project-level data - whether it is Complete, Partial, or Missing, as well as summarized statistics for the count of projects, the type of project, whether the applicants are state-defined DACs, and whether they are expecting funding or not. Percentages and other statistics are then generated from these results. See the full data dictionary [here](https://docs.google.com/document/d/1tdFCCTpw3GKs9-kmNnRkm7tjK-yktozTeO3fA9ejuPg/edit?usp=sharing).

The resulting dataset can answer a number of high-level questions for each state in a given state fiscal year, like "what percentage of DACs are expecting funding?" or the inverse - "how many projects expecting funding are supporting DACs?". The dataset is used extensively in `year-over-year.Rmd` to create the data visualizations used in the SRF Funding Tracker state landing pages.

## Source scripts
```{r, warning=FALSE}
source("resources/inputs.R")
source("resources/cleaning.R")
source("resources/testing.R")
run_code_from_file("aws.txt")

```

## Read Clean Data (All Projects)
```{r}
# import project-level data
all_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() |>
  dplyr::mutate(state_fiscal_year = as.character(state_fiscal_year))
```

## Create Column Status Features
```{r}
# only used here for data creation once, no need to reference from resources
create_status_cols <- function(df, col_names) {
  
  # create the returned dataframe of only state and SFY
  output_df <- df %>%
    select(state, state_fiscal_year) %>%
    distinct()
  
  # for each column passed through, make a new column name and test coverage
  for (col_name in col_names) {
  
  new_col <- paste0(col_name, "_status")
  col_sym <- sym(col_name)

  new_df <- df %>%
    group_by(state, state_fiscal_year) %>%
    mutate(
      # save new column name as one of these features based on whether it's NA or includes 'No Information'
      !!new_col := case_when(
        all(is.na(!!col_sym)) ~ 'Missing',
        any(!!col_sym == "No Information") ~ 'Partial',
        TRUE ~ 'Complete'
      )
    ) %>% 
  select(state, state_fiscal_year, !!new_col) %>%
  distinct()
  
  output_df <- output_df %>%
    left_join(new_df, by=c("state", "state_fiscal_year"))
 }
  return(output_df)
}


feature_coverage <- create_status_cols(all_projects,
                                       col_names=c("community_served", "borrower", "pwsid", "project_id", 
                                                   "project_name", "project_type", "project_cost", 
                                                   "requested_amount", "funding_amount", "principal_forgiveness", 
                                                   "population", "project_description", "disadvantaged", 
                                                   "project_rank", "project_score", "expecting_funding" ))

```

## Create Numeric Columns

### Project Counts
```{r}
# NOTE: replacing NAs for math when calculating percents, 
# but when this merges in, we need to go back and add "Missing" when it was actually NA in the project data
# for project_type calculations, NA represents when we didn't have this data
project_counts <- all_projects %>%
  group_by(state, state_fiscal_year) %>%
  summarize(projects = n(),

            # number of projects expecting funding and percent of all projects expecting funding or not or where info is missing
            ef_projects = replace_na(sum(expecting_funding == "Yes"),0),
            nef_projects = replace_na(sum(expecting_funding == "No"),0),
            ef_ni_projects = replace_na(sum(expecting_funding == "No Information"),0),

            # number of disadvantaged projects and percent of all projects that are DACs
            dac_projects = replace_na(sum(disadvantaged == "Yes"),0),
            ndac_projects = replace_na(sum(disadvantaged == "No"),0),
            dac_ni_projects = replace_na(sum(disadvantaged == "No Information"),0),
            
            # dacs x expecting funding outcomes
            dac_ef_projects = replace_na(sum(disadvantaged == "Yes" & expecting_funding == "Yes"),0),
            dac_nef_projects = replace_na(sum(disadvantaged == "Yes" & expecting_funding=="No"),0),
            dac_ef_ni_projects = replace_na(sum(disadvantaged=="Yes" & expecting_funding=="No Information"),0),
            
            # non-dacs x expecting funding outcomes
            ndac_ef_projects = replace_na(sum(disadvantaged=="No" & expecting_funding=="Yes"),0),
            ndac_nef_projects = replace_na(sum(disadvantaged=="No" & expecting_funding=="No"),0),
            ndac_ef_ni_projects = replace_na(sum(disadvantaged=="No" & expecting_funding=="No Information"),0),
            
            # no info dacs x expecting funding outcomes
            dac_ni_ef_projects = replace_na(sum(disadvantaged=="No Information" & expecting_funding=="Yes"),0),
            dac_ni_nef_projects = replace_na(sum(disadvantaged=="No Information" & expecting_funding=="No"),0),
            dac_ni_ef_ni_projects = replace_na(sum(disadvantaged=="No Information" & expecting_funding=="No Information"),0)
  )

# subset to Lead projects to identify inventory and replacement projects
# 2025-07-11: added survey to lead_type designation for lsli
# see: https://enviropolicyinno.slack.com/archives/C08LXGF02AE/p1752240945744229?thread_ts=1752092342.154789&cid=C08LXGF02AE
lsl_projects <- all_projects |>
  dplyr::filter(project_type=="Lead") |>
  dplyr::mutate(
    lead_type = dplyr::case_when(
    stringr::str_detect(tolower(project_description), lsli_str) & stringr::str_detect(tolower(project_description), lslr_str) ~ "both",
    stringr::str_detect(tolower(project_description), lsli_str) ~ "lsli",
    stringr::str_detect(tolower(project_description), lslr_str) ~ "lslr",
   # catch weird exceptions where replacement/inventory doesn't appear next to LSL but should still be marked lslr/i
   stringr::str_detect(tolower(project_description), "replacement") & stringr::str_detect(tolower(project_description), lead_str) ~ "lslr",
   stringr::str_detect(tolower(project_description), "inventory") & stringr::str_detect(tolower(project_description), lead_str) ~ "lsli",
    TRUE ~ "unknown"
  )) |>
    # lead sub type exceptions (after inspection of project description classified as "both" lead types) 
    dplyr::left_join(
      tibble::tibble(
        borrower = c("Mart", "Alice", "Dog Ridge WSC", "Texarkana", "Bandera", "Varner Creek UD", "New Orleans - Carrolton Water Works (Loan 3)", "Goliad", "Brazoria", "Alice", "Cumby", "Hempstead"),
        project_id = c("No Information", "No Information", "No Information", "No Information", "No Information", "No Information", "1071009-03-LSLR", "16050", "16081", "16057", "16082", "16083"),
        pwsid = c("TX1550005", "TX1250001", "TX0140044", "TX0190004", "TX0100012", "TX0200070", "LA1071009", "TX0880001", "TX0200003", "TX1250001", "TX1120001", "TX2370001"),
        project_type = c("Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead", "Lead"),
        project_score = c("48.49", "42.34", "40.42", "16.43", "15.91", "13.32", "12", "62.99", "58.17", "46.20", "29.29", "25.26"),
        state = c("Texas", "Texas", "Texas", "Texas", "Texas", "Texas", "Louisiana", "Texas", "Texas", "Texas", "Texas", "Texas"),
        state_fiscal_year = c("2023", "2023", "2023", "2023", "2023", "2023", "2025", "2025", "2025", "2025", "2025", "2025"),
        lead_type_new = c("lsli", "lsli", "lsli", "lsli", "lsli", "lsli", "lslr", "lslr", "lsli", "lsli", "lslr", "lslr")
        ) 
    ) |>
    dplyr::mutate(lead_type = if_else(lead_type == "both" & !is.na(lead_type_new), 
                              lead_type_new, 
                              lead_type)) |>
    dplyr::select(-lead_type_new)


lead_type_count_summary <- lsl_projects |>
  dplyr::group_by(state, state_fiscal_year, lead_type) |>
  dplyr::summarise(
    lslb_projects = sum(lead_type == "both"),
    lsli_projects = sum(lead_type == "lsli"),
    lslr_projects = sum(lead_type == "lslr"),
    lslu_projects = sum(lead_type == "unknown"),

    lslb_ef_projects = sum(lead_type == "both" & expecting_funding=="Yes"),
    lsli_ef_projects = sum(lead_type == "lsli" & expecting_funding=="Yes"),
    lslr_ef_projects = sum(lead_type == "lslr" & expecting_funding=="Yes"),
    lslu_ef_projects = sum(lead_type == "unknown" & expecting_funding=="Yes")
    ) |>
  dplyr::mutate(dplyr::across(dplyr::everything(), ~tidyr::replace_na(., 0))) |>
  dplyr::select(-lead_type) |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarise(
    lslb_projects = sum(lslb_projects),
    lsli_projects = sum(lsli_projects),
    lslr_projects = sum(lslr_projects),
    lslu_projects = sum(lslu_projects),

    lslb_ef_projects = sum(lslb_ef_projects),
    lsli_ef_projects = sum(lsli_ef_projects),
    lslr_ef_projects = sum(lslr_ef_projects),
    lslu_ef_projects = sum(lslu_ef_projects)
  )


project_counts <- project_counts |>
  dplyr::left_join(lead_type_count_summary)
```

### Small Community
```{r}
small_community_df <- all_projects %>%
  mutate(small_community = case_when(
    population == "No Information" ~ "No Information",
    is.na(population) ~ "No Information", 
    convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 ~ "Yes",
    TRUE ~ "No"
  )) %>%
  group_by(state, state_fiscal_year, small_community) %>%
  summarize(projects = n(),
            dac_projects = sum(disadvantaged=="Yes", na.rm=TRUE),
            ndac_projects = sum(disadvantaged=="No", na.rm=TRUE),
            dac_ni_projects = sum(disadvantaged=="No Information", na.rm=TRUE),
            
            ef_projects = sum(expecting_funding=="Yes", na.rm=TRUE),
            nef_projects = sum(expecting_funding=="No", na.rm=TRUE),
            ef_ni_projects = sum(expecting_funding=="No Information", na.rm=TRUE),
            
            dac_ef_projects = sum(disadvantaged=="Yes" & expecting_funding=="Yes", na.rm=TRUE),
            dac_nef_projects = sum(disadvantaged=="Yes" & expecting_funding=="No", na.rm=TRUE),
            dac_ef_ni_projects = sum(disadvantaged=="Yes" & expecting_funding=="No Information", na.rm=TRUE),
            
            ndac_ef_projects = sum(disadvantaged=="No" & expecting_funding=="Yes", na.rm=TRUE),
            ndac_nef_projects = sum(disadvantaged=="No" & expecting_funding=="No", na.rm=TRUE),
            ndac_ef_ni_projects = sum(disadvantaged=="No" & expecting_funding=="No Information", na.rm=TRUE),
            
            dac_ni_ef_projects = sum(disadvantaged=="No Information" & expecting_funding=="Yes", na.rm=TRUE),
            dac_ni_nef_projects = sum(disadvantaged=="No Information" & expecting_funding=="No", na.rm=TRUE),
            dac_ni_ef_ni_projects = sum(disadvantaged=="No Information" & expecting_funding=="No Information", na.rm=TRUE),
            )
  
```

### Project Type
```{r}
project_type_counts <- all_projects %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(projects = n(),
            dac_projects = sum(disadvantaged=="Yes", na.rm=TRUE),
            ndac_projects = sum(disadvantaged=="No", na.rm=TRUE),
            dac_ni_projects = sum(disadvantaged=="No Information", na.rm=TRUE),
            
            ef_projects = sum(expecting_funding=="Yes", na.rm=TRUE),
            nef_projects = sum(expecting_funding=="No", na.rm=TRUE),
            ef_ni_projects = sum(expecting_funding=="No Information", na.rm=TRUE),
            
            dac_ef_projects = sum(disadvantaged=="Yes" & expecting_funding=="Yes", na.rm=TRUE),
            dac_nef_projects = sum(disadvantaged=="Yes" & expecting_funding=="No", na.rm=TRUE),
            dac_ef_ni_projects = sum(disadvantaged=="Yes" & expecting_funding=="No Information", na.rm=TRUE),
            
            ndac_ef_projects = sum(disadvantaged=="No" & expecting_funding=="Yes", na.rm=TRUE),
            ndac_nef_projects = sum(disadvantaged=="No" & expecting_funding=="No", na.rm=TRUE),
            ndac_ef_ni_projects = sum(disadvantaged=="No" & expecting_funding=="No Information", na.rm=TRUE),
            
            dac_ni_ef_projects = sum(disadvantaged=="No Information" & expecting_funding=="Yes", na.rm=TRUE),
            dac_ni_nef_projects = sum(disadvantaged=="No Information" & expecting_funding=="No", na.rm=TRUE),
            dac_ni_ef_ni_projects = sum(disadvantaged=="No Information" & expecting_funding=="No Information", na.rm=TRUE)) 

```

### Project Cost
```{r}
project_cost_aggregation <- all_projects %>%
  filter(!is.na(project_cost) & project_cost != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost)),
            median_project_cost = median_wo_zero(as.numeric(project_cost)),
            avg_project_cost = mean(as.numeric(project_cost)))

pc_by_type <- all_projects %>%
  filter(!is.na(project_cost) & project_cost != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost)))


sc_projects <- all_projects %>%
  dplyr::mutate(small_community = case_when(
    population == "No Information" ~ "No Information",
    convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 ~ "Yes",
    TRUE ~ "No"
  ))

pc_by_sc <-  sc_projects |>
  dplyr::filter(!is.na(project_cost) & project_cost != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community) |>
  dplyr::summarize(total_project_cost = sum(as.numeric(project_cost)))


pc_by_ef_sc <- sc_projects |>
  dplyr::filter(!is.na(project_cost) & project_cost != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, expecting_funding) |>
  dplyr::summarize(total_project_cost = sum(as.numeric(project_cost))) %>%
  tidyr::pivot_wider(names_from="expecting_funding", values_from="total_project_cost") %>%
  rename(project_cost_ef = Yes,
         project_cost_nef = No,
         project_cost_ef_ni = `No Information`,
         project_cost_ef_na = `NA`
         )


pc_by_dac_sc <- sc_projects |>
  dplyr::filter(!is.na(project_cost) & project_cost != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, disadvantaged) |>
  dplyr::summarize(total_project_cost = sum(as.numeric(project_cost))) |>
  tidyr::pivot_wider(names_from="disadvantaged", values_from="total_project_cost") %>%
  rename(project_costs_dac = Yes,
         project_costs_ndac = No,
         project_costs_dac_ni = `No Information`,
         project_costs_dac_na = `NA`
         ) 
 
pc_by_dac_ef_sc <- sc_projects |>
  dplyr::filter(!is.na(project_cost) & project_cost != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, disadvantaged, expecting_funding) |>
  dplyr::summarize(total_project_cost = sum(as.numeric(project_cost))) |>
  tidyr::pivot_wider(names_from=c("expecting_funding", "disadvantaged"), values_from="total_project_cost")|> 
  dplyr::rename(
    project_costs_ef_dac = Yes_Yes,
    project_costs_ef_ndac = Yes_No,
    project_costs_ef_dac_ni = `Yes_No Information`,
    # project_costs_ef_dac_na = `Yes_NA`,

    project_costs_nef_dac = No_Yes,
    project_costs_nef_ndac = No_No,
    project_costs_nef_dac_ni = `No_No Information`,
    # project_costs_nef_dac_na = `No_NA`,

    project_costs_ef_ni_dac = `No Information_Yes`,
    project_costs_ef_na_dac_na = `NA_NA`,
    project_costs_ef_na_dac = `NA_Yes`,
    project_costs_ef_na_ndac = `NA_No`
    ) 

# filter project costs by expecting funding for both tables
pc_by_ef <- all_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year, expecting_funding) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost))) %>%
  pivot_wider(names_from="expecting_funding", values_from="total_project_cost") %>%
  rename(project_cost_ef = Yes,
         project_cost_nef = No,
         project_cost_ef_ni = `No Information`,
         project_cost_ef_na = `NA`
         )

pc_by_dac <- all_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year,disadvantaged) %>%
  summarize(project_cost_dac = sum(as.numeric(project_cost)), .groups = "drop") |>
  tidyr::pivot_wider(names_from="disadvantaged", values_from="project_cost_dac") %>%
  rename(project_cost_dac = Yes,
         project_cost_ndac = No,
         project_cost_dac_ni = `No Information`,
         project_cost_dac_na = `NA`
         )

pc_by_dac_type <- all_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year,disadvantaged, project_type) %>%
  filter(disadvantaged == "Yes") %>%
  summarize(project_cost_dac = sum(as.numeric(project_cost))) %>%
  ungroup() %>%
  select(state, state_fiscal_year, project_cost_dac, project_type)

pc_by_ef_dac <- all_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year, expecting_funding, disadvantaged) %>%
  filter(expecting_funding == "Yes" & disadvantaged == "Yes") %>%
  summarize(project_cost_ef_dac = sum(as.numeric(project_cost))) %>%
  ungroup() %>%
  select(state, state_fiscal_year, project_cost_ef_dac)

pc_by_ef_type <- all_projects %>%
  filter(!is.na(project_cost) & project_cost != "No Information") %>%
  group_by(state, state_fiscal_year, project_type, expecting_funding) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost))) %>%
  filter(expecting_funding == "Yes") %>%
  select(-expecting_funding) %>%
  rename(total_project_cost_ef = total_project_cost)



# group by lsl project type for project cost calculations
pc_by_lsl <- lsl_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year, lead_type) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost))) %>%
  tidyr::pivot_wider(
    names_from = "lead_type", 
    values_from = "total_project_cost", 
    values_fill = 0,
    names_prefix = "project_cost_",
    names_glue = "project_cost_{ifelse(lead_type == 'both', 'lslb', ifelse(lead_type == 'lslr', 'lslr', ifelse(lead_type == 'lsli', 'lsli', ifelse(lead_type == 'unknown', 'lslu', lead_type))))}"
  )
  # pivot_wider(names_from="lead_type", values_from="total_project_cost") %>%
  # rename(
  #       project_cost_lslb = both,
  #       project_cost_lslr = lslr,
  #        project_cost_lsli = lsli,
  #        project_cost_lslu = unknown)

# group by lsl project type AND whether they are expecting funding or not
pc_by_lsl_ef <- lsl_projects %>%
  filter(!is.na(project_cost) & project_cost !="No Information") %>%
  group_by(state, state_fiscal_year, expecting_funding, lead_type) %>%
  summarize(total_project_cost = sum(as.numeric(project_cost))) %>% 
  filter(expecting_funding == "Yes") %>%
  tidyr::pivot_wider(
    names_from = "lead_type", 
    values_from = "total_project_cost", 
    values_fill = 0,
    names_prefix = "project_cost_ef",
    names_glue = "project_cost_ef_{ifelse(lead_type == 'both', 'lslb', ifelse(lead_type == 'lslr', 'lslr', ifelse(lead_type == 'lsli', 'lsli', ifelse(lead_type == 'unknown', 'lslu', lead_type))))}"
  ) |>
  # pivot_wider(names_from="lead_type", values_from="total_project_cost") %>%
  # rename(
  #   project_cost_ef_lslb = both, 
  #   project_cost_ef_lslr = lslr,
  #        project_cost_ef_lsli = lsli,
  #        project_cost_ef_lslu = unknown) %>%
  ungroup() %>%
  select(-expecting_funding)

```

### Project Funding
```{r}
funding_amount_aggregation <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount)),
            median_funding_amount = median_wo_zero(as.numeric(funding_amount)),
            avg_funding_amount = mean(as.numeric(funding_amount)))

# filter project costs by expecting funding
fa_by_ef <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount !="No Information") %>%
  group_by(state, state_fiscal_year, expecting_funding) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount))) %>%
  pivot_wider(names_from="expecting_funding", values_from="total_funding_amount") %>%
  rename(funding_amount_ef = Yes,
         funding_amount_nef = No)

# filter funding amount by expecting funding and dac status
fa_by_ef_dac <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount !="No Information") %>%
  group_by(state, state_fiscal_year, expecting_funding, disadvantaged) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount), na.rm=TRUE)) %>%
  filter(expecting_funding == "Yes" & disadvantaged %in% c("Yes", "No", "No Information")) %>%
  pivot_wider(names_from="disadvantaged", values_from="total_funding_amount") %>%
  rename(funding_amount_ef_dac = Yes,
         funding_amount_ef_ndac = No,
         funding_amount_ef_dac_ni = `No Information`) %>%
  ungroup() %>%
  select(state, state_fiscal_year, funding_amount_ef_dac, funding_amount_ef_ndac, funding_amount_ef_dac_ni)

fa_by_type <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount)))

fa_by_type_dac <- all_projects %>%
  filter(!is.na(funding_amount) & funding_amount !="No Information") %>%
  group_by(state, state_fiscal_year, disadvantaged, project_type) %>%
  summarize(total_funding_amount = sum(as.numeric(funding_amount), na.rm=TRUE)) %>%
  filter(disadvantaged %in% c("Yes", "No", "No Information")) %>%
  pivot_wider(names_from="disadvantaged", values_from="total_funding_amount") %>%
  rename(funding_amount_dac = Yes,
         funding_amount_ndac = No,
         funding_amount_dac_ni = `No Information`) %>%
  ungroup() %>%
  select(state, state_fiscal_year, project_type, funding_amount_dac, funding_amount_ndac, funding_amount_dac_ni)


fa_by_lsl <- lsl_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount !="No Information") |>
  dplyr::group_by(state, state_fiscal_year, lead_type) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount), na.rm=TRUE), .groups = "drop") |>
   tidyr::pivot_wider(
    names_from = "lead_type", 
    values_from = "total_funding_amount", 
    values_fill = 0,
    names_prefix = "funding_amount_",
    names_glue = "funding_amount_{ifelse(lead_type == 'both', 'lslb', ifelse(lead_type == 'lslr', 'lslr', ifelse(lead_type == 'lsli', 'lsli', ifelse(lead_type == 'unknown', 'lslu', lead_type))))}"
  ) |>
#   tidyr::pivot_wider(names_from = "lead_type", values_from = "total_funding_amount", values_fill = 0) |>
# dplyr::mutate(
#     funding_amount_lslb = if(exists("both", inherits = FALSE)) both else NA_real_,
#     funding_amount_lslr = if(exists("lslr", inherits = FALSE)) lslr else NA_real_,
#     funding_amount_lsli = if(exists("lsli", inherits = FALSE)) lsli else NA_real_,
#     funding_amount_lslu = if(exists("unknown", inherits = FALSE)) unknown else NA_real_
#   ) |>
  dplyr::select(state, state_fiscal_year, 
  dplyr::any_of(c("funding_amount_lslb", 
                             "funding_amount_lslr", 
                             "funding_amount_lsli", 
                             "funding_amount_lslu")))

fa_by_lsl_ef <-  lsl_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount !="No Information") |>
  dplyr::group_by(state, state_fiscal_year, expecting_funding, lead_type) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount), na.rm=TRUE), .groups = "drop") |>
  dplyr::filter(expecting_funding == "Yes") |>
   tidyr::pivot_wider(
    names_from = "lead_type", 
    values_from = "total_funding_amount", 
    values_fill = 0,
    names_prefix = "funding_amount_ef_",
    names_glue = "funding_amount_ef_{ifelse(lead_type == 'both', 'lslb', ifelse(lead_type == 'lslr', 'lslr', ifelse(lead_type == 'lsli', 'lsli', ifelse(lead_type == 'unknown', 'lslu', lead_type))))}"
  ) |> 
 # tidyr::pivot_wider(names_from = "lead_type", values_from = "total_funding_amount", values_fill = 0) |>
  # dplyr::mutate(
  #   funding_amount_ef_lslb = if("both" %in% dplyr::cur_data()) both else NA_real_,
  #   funding_amount_ef_lslr = if("lslr" %in% dplyr::cur_data()) lslr else NA_real_,
  #   funding_amount_ef_lsli = if("lsli" %in% dplyr::cur_data()) lsli else NA_real_,
  #   funding_amount_ef_lslu = if("unknown" %in% dplyr::cur_data()) unknown else NA_real_
  # ) |>
  dplyr::select(state, state_fiscal_year, 
  dplyr::any_of(c("funding_amount_ef_lslb", 
                             "funding_amount_ef_lslr", 
                             "funding_amount_ef_lsli", 
                             "funding_amount_ef_lslu")))


fa_by_sc <- sc_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount)))

fa_by_ef_sc <- sc_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, expecting_funding) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount))) |>
  tidyr::pivot_wider(names_from="expecting_funding", values_from="total_funding_amount") |>
  rename(funding_amount_ef = Yes,
         funding_amount_nef = No)

fa_by_dac_sc <- sc_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, disadvantaged) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount))) |>
  tidyr::pivot_wider(names_from="disadvantaged", values_from="total_funding_amount") |>
  rename(funding_amount_dac = Yes,
         funding_amount_ndac = No,
         funding_amount_dac_ni = `No Information`,
         funding_amount_dac_na = `NA`
         )

fa_by_dac_ef_sc <- sc_projects |>
  dplyr::filter(!is.na(funding_amount) & funding_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, small_community, disadvantaged, expecting_funding) |>
  dplyr::summarize(total_funding_amount = sum(as.numeric(funding_amount)), .groups = "drop") |>
  tidyr::pivot_wider(names_from= c("expecting_funding", "disadvantaged"), values_from="total_funding_amount") |>
  dplyr::rename(
    funding_amount_ef_dac = Yes_Yes,
    funding_amount_ef_ndac = Yes_No,
    funding_amount_ef_dac_ni = `Yes_No Information`,
    funding_amount_ef_dac_na = `Yes_NA`,

    funding_amount_nef_dac = No_Yes,
    funding_amount_nef_ndac = No_No,
    # funding_amount_nef_dac_ni = `No_No Information`,
    # funding_amount_nef_dac_na = `No_NA`
  )
```

```{r}
requested_amount_aggregation <- all_projects %>%
  filter(!is.na(requested_amount) & requested_amount != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_requested_amount = sum(as.numeric(requested_amount)),
            median_requested_amount = median_wo_zero(as.numeric(requested_amount)),
            avg_requested_amount = mean(as.numeric(requested_amount)))

ra_by_ef <- all_projects |>
  dplyr::filter(!is.na(requested_amount) & requested_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, expecting_funding) |>
  dplyr::summarize(total_requested_amount_ef = sum(as.numeric(requested_amount), na.rm = TRUE))|> 
  dplyr::filter(expecting_funding == "Yes") |>
  dplyr::select(-expecting_funding) 

ra_by_type <- all_projects %>%
  filter(!is.na(requested_amount) & requested_amount != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_requested_amount = sum(as.numeric(requested_amount)))

ra_by_ef_type <- all_projects |>
  dplyr::filter(!is.na(requested_amount) & requested_amount != "No Information") |>
  dplyr::group_by(state, state_fiscal_year, project_type, expecting_funding) |>
  dplyr::summarize(total_requested_amount_ef = sum(as.numeric(requested_amount), na.rm = TRUE))|> 
  dplyr::filter(expecting_funding == "Yes") |>
  dplyr::select(-expecting_funding) 

```

### Principal Forgiveness
```{r}

principal_forgiveness_aggregation <- all_projects %>%
  filter(!is.na(principal_forgiveness) & principal_forgiveness != "No Information") %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_principal_forgiveness = sum(as.numeric(principal_forgiveness)),
            median_principal_forgiveness = median_wo_zero(as.numeric(principal_forgiveness)),
            avg_principal_forgiveness = mean(as.numeric(principal_forgiveness))
            )

pf_projects <- all_projects %>%
  group_by(state, state_fiscal_year) %>%
  summarize(projects_receiving_pf = sum(as.numeric(principal_forgiveness > 0 )),
            dac_receiving_pf = sum(disadvantaged == "Yes" & as.numeric(principal_forgiveness) > 0),
            ndac_receiving_pf = sum(disadvantaged == "No" & as.numeric(principal_forgiveness) > 0),
            dac_ni_receiving_pf = sum(disadvantaged == "No Information" & as.numeric(principal_forgiveness) > 0),
            pf_ni = sum(principal_forgiveness == "No Information"),
            pf_na = sum(is.na(principal_forgiveness))
            )

pf_by_type <- all_projects %>%
  filter(!is.na(principal_forgiveness) & principal_forgiveness != "No Information") %>%
  group_by(state, state_fiscal_year, project_type) %>%
  summarize(total_principal_forgiveness = sum(as.numeric(principal_forgiveness)))
  
```

```{r}

project_dollars <- feature_coverage %>%
  select(state, state_fiscal_year) %>%
  left_join(funding_amount_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(fa_by_ef, by=c("state", "state_fiscal_year")) %>%
  left_join(fa_by_ef_dac, by=c("state", "state_fiscal_year")) %>%
  left_join(fa_by_lsl, by=c("state", "state_fiscal_year")) |>
  left_join(fa_by_lsl_ef, by=c("state", "state_fiscal_year")) |>
  left_join(principal_forgiveness_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(project_cost_aggregation, by=c("state", "state_fiscal_year")) %>%
  left_join(pc_by_ef, by=c("state", "state_fiscal_year")) %>%
  left_join(pc_by_dac, by=c("state", "state_fiscal_year")) %>%
  left_join(pc_by_ef_dac, by=c("state", "state_fiscal_year")) %>%
  left_join(pc_by_lsl, by=c("state", "state_fiscal_year")) %>%
  left_join(pc_by_lsl_ef , by=c("state", "state_fiscal_year")) %>%
  left_join(requested_amount_aggregation, by=c("state", "state_fiscal_year")) |>
  left_join(ra_by_ef, by=c("state", "state_fiscal_year")) |>
  dplyr::arrange(state, state_fiscal_year)

project_type_df <- project_type_counts %>%
  left_join(fa_by_type, by=c("state", "state_fiscal_year", "project_type")) %>%
  left_join(pc_by_type, by=c("state", "state_fiscal_year", "project_type")) %>% 
  left_join(pc_by_ef_type, by=c("state", "state_fiscal_year", "project_type")) %>%
  left_join(ra_by_type, by=c("state", "state_fiscal_year", "project_type")) %>%
  left_join(ra_by_ef_type, by=c("state", "state_fiscal_year", "project_type")) |>
  left_join(pf_by_type, by=c("state", "state_fiscal_year", "project_type")) |>
  left_join(pc_by_dac_type, by=c("state", "state_fiscal_year", "project_type")) |>
  left_join(fa_by_type_dac,  by=c("state", "state_fiscal_year", "project_type")) |>
  dplyr::arrange(state, state_fiscal_year)
  
project_counts <- project_counts %>%
  left_join(pf_projects, by=c("state", "state_fiscal_year")) |>
  dplyr::arrange(state, state_fiscal_year)

small_community_df <- small_community_df |>
  left_join(pc_by_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(pc_by_ef_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(pc_by_dac_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(pc_by_dac_ef_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(fa_by_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(fa_by_ef_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(fa_by_dac_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  left_join(fa_by_dac_ef_sc, by=c("state", "state_fiscal_year", "small_community") ) |>
  dplyr::arrange(state, state_fiscal_year)
```

## Exporting AWS
```{r}

# Writing to temp
write.csv(feature_coverage, file.path(tempdir(), "dwsrf-funding-tracker-feature-coverage.csv"), row.names = FALSE)

#Putting in Bucket
put_object(
  file = file.path(tempdir(), "dwsrf-funding-tracker-feature-coverage.csv"),
  object = "state-stats/dwsrf-funding-tracker-feature-coverage.csv",
  bucket = "funding-tracker",
  acl = "public-read"
)

# Writing to temp
write.csv(project_counts, file.path(tempdir(), "dwsrf-funding-tracker-project-counts.csv"), row.names = FALSE)

#Putting in Bucket
put_object(
  file = file.path(tempdir(), "dwsrf-funding-tracker-project-counts.csv"),
  object = "state-stats/dwsrf-funding-tracker-project-counts.csv",
  bucket = "funding-tracker",
  acl = "public-read"
)

# Writing to temp
write.csv(project_type_df, file.path(tempdir(), "dwsrf-funding-tracker-project-types.csv"), row.names = FALSE)

#Putting in Bucket
put_object(
  file = file.path(tempdir(), "dwsrf-funding-tracker-project-types.csv"),
  object = "state-stats/dwsrf-funding-tracker-project-types.csv",
  bucket = "funding-tracker",
  acl = "public-read"
)

# Writing to temp
write.csv(small_community_df, file.path(tempdir(), "dwsrf-funding-tracker-small-community.csv"), row.names = FALSE)

#Putting in Bucket
put_object(
  file = file.path(tempdir(), "dwsrf-funding-tracker-small-community.csv"),
  object = "state-stats/dwsrf-funding-tracker-small-community.csv",
  bucket = "funding-tracker",
  acl = "public-read"
)

# Writing to temp
write.csv(project_dollars, file.path(tempdir(), "dwsrf-funding-tracker-project-dollars.csv"), row.names = FALSE)

#Putting in Bucket
put_object(
  file = file.path(tempdir(), "dwsrf-funding-tracker-project-dollars.csv"),
  object = "state-stats/dwsrf-funding-tracker-project-dollars.csv",
  bucket = "funding-tracker",
  acl = "public-read"
)

```

## Exporting Google Sheets
```{r}
write_sheet(feature_coverage, "https://docs.google.com/spreadsheets/d/1MtnflPSLXBvcPMGLmwwCtMe21wtXxGGKswzaE6m-GBw", sheet = "Feature Coverage")
write_sheet(project_type_df, "https://docs.google.com/spreadsheets/d/1MtnflPSLXBvcPMGLmwwCtMe21wtXxGGKswzaE6m-GBw", sheet = "Project Type")
write_sheet(small_community_df, "https://docs.google.com/spreadsheets/d/1MtnflPSLXBvcPMGLmwwCtMe21wtXxGGKswzaE6m-GBw", sheet="Small Community")
write_sheet(project_counts, "https://docs.google.com/spreadsheets/d/1MtnflPSLXBvcPMGLmwwCtMe21wtXxGGKswzaE6m-GBw", sheet="Project Counts")
write_sheet(project_dollars, "https://docs.google.com/spreadsheets/d/1MtnflPSLXBvcPMGLmwwCtMe21wtXxGGKswzaE6m-GBw", sheet="Project Dollars")

```

```{r}
```