---
title: "State Funding Pages"
author: "Phil Cork"
date: "2024-07-08"
---

This notebook is a draft of analysis for creating statistics and data viz for the SRF Funding Tracker's individual state funding pages (those which cover a single state fiscal year for a particular state.) It pulls data from both the project-level data and the pre-generated state-level aggregate data.

# Setup

## Set Variables

```{r, warning=FALSE}

state_name <- "Texas"
state_abbr <- "TX"

sfy <- "2023"

save_plots <- FALSE

```

## Imports

```{r}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))
```



### Import Projects & State Data

```{r}
# import project-level data
state_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_counts <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-counts.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_dollars <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-dollars.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_types <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

general <- project_types %>%
    filter(project_type == "General")

lead <- project_types %>%
    filter(project_type == "Lead")

ec <- project_types %>%
    filter(project_type == "Emerging Contaminants")

```


```{r}
financial <- get_financial(state_name) %>%
  filter(state_fiscal_year == sfy)

set_asides <- get_set_asides(state_name) %>%
  filter(state_fiscal_year == sfy)

pf <- get_pf(state_name) %>%
  filter(state_fiscal_year == sfy)
```


# Funding Data

## Grants Allocated

```{r}
grants <- financial %>%
  select(fed_cap_grant, ffy21_fcg, ffy22_fcg, ffy23_fcg, ffy24_fcg, ffy25_fcg, ffy26_fcg, reallotments) %>%
  pivot_longer(cols=c(ffy21_fcg, ffy22_fcg, ffy23_fcg, ffy24_fcg, ffy25_fcg, ffy26_fcg, reallotments)) %>%
  filter(!is.na(value) & value > 0) %>%
  mutate(name = str_to_sentence(str_remove(name, "_fcg")),
         name = str_replace(name, "Ffy", "FFY"),
         name = paste0(fed_cap_grant, ", ", name)
         )
```


```{r}

# display fed cap grants by source
if (sum(grants$value, na.rm=TRUE)>0) {
  
  grants_p <- ggplot(grants, aes(x=value, y=reorder(name, value), text=format_currency(value))) + 
    geom_bar(stat="identity", fill="#4ea324") +
    labs(x="", y="", 
         title="Federal Capitalization Grants", 
         subtitle=get_subtitle_str(c(sfy), state_name)) + 
    scale_x_continuous(labels=label_dollar()) + 
    epic_chart_theme
  
  grants_gp <-  ggplotly(grants_p, tooltip="text") %>%
    layout(title = list(text = paste0('Federal Capitalization Grants',
                                      '<br>',
                                      get_subtitle_str(c(sfy), state_name))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = grants_p,
                   gp_object = grants_gp,
                   name = paste0("fed-cap-grant-amount-",sfy))
    }
  
  grants_gp
}
```



## Principal Forgiveness

```{r}
total_pf_sfy <- pf %>%
    filter(fed_cap_grant != "Base & IIJA General Supplemental") %>%
    select(fed_cap_grant, ffy21_amt, ffy22_amt, ffy23_amt, ffy24_amt, ffy25_amt, ffy26_amt) %>%
  pivot_longer(cols=c(ffy21_amt, ffy22_amt, ffy23_amt, ffy24_amt, ffy25_amt, ffy26_amt)) %>%
  filter(!is.na(value) & value > 0) %>%
  mutate(name = str_to_upper(str_remove(name, "_amt")),
         name = paste0(fed_cap_grant, ", ", name)
         ) %>%
  group_by(name) %>%
  summarize(value = sum(value, na.rm=TRUE))

```


```{r}

if(sum(pf$total_pf_amt, na.rm=TRUE) > 0) {
  
  all_pf_p <- ggplot(pf, aes(x=total_pf_amt, y=reorder(pf_category, total_pf_amt), text=format_currency(total_pf_amt))) + 
    geom_bar(stat="identity", fill="#4ea324") +
    labs(x="", y="", title="Expected Principal Forgiveness from Fed Cap Grants", 
         subtitle=get_subtitle_str(c(sfy), state_name)) + 
    scale_x_continuous(labels=label_dollar()) + 
    epic_chart_theme
  
  all_pf_gp <-  ggplotly(all_pf_p, tooltip="text") %>%
    layout(title = list(text = paste0("Expected Principal Forgiveness from Fed Cap Grants",
                                      "<br>",
                                      get_subtitle_str(c(sfy), state_name))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = all_pf_p,
                   gp_object = all_pf_gp,
                   name = paste0("fed-cap-grant-pf-amount-",sfy))
    }
  
  all_pf_gp
}
```

## Base, Set Asides, PF Percents

```{r}

base_distribution <-data.frame(name=c(""), value=c(0))

base_distribution <- base_distribution %>%
  add_row(name="DAC PF", 
          value=pf$total_pf_amt[pf$pf_category=="DAC"]) %>%
  add_row(name="Discretionary PF",
          value=pf$total_pf_amt[pf$pf_category=="Discretionary"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Local Assistance & Other"])

base_distribution <- base_distribution %>%
  add_row(name="Remaining Base Grant, for Loans",
          value = max(financial$total_fcg[financial$fed_cap_grant=="Base"] - sum(base_distribution$value),0)) %>%
  filter(name!="") %>%
  mutate(name = factor(name, levels=c("Remaining Base Grant, for Loans", "Local Assistance & Other Set Asides", 
                                      "State Program Management Set Asides", "Small System TA Set Asides", 
                                      "Admin & TA Set Asides",  "Discretionary PF", "DAC PF")),
         plot_str = ifelse(value > 0, paste0("$", as.character(round(value/1000000,2)), "M"), ""))

```


```{r}
if (nrow(base_distribution %>% filter(!is.na(value))) > 0) {
  
  base_all_p <- ggdonutchart(base_distribution,
                             "value",
                             fill="name",
                             color="white",
                             palette=distribution_colors,
                             label="plot_str",
                             lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How is the Base grant allocated?",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  base_all_fig <- base_distribution %>% 
    plot_ly(labels = ~name, values = ~value,
            hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, 
            marker=list(colors = distribution_colors[as.character(base_distribution$name)])) %>% 
    layout(title = paste0("How is the Base grant allocated?", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),  
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  if (save_plots) {
    run_save_plots(gg_plot = base_all_p,
                   gp_object = base_all_fig,
                   name = paste0("base-pf-sa-pct-",sfy))
    }
  
  base_all_fig
}
```


## Gen Supp, Set Asides, PF Percents

```{r}

gensupp_distribution <-data.frame(name=c(""), value=c(0))

gensupp_distribution <- gensupp_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA Gen Supp"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Local Assistance & Other"])

gensupp_distribution <- gensupp_distribution %>%
  add_row(name="Remaining Gen Supp Grant, for Loans",
          value = max(financial$total_fcg[financial$fed_cap_grant=="IIJA Gen Supp"] - sum(gensupp_distribution$value, na.rm=TRUE),0)) %>%
  filter(name!="") %>%
  mutate(name = factor(name, levels=c("Remaining Gen Supp Grant, for Loans", 
                                      "Local Assistance & Other Set Asides", 
                                      "State Program Management Set Asides", 
                                      "Small System TA Set Asides", 
                                      "Admin & TA Set Asides",  "Principal Forgiveness")),
         plot_str = ifelse(value > 0, paste0("$", as.character(round(value/1000000,2)), "M"), ""))

```


```{r}
if (nrow(gensupp_distribution %>% filter(!is.na(value))) > 0) {
  
  gensupp_all_p <- ggdonutchart(gensupp_distribution,
                                "value",
                                fill="name",
                                color="white",
                                palette=distribution_colors,
                                label="plot_str",
                                lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How is the IIJA Gen Supp grant allocated?",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  gensupp_all_fig <- gensupp_distribution %>% 
    plot_ly(labels = ~name, values = ~value,
            hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = distribution_colors[as.character(gensupp_distribution$name)])) %>% 
    layout(title = paste0("How is the IIJA Gen Supp grant allocated?", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),  
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  if (save_plots) {
    run_save_plots(gg_plot = gensupp_all_p,
                   gp_object = gensupp_all_fig,
                   name = paste0("iija-gen-supp-pf-sa-pct-",sfy))
    }
  
  gensupp_all_fig
}
```

## Lead, Set Asides, PF Percents

```{r}

lead_distribution <-data.frame(name=c(""), value=c(0))

lead_distribution <- lead_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA - LSLR"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Local Assistance & Other"])

lead_distribution <- lead_distribution %>%
  add_row(name="Remaining LSLR Grant, for Loans",
          value = max(financial$total_fcg[financial$fed_cap_grant=="LSLR"] - sum(lead_distribution$value, na.rm=TRUE),0)) %>%
  filter(name!="") %>%
  mutate(name = factor(name, levels=c("Remaining LSLR Grant, for Loans", 
                                      "Local Assistance & Other Set Asides", 
                                      "State Program Management Set Asides", 
                                      "Small System TA Set Asides", 
                                      "Admin & TA Set Asides", 
                                      "Principal Forgiveness")),
         plot_str = ifelse(value > 0, paste0("$", as.character(round(value/1000000,2)), "M"), ""))

```


```{r}
if (nrow(lead_distribution %>% filter(!is.na(value))) > 0) {
  
  lead_all_p <- ggdonutchart(lead_distribution,
                             "value",
                             fill="name",
                             color="white",
                             palette=distribution_colors,
                             label="plot_str",
                             lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How is the LSLR grant allocated?",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  lead_all_fig <- lead_distribution %>% 
    plot_ly(labels = ~name, values = ~value,
            hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = distribution_colors[as.character(lead_distribution$name)])) %>% 
    layout(title = paste0("How is the LSLR grant allocated?", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),  
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  if (save_plots) {
    run_save_plots(gg_plot = lead_all_p,
                   gp_object = lead_all_fig,
                   name = paste0("lead-pf-sa-pct-",sfy)) }
  
  lead_all_fig
}
```


## EC, Set Asides, PF Percents

```{r}

ec_distribution <-data.frame(name=c(""), value=c(0))

ec_distribution <- ec_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA - EC"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Local Assistance & Other"])

ec_distribution <- ec_distribution %>%
  add_row(name="Remaining EC Grant, for Loans",
          value = max(financial$total_fcg[financial$fed_cap_grant=="EC"] - sum(ec_distribution$value, na.rm=TRUE),0)) %>%
  filter(name!="") %>%
  mutate(name = factor(name, levels=c("Remaining EC Grant, for Loans", 
                                      "Local Assistance & Other Set Asides", 
                                      "State Program Management Set Asides", 
                                      "Small System TA Set Asides", 
                                      "Admin & TA Set Asides",  
                                      "Principal Forgiveness")),
         plot_str = ifelse(value > 0, paste0("$", as.character(round(value/1000000,2)), "M"), ""))

```


```{r}
if (nrow(ec_distribution %>% filter(!is.na(value))) > 0) {
  
  ec_all_p <- ggdonutchart(ec_distribution,
                           "value",
                           fill="name",
                           color="white",
                           palette=distribution_colors,
                           label="plot_str",
                           lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How is the EC grant allocated?",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  ec_all_fig <- ec_distribution %>% 
    plot_ly(labels = ~name, values = ~value,
            hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = distribution_colors[as.character(ec_distribution$name)])) %>% 
    layout(title = paste0("How is the EC grant allocated?", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),  
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  if (save_plots) {
    run_save_plots(gg_plot = ec_all_p,
                   gp_object = ec_all_fig,
                   name = paste0("ec-pf-sa-pct-",sfy))
    }
  
  ec_all_fig
}
```

# Comparing Funding and Project Data

## Funds by Process

```{r}

if (!is.na(project_dollars$total_project_cost)) {
  
  funds_process_comp <- data.frame(name=c("Available to Fund Projects", 
                                          "Project Demand", "Expected Funding"),
                                   value=c(sum(financial$total_funding_available, na.rm = TRUE),
                                           project_dollars$total_project_cost,
                                           project_dollars$total_funding_amount)) %>%
    mutate(category = factor(name, levels=c("Expected Funding", 
                                            "Available to Fund Projects", "Project Demand")))
  
} else {
  
  funds_process_comp <- data.frame(name=c("Available to Fund Projects", 
                                          "Project Demand", "Expected Funding"),
                                   value=c(sum(financial$total_funding_available, na.rm = TRUE),
                                           project_dollars$total_requested_amount,
                                           project_dollars$total_funding_amount)) %>%
    mutate(category = factor(name, levels=c("Expected Funding", 
                                            "Available to Fund Projects", "Project Demand"))) 
  
}
```


```{r}

if(nrow(funds_process_comp %>% filter(!is.na(value))) == nrow(funds_process_comp)) {
  
  funds_comp_p <- ggplot(funds_process_comp, 
                         aes(y=category, x=value, text=format_currency(value))) + 
    geom_bar(stat="identity", fill="#4ea324") +
    labs(x="", y="", 
         title="How do available funds compare to demand and funding expectations?", 
         subtitle=get_subtitle_str(c(sfy), state_name)) + 
    scale_x_continuous(labels=label_dollar()) + 
    epic_chart_theme
  
  funds_comp_gp <- ggplotly(funds_comp_p, tooltip="text") %>%
    layout(title = list(text = paste0("How do available funds compare to demand and funding expectations?", '<br>',get_subtitle_str(c(sfy), state_name))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = funds_comp_p,
                   gp_object = funds_comp_gp,
                   name = paste0("compare-available-demand-expecting-",sfy)) }
  
  funds_comp_gp
}
```

## Project Costs & Available Funds

### General

```{r}
# even though gen funds in 2022, not making plots
if (sfy != "2022") {
  
  base_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="Base"]),
                                 0,financial$total_funding_available[financial$fed_cap_grant=="Base"])
  
  iija_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="IIJA Gen Supp"]),
                                 0,financial$total_funding_available[financial$fed_cap_grant=="IIJA Gen Supp"])
  
  base_iija_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="Base & IIJA Gen Supp"]),
                                      0,financial$total_funding_available[financial$fed_cap_grant=="Base & IIJA Gen Supp"])
  
  
  gen_funds_available <- max(base_iija_funds_available, base_funds_available+iija_funds_available)
  
  # if general project cost is missing, switch to requested amount
  # NOTE may manually need to use Requested Amount instead when both are present but the latter is preferred
  
  if (!is.na(general$total_project_cost)) {
    gen_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                            value=c(gen_funds_available,
                                    max(general$total_project_cost - gen_funds_available,0)))
  } else {
    gen_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                            value=c(gen_funds_available,
                                    max(general$total_requested_amount - gen_funds_available,0)))
  }
  
  gen_costs <- gen_costs %>%
    mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
           value_str = format_currency(value),
           pct_value = round(value / sum(gen_costs$value),2),
           pct_value_str = format_percent(pct_value))
  
  # if we have gen data for given sfy, make plot
  if ( nrow(gen_costs %>% filter(!is.na(value))) == nrow(gen_costs))  {
    
    gen_comp_p <- ggdonutchart(gen_costs,
                               "value",
                               fill="name",
                               color="white",
                               palette=comp_colors,
                               label="value_str",
                               lab.pos="out") + 
      theme(
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_blank(), 
        legend.position = "right",  
        legend.direction = "vertical") +
      guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
      ) + 
      labs(title="Unmet demand compared to available Base & IIJA Gen Supp funds",
           subtitle=get_subtitle_str(c(sfy), state_name))
    
    
    gen_comp_fig <- gen_costs %>% plot_ly(labels = ~name, values = ~value,
                                          hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
      add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = comp_colors[as.character(gen_costs$name)])) %>% 
      layout(title = paste0("Unmet demand compared to available Base & IIJA Gen Supp funds", "<br>", get_subtitle_str(c(sfy), state_name)),  
             margin = list(t = 100, l = 10),  
             xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
    
    
    if (save_plots) {
      run_save_plots(gg_plot = gen_comp_p,
                     gp_object = gen_comp_fig,
                     name = paste0("compare-available-demand-general-",sfy))
      }
    
    gen_comp_fig
    
  } # if gen_costs > 0, make plot
  
} # if sfy!=2022
```

### Lead

```{r}
# no lead funds before year 0, skip data and plot
if (sfy != "2022") {
  
  lead_funds_available <- financial$total_funding_available[financial$fed_cap_grant=="LSLR"]
  
  # if project costs is NA, use Requested Amount instead
  # NOTE: may need to manually change to requested amount if both are valid but the latter is more worth highlighting
  
  if (nrow(lead) > 0) {
  
  if (!is.na(lead$total_project_cost)) {
    
    lead_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                             value=c(lead_funds_available,
                                     max(lead$total_project_cost - lead_funds_available, 0)))
  } else {
    lead_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                             value=c(lead_funds_available,
                                     max(lead$total_requested_amount - lead_funds_available, 0)))
  }
  
  lead_costs <- lead_costs %>%
    mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
           value_str = format_currency(value),
           pct_value = round(value / sum(lead_costs$value),2),
           pct_value_str = format_percent(pct_value))
  
  # if we have lead data for given sfy, make plot
  if ( nrow(lead_costs %>% filter(!is.na(value))) == nrow(lead_costs))  {
    
    lead_comp_p <- ggdonutchart(lead_costs,
                                "value",
                                fill="name",
                                color="white",
                                palette=comp_colors,
                                label="value_str",
                                lab.pos="out") + 
      theme(
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_blank(), 
        legend.position = "right",  
        legend.direction = "vertical") +
      guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
      ) + 
      labs(title="Unmet demand compared to available LSLR funds",
           subtitle=get_subtitle_str(c(sfy), state_name))
    
    
    lead_comp_fig <- lead_costs %>% plot_ly(labels = ~name, values = ~value,
                                            hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
      add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = comp_colors[as.character(lead_costs$name)])) %>% 
      layout(title = paste0("Unmet demand compared to available LSLR funds", "<br>", get_subtitle_str(c(sfy), state_name)),  
             margin = list(t = 100, l = 10),  
             xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
    
    
    if (save_plots) {
      run_save_plots(gg_plot = lead_comp_p,
                     gp_object = lead_comp_fig,
                     name = paste0("compare-available-demand-lead-",sfy))
      }
    
    
    lead_comp_fig
    
  } # if lead_costs > 0, make plot
  
  } # if lead has length
} # if sfy!=2022, check for lead


```

### EC
```{r}

# no ec funds in y0
if (sfy != "2022") {
  
  ec_funds_available <- financial$total_funding_available[financial$fed_cap_grant=="EC"]
  
  ec <- project_types %>%
    filter(project_type == "Emerging Contaminants")
  
  if (nrow(ec) > 0) {
  
  if (!is.na(ec$total_project_cost)) {
    ec_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                           value=c(ec_funds_available,
                                   max(ec$total_project_cost - ec_funds_available,0)))
  } else {
    ec_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                           value=c(ec_funds_available,
                                   max(ec$total_requested_amount - ec_funds_available,0)))
  }
  
  ec_costs <- ec_costs %>%
    mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
           value_str = format_currency(value),
           pct_value = round(value / sum(ec_costs$value),2),
           pct_value_str = format_percent(pct_value))
  
  # if we have ec data for sfy, make plot
  if ( nrow(ec_costs %>% filter(!is.na(value))) == nrow(ec_costs))  {
    
    ec_comp_p <- ggdonutchart(ec_costs,
                              "value",
                              fill="name",
                              color="white",
                              palette=comp_colors,
                              label="value_str",
                              lab.pos="out") + 
      theme(
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_blank(), 
        legend.position = "right",  
        legend.direction = "vertical") +
      guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
      ) + 
      labs(title="Unmet demand compared to available EC funds",
           subtitle=get_subtitle_str(c(sfy), state_name))
    
    
    ec_comp_fig <- ec_costs %>% plot_ly(labels = ~name, values = ~value,
                                        hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
      add_pie(hole = 0.6, direction="clockwise", sort=FALSE,marker=list(colors = comp_colors[as.character(ec_costs$name)])) %>% 
      layout(title = paste0("Unmet demand compared to available EC funds", "<br>", get_subtitle_str(c(sfy), state_name)),  
             margin = list(t = 100, l = 10),  
             xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
    
    if (save_plots) {
      run_save_plots(gg_plot = ec_comp_p,
                     gp_object = ec_comp_fig,
                     name = paste0("compare-available-demand-ec-",sfy))
      }
    
    ec_comp_fig
  } # end if we have ec data, make plot
  } # end if ec data
} # end if sfy !=2022
```

## DAC & EF Comparisons

### DACs EF / All DACs

```{r}
dacs_df <- data.frame(name=c("DACs, Expecting Funding", 
                             "DACs, Not Expecting Funding", "DACs, No Info"),
                      value=c(project_counts$dac_ef_projects, 
                              project_counts$dac_nef_projects, 
                              project_counts$dac_ef_ni_projects),
                      value_str=c(as.character(project_counts$dac_ef_projects),
                                  as.character(project_counts$dac_nef_projects), 
                                  as.character(project_counts$dac_ef_ni_projects))) %>%
  mutate(value_str = ifelse(value >0, value_str, ""),
         name = factor(name, levels=c("DACs, No Info", 
                                      "DACs, Not Expecting Funding", 
                                      "DACs, Expecting Funding")))


if ( nrow(dacs_df %>% filter(!is.na(name))) == nrow(dacs_df))  {
  
  dacs_ef_p <- ggdonutchart(dacs_df,
                            "value",
                            fill="name",
                            color="white",
                            palette=ef_dac_comp_colors,
                            label="value_str",
                            lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How many DACs could expect funding?",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  dacs_ef_fig <- dacs_df %>% plot_ly(labels = ~name, values = ~value,
                                     hovertemplate="%{label}<br>%{value} projects<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = ef_dac_comp_colors[as.character(dacs_df$name)])) %>% 
    layout(title = paste0("How many DACs could expect funding?", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = dacs_ef_p,
                   gp_object = dacs_ef_fig,
                   name = paste0("dacs-expecting-funding-",sfy))
    }
  
  dacs_ef_fig
}
```


### DACs / All EF

```{r}

ef_df <- data.frame(name=c("Expecting Funding, DAC", 
                               "Expecting Funding, Not DAC", 
                               "Expecting Funding, No Info"),
                         value=c(project_counts$dac_ef_projects,
                                  project_counts$ndac_ef_projects,
                                  project_counts$dac_ni_ef_projects),
                         value_str=c(as.character(project_counts$dac_ef_projects),
                                      as.character(project_counts$ndac_ef_projects),
                                      as.character(project_counts$dac_ni_ef_projects))
                                      ) %>%
  mutate(value_str = ifelse(value >0, value_str, ""),
         name = factor(name, levels=c("Expecting Funding, No Info",
                                      "Expecting Funding, Not DAC", 
                                      "Expecting Funding, DAC")))


if ( nrow(ef_df %>% filter(!is.na(value))) == nrow(ef_df))  {
  
  ef_df_p <- ggdonutchart(ef_df,
                          "value",
                          fill="name",
                          color="white",
                          palette=ef_dac_comp_colors,
                          label="value_str",
                          lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="How many projects expected to receive funding are for DACs",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  ef_df_fig <- ef_df %>% plot_ly(labels = ~name, values = ~value,
                                 hovertemplate = '%{label}<br>%{value} projects<extra></extra>') %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = ef_dac_comp_colors[as.character(ef_df$name)])) %>% 
    layout(title = paste0("How many projects expected to receive funding are for DACs?", "<br>", get_subtitle_str(c(sfy), state_name)),
           margin = list(t = 100, l = 10),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  if (save_plots) {
    run_save_plots(gg_plot = ef_df_p,
                   gp_object = ef_df_fig, 
                   name = paste0("expecting-funding-projects-dacs-",sfy))
    }
  
  ef_df_fig
}
```

### DAC/EF Heat Map
```{r}
dac_ef_hm <- project_counts %>%
  select(state_fiscal_year, 
         dac_ef_projects, dac_nef_projects, dac_ef_ni_projects,
         ndac_ef_projects, ndac_nef_projects, ndac_ef_ni_projects,
         dac_ni_ef_projects, dac_ni_nef_projects, dac_ni_ef_ni_projects
  ) %>%
  pivot_longer(cols=c("dac_ef_projects", "dac_nef_projects", "dac_ef_ni_projects",
                      "ndac_ef_projects", "ndac_nef_projects", "ndac_ef_ni_projects",
                      "dac_ni_ef_projects", "dac_ni_nef_projects", "dac_ni_ef_ni_projects")) %>%
  mutate(dac = case_when(
    name %in% c("dac_ef_projects", "dac_nef_projects", "dac_ef_ni_projects") ~ "Yes",
    name %in% c("ndac_ef_projects", "ndac_nef_projects", "ndac_ef_ni_projects") ~ "No",
    TRUE ~ "No Info"),
    expecting_funding = case_when(
      name %in% c("dac_ef_projects", "ndac_ef_projects", "dac_ni_ef_projects") ~ "Yes",
      name %in% c("dac_nef_projects", "ndac_nef_projects", "dac_ni_nef_projects") ~ "No",
      TRUE ~ "No Info") 
  ) %>%
  mutate(plot_str = paste0("DAC: ", dac, ", Expecting Funding: ", expecting_funding, "<br>",
                           value, " projects"),
         dac = factor(dac, levels=c("Yes", "No", "No Info")),
         expecting_funding = factor(expecting_funding, levels=c("Yes", "No", "No Info"))) %>%
  select(-name)

if (sum(dac_ef_hm$value, na.rm=TRUE) > 0) {
  
  dac_ef_hm_p <- ggplot(dac_ef_hm, aes(x=expecting_funding, y=dac, fill=value, text=plot_str)) + 
    geom_tile(color="white") + 
    labs(title="Projects by DAC Status and Expecting Funding", subtitle=get_subtitle_str(c(sfy), state_name),
         x="Expecting Funding", y="Disadvantaged", fill="Projects") + 
    scale_fill_gradient(low=ef_dac_hm_colors[1], high=ef_dac_hm_colors[2]) + 
    theme_minimal()
  
  dac_ef_hm_gp <- ggplotly(dac_ef_hm_p, tooltip="text") %>%
    layout(title = list(text = paste0("Projects by DAC Status and Expecting Funding",
                                      '<br>', get_subtitle_str(c(sfy), state_name))))
  
  
  if (save_plots) {
    
    run_save_plots(gg_plot = dac_ef_hm_p,
                   gp_object = dac_ef_hm_gp,
                   name = paste0("dac-expecting-funding-heatmap-",sfy))
    }
  
  dac_ef_hm_gp
}
```

## DAC by Project Type

### General

```{r}

if (nrow(general) > 0) {

 gen_dacs_df <- data.frame(name=c("DACs, Expecting Funding", 
                             "DACs, Not Expecting Funding", "DACs, No Info"),
                      value=c(general$dac_ef_projects, 
                              general$dac_nef_projects, 
                              general$dac_ef_ni_projects),
                      value_str=c(as.character(general$dac_ef_projects),
                                  as.character(general$dac_nef_projects), 
                                  as.character(general$dac_ef_ni_projects))) %>%
  mutate(value_str = ifelse(value >0, value_str, ""),
         name = factor(name, levels=c("DACs, No Info", 
                                      "DACs, Not Expecting Funding", 
                                      "DACs, Expecting Funding")))


if ( nrow(gen_dacs_df %>% filter(!is.na(name))) == nrow(gen_dacs_df))  {
  
  gen_dacs_ef_p <- ggdonutchart(gen_dacs_df,
                            "value",
                            fill="name",
                            color="white",
                            palette=ef_dac_comp_colors,
                            label="value_str",
                            lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="DACs Expecting Funding, General Projects",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  gen_dacs_ef_fig <- gen_dacs_df %>% plot_ly(labels = ~name, values = ~value,
                                     hovertemplate="%{label}<br>%{value} projects<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = ef_dac_comp_colors[as.character(gen_dacs_df$name)])) %>% 
    layout(title = paste0("DACs Expecting Funding, General Projects", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = gen_dacs_ef_p,
                   gp_object = gen_dacs_ef_fig,
                   name = paste0("dacs-expecting-funding-general-",sfy))
    }
  
  gen_dacs_ef_fig
}
}
```

### Lead

```{r}

if (nrow(lead) > 0) {

 lead_dacs_df <- data.frame(name=c("DACs, Expecting Funding", 
                             "DACs, Not Expecting Funding", "DACs, No Info"),
                      value=c(lead$dac_ef_projects, 
                              lead$dac_nef_projects, 
                              lead$dac_ef_ni_projects),
                      value_str=c(as.character(lead$dac_ef_projects),
                                  as.character(lead$dac_nef_projects), 
                                  as.character(lead$dac_ef_ni_projects))) %>%
  mutate(value_str = ifelse(value >0, value_str, ""),
         name = factor(name, levels=c("DACs, No Info", 
                                      "DACs, Not Expecting Funding", 
                                      "DACs, Expecting Funding")))


if ( nrow(lead_dacs_df %>% filter(!is.na(name))) == nrow(lead_dacs_df))  {
  
  lead_dacs_ef_p <- ggdonutchart(lead_dacs_df,
                            "value",
                            fill="name",
                            color="white",
                            palette=ef_dac_comp_colors,
                            label="value_str",
                            lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="DACs Expecting Funding, Lead Projects",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  lead_dacs_ef_fig <- lead_dacs_df %>% plot_ly(labels = ~name, values = ~value,
                                     hovertemplate="%{label}<br>%{value} projects<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = ef_dac_comp_colors[as.character(lead_dacs_df$name)])) %>% 
    layout(title = paste0("DACs Expecting Funding, Lead Projects", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = lead_dacs_ef_p,
                   gp_object = lead_dacs_ef_fig,
                   name = paste0("dacs-expecting-funding-lead-",sfy))
    }
  
  lead_dacs_ef_fig
}
}
```

### EC

```{r}

if (nrow(ec) > 0) {

 ec_dacs_df <- data.frame(name=c("DACs, Expecting Funding", 
                             "DACs, Not Expecting Funding", "DACs, No Info"),
                      value=c(ec$dac_ef_projects, 
                              ec$dac_nef_projects, 
                              ec$dac_ef_ni_projects),
                      value_str=c(as.character(ec$dac_ef_projects),
                                  as.character(ec$dac_nef_projects), 
                                  as.character(ec$dac_ef_ni_projects))) %>%
  mutate(value_str = ifelse(value >0, value_str, ""),
         name = factor(name, levels=c("DACs, No Info", 
                                      "DACs, Not Expecting Funding", 
                                      "DACs, Expecting Funding")))


if ( nrow(ec_dacs_df %>% filter(!is.na(name))) == nrow(ec_dacs_df))  {
  
  ec_dacs_ef_p <- ggdonutchart(ec_dacs_df,
                            "value",
                            fill="name",
                            color="white",
                            palette=ef_dac_comp_colors,
                            label="value_str",
                            lab.pos="out") + 
    theme(
      text = element_text(size = 11, family = "Lato"), 
      legend.text = element_text(size = 10, family = "Lato"), 
      legend.title = element_blank(), 
      legend.position = "right",  
      legend.direction = "vertical") +
    guides(fill = guide_legend(ncol = 1, byrow = TRUE, reverse=TRUE),
    ) + 
    labs(title="DACs Expecting Funding, EC Projects",
         subtitle=get_subtitle_str(c(sfy), state_name))
  
  
  ec_dacs_ef_fig <- ec_dacs_df %>% plot_ly(labels = ~name, values = ~value,
                                     hovertemplate="%{label}<br>%{value} projects<extra></extra>") %>%
    add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors = ef_dac_comp_colors[as.character(ec_dacs_df$name)])) %>% 
    layout(title = paste0("DACs Expecting Funding, EC Projects", "<br>", get_subtitle_str(c(sfy), state_name)),  
           margin = list(t = 100, l = 10),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = ec_dacs_ef_p,
                   gp_object = ec_dacs_ef_fig,
                   name = paste0("dacs-expecting-funding-ec-",sfy))
    }
  
  ec_dacs_ef_fig
}
}
```


## Project Score Range

```{r}
# range of project scores for projects expecting funding
tibble(
  state_projects %>%
    filter(expecting_funding == "Yes") %>%
    mutate(score_numeric = as.numeric(project_score)) %>%
    group_by(state, state_fiscal_year) %>%
    summarize(min_score = min(score_numeric),
              max_score = max(score_numeric))
)
```