
# Combined Funding + Projects

This notebook combines data used in the `funding-data-yoy` and `project-data-yoy` notebooks and combines them for data viz and analysis where the two overlap. These include comparing the total funding available to requested amounts or project costs to consider oversubscription and other considerations.

# Setup

## Local Variables

```{r}
# filters data and changes plot names
state_name <- "Pennsylvania"
state_abbr <- "PA"

# export plots to AWS and Drive?
save_plots <- FALSE

```

## Imports & Settings

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))

```

## Create Data Subsets

```{r, warning=FALSE}

financial <- get_financial(state_name)

financial_by_project_type <- financial |>
  dplyr::mutate(project_type = case_when(
    fed_cap_grant %in% c("LSLR", "LSLR_FFY22","LSLR_FFY23") ~ "Lead",
    fed_cap_grant %in% c("EC", "EC_FFY22", "EC_FFY23") ~ "Emerging Contaminants (EC)",
    fed_cap_grant %in% c("Base", "IIJA Gen Supp", "Base & IIJA Gen Supp", "Base and IIJA Gen Supp") ~ "General", 
    TRUE ~ paste0("Combined: ", fed_cap_grant)
  )) |>
  dplyr::group_by(state, state_fiscal_year, project_type) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE),
            .groups = "drop") 

```


```{r}
# Funding Data Aggregates (state level sheet)
project_type_df <- aws.s3::s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") |>
  janitor::clean_names() |>
  dplyr::filter(state == state_name) |>
  dplyr::mutate(
    state_fiscal_year = as.factor(state_fiscal_year),
    project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type),
    project_type = factor(project_type, levels=c("Both Lead and EC","Emerging Contaminants (EC)", "Lead", "General")),
    )

# Project-level data (clean full dataset)
all_projects <- aws.s3::s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") |>
  janitor::clean_names() |>
  dplyr::mutate(state_fiscal_year = as.character(state_fiscal_year))

# Funding Available (financial sheet)
all_funds <- financial |>
  dplyr::select(state, state_fiscal_year, total_funding_available) |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE))

include_y0 <- ifelse("2022" %in% unique(project_type_df$state_fiscal_year), TRUE, FALSE)
```

```{r}
# combine all years included in both sets, using characters to preserve years rather than factor levels
year_table <- as.vector(rbind(as.character(financial$state_fiscal_year), 
                               as.character(project_type_df$state_fiscal_year)))
# get rid of all duplicates
year_table <- unique(year_table)
# reset as factor
year_table <- factor(year_table)
year_table <- data.frame(state_fiscal_year=year_table)
```

# All Sources & Projects --

# DAC
## Comparing Project Cost and Funding Available

```{r}
pc_vs_fa_dac <- all_funds |>
  dplyr::left_join(
    project_type_df |>
      dplyr::select(state, state_fiscal_year, project_cost_dac) |>
      dplyr::group_by(state, state_fiscal_year) |>
      dplyr::summarise(total_project_cost_dac = sum(project_cost_dac), .groups = "drop") 
  ) |>
  dplyr::mutate(
    ratio = total_project_cost_dac/total_funding_available,
    plot_str = paste0("Project Cost: ", format_currency(total_project_cost_dac),"<br>", "Funding Available: ",format_currency(total_funding_available))
  ) 
  
pc_vs_fa_dac_p <- pc_vs_fa_dac |>
  ggplot2::ggplot(aes(state_fiscal_year, ratio, text=plot_str)) +
  ggplot2::geom_point(color = "#4ea324", size = 3, show.legend = FALSE) +
  ggplot2::geom_hline(yintercept = 1) +
  ggplot2::labs(x="State Fiscal Year",
      y="",
      title="What's the ratio of project cost to total available funding for DAC projects?",
      ) +
  ggplot2::scale_y_continuous(breaks = function(y) seq(0, ceiling(max(y)), by = 1)) +
  epic_chart_theme

pc_vs_fa_dac_gp <- ggplotly(pc_vs_fa_dac_p, tooltip="plot_str") |>
    layout(title = list(text = paste0("What's the ratio of project cost to total available funding for DAC projects?",
                                      '<br>',
                                      get_subtitle_str(pc_vs_fa_dac$state_fiscal_year, state_name))))

if (save_plots) {
    run_save_plots(gg_plot = pc_vs_fa_dac_p,
                   gp_object = pc_vs_fa_dac_gp, 
                   name = "dac-project-cost-vs-available-funding-yoy")
    }
  
pc_vs_fa_dac_gp
```

# Demand for funds project type breakdown (Updated)

```{r}

## Separate by bar
# funding amount
funding_amount_df <- all_projects |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding,
    funding_amount,
    requested_amount,
    project_cost
  ) |>
   dplyr::mutate(
    # convert $ data to numeric (No Information --> NA)
    dplyr::across(c(funding_amount, requested_amount, project_cost), as.numeric),
    # Make any $0 NA for coalesce operation
    dplyr::across(c(funding_amount, requested_amount, project_cost), \(x) ifelse(x == 0, NA, x)),
    requested_amount_ef = requested_amount * (expecting_funding == "Yes"),
    total_project_cost_ef = project_cost * (expecting_funding == "Yes"),
    # Choose first non NA value following hierarchy
    selected_funding_amount = dplyr::coalesce(funding_amount, requested_amount_ef, total_project_cost_ef),
    # Note funding source of selected funding amount (string check will flag any issue)
    source_funding = dplyr::case_when(
      funding_amount > 0 ~ "funding_amount", 
      requested_amount_ef > 0 ~ "requested_amount_ef", 
      total_project_cost_ef > 0 ~ "total_project_cost_ef",
      .default = "check"
    )
  )

funding_amount_df_summary <- funding_amount_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_funding) |>
  dplyr::summarise(funding_amount=sum(selected_funding_amount, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount == 0) |>
  tidyr::pivot_longer(cols = "funding_amount") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

# demand
demand_df <- all_projects |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding,
    funding_amount,
    requested_amount,
    project_cost
  ) |>
   dplyr::mutate(
    dplyr::across(c(funding_amount, requested_amount, project_cost), as.numeric),
    dplyr::across(c(funding_amount, requested_amount, project_cost), \(x) ifelse(x == 0, NA, x)),
    selected_demand = coalesce(requested_amount, project_cost, funding_amount), 
    source_demand = dplyr::case_when(
      requested_amount > 0 ~ "requested_amount", 
      project_cost > 0 ~ "project_cost", 
      funding_amount > 0 ~ "funding_amount", 
      .default = "check"
    )
  )

demand_df_summary <- demand_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_demand) |>
  dplyr::summarise(demand=sum(selected_demand, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand == 0) |>
  tidyr::pivot_longer(cols = "demand") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

# concatenate summary dfs
# summary all -----
funding_amount_by_source <- funding_amount_df_summary |>
  dplyr::group_by(state_fiscal_year, source_funding) |>
  dplyr::summarise(`On fundable list` = sum(funding_amount)) |>
  dplyr::ungroup()

demand_by_source <- demand_df_summary |>
  dplyr::group_by(state_fiscal_year, source_demand) |>
  dplyr::summarise(Demand = sum(demand)) |>
  dplyr::ungroup()

# Combine into one data frame with a category column
combined_bars <- dplyr::bind_rows(
  all_funds |> 
    dplyr::transmute(state_fiscal_year, value = total_funding_available, category = "Available Funds", alpha_var = NA_character_),
  funding_amount_by_source |> 
    dplyr::transmute(state_fiscal_year, value = `On fundable list`, category = "On Fundable List", alpha_var = source_funding),
  demand_by_source |> 
    dplyr::transmute(state_fiscal_year, value = Demand, category = "Demand for Funds", alpha_var = source_demand)
)

# Stacked layers (the ones with alpha)
stacked_bars <- dplyr::bind_rows(
  funding_amount_by_source |>
    dplyr::transmute(state_fiscal_year, value = `On fundable list`, category = "On Fundable List", stack_var = source_funding),
  demand_by_source |>
    dplyr::transmute(state_fiscal_year, value = Demand, category = "Demand for Funds", stack_var = source_demand)
)

# Pre-compute stacked positions within each category
stacked_positioned <- stacked_bars |>
  dplyr::group_by(state_fiscal_year, category) |>
  dplyr::mutate(
    ymax = cumsum(value),
    ymin = ymax - value,
    stack_var = dplyr::case_when(
      stack_var == "requested_amount" ~ "Requested Amount",
      stack_var == "project_cost" ~ "Project Cost",
      stack_var == "funding_amount" ~ "Funding Amount", 
      stack_var == "requested_amount_ef" ~ "Requested Amount", 
      stack_var == "total_project_cost_ef" ~ "Project Cost",
      .default = "check"
    )
  ) |>
  dplyr::ungroup() |>
dplyr::mutate(
  state_fiscal_year = as.factor(
    state_fiscal_year
  ),
  plt_str = paste0(category, ": ", scales::label_currency()(value), " (", stack_var, ")")
) 

plot_tile_demand = "Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List"

demand_p <- ggplot2::ggplot() +
  # Available Funds — left
  ggplot2::geom_rect(
    data = all_funds |>
      dplyr::ungroup() |>
      dplyr::mutate(
        category = "Available Funds",
        plt_str = paste0(category, ": ", scales::label_currency()(total_funding_available))
      ),
    ggplot2::aes(
      xmin = as.numeric(state_fiscal_year) - 0.275 - 0.125,
      xmax = as.numeric(state_fiscal_year) - 0.275 + 0.125,
      ymin = 0,
      ymax = total_funding_available,
      fill = category,
      text = plt_str
    )
  )  +
  # On Fundable List — center
  ggplot2::geom_rect(
    data = dplyr::filter(stacked_positioned, category == "On Fundable List"),
    ggplot2::aes(
      xmin = as.numeric(state_fiscal_year) - 0.125,
      xmax = as.numeric(state_fiscal_year) + 0.125,
      ymin = ymin, ymax = ymax,
      fill = category, alpha = stack_var, text = plt_str
    )
  ) +
  # Demand for Funds — right
  ggplot2::geom_rect(
    data = dplyr::filter(stacked_positioned, category == "Demand for Funds"),
    ggplot2::aes(
      xmin = as.numeric(state_fiscal_year) + 0.275 - 0.125,
      xmax = as.numeric(state_fiscal_year) + 0.275 + 0.125,
      ymin = ymin, ymax = ymax,
      fill = category, alpha = stack_var, text = plt_str
    )
  ) +
  ggplot2::scale_fill_manual(values = c(
    "Available Funds"              = "#4ea324",
    "Demand for Funds"             = "#172f60",
    "On Fundable List"             = "#791a7b",
    "Demand for Funds (Lead and EC)" = "#B15712",
    "On Fundable List (Lead and EC)" = "#ff6f00ff"
  )) +
  ggplot2::scale_alpha_manual(values = c(
    "Funding Amount" = 1,
    "Project Cost" = 1,
    "Requested Amount" = 0.8
  ), ) +
  ggplot2::scale_x_continuous(
    breaks = seq_along(levels(all_funds$state_fiscal_year)),
    labels = levels(all_funds$state_fiscal_year)
    ) +  
  ggplot2::scale_y_continuous(labels = label_currency()) + 
  ggplot2::labs(
    x = "",
    y = "",
    fill = "",
    title = plot_tile_demand,
    subtitle = get_subtitle_str(stacked_positioned$state_fiscal_year, state_name)
    ) +
      guides(alpha = "none") +
  epic_chart_theme


demand_gp <- ggplotly(demand_p, tooltip="text") |>
    layout(title = list(text = paste0(plot_tile_demand,
                                      '<br>',
                                      get_subtitle_str(stacked_positioned$state_fiscal_year, state_name)
                                      )))
  
  if (save_plots) {
    run_save_plots(gg_plot=demand_p,
                   gp_object=demand_gp,
                   name="prelim-compare-unmet-demand-available-funds-expected-funds-yoy")
  }
  
  demand_gp

```

# Demand for funds project type breakdown (by type UPDATED)

```{r}
# Combine into one data frame with a category column
combined_project_type <- dplyr::bind_rows(
  financial_by_project_type |> 
    dplyr::transmute(state_fiscal_year, project_type,  value = total_funding_available, category = "Available Funds", alpha_var = NA_character_),
  funding_amount_df_summary |> 
    dplyr::rename(`On fundable list` = funding_amount) |>
    dplyr::transmute(state_fiscal_year, project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type),  value = `On fundable list`, category = "On Fundable List", alpha_var = source_funding),
  demand_df_summary |> 
    dplyr::rename(`Demand for Funds` = demand) |>
    dplyr::transmute(state_fiscal_year, project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type),  value = `Demand for Funds`, category = "Demand for Funds", alpha_var = source_demand)
  ) |>
  dplyr::filter(
      !value == 0
    ) |>
  dplyr::mutate(
    project_type = forcats::fct(project_type)
  ) 

for (type in levels(combined_project_type$project_type)){
  
  stacked_bars_type <- dplyr::bind_rows(
    funding_amount_df_summary |> 
      dplyr::rename(`On fundable list` = funding_amount) |>
      dplyr::transmute(
        state_fiscal_year, 
        project_type = dplyr::case_when(
            project_type == "Emerging Contaminants" ~ "Emerging Contaminants (EC)",
            .default = project_type
        ),  
        value = `On fundable list`, 
        category = "On Fundable List", 
        alpha_var = source_funding, 
        stack_var = source_funding
        ),
    demand_df_summary |> 
      dplyr::rename(`Demand for Funds` = demand) |>
      dplyr::transmute(
        state_fiscal_year, 
          project_type = dplyr::case_when(
            project_type == "Emerging Contaminants" ~ "Emerging Contaminants (EC)",
            .default = project_type
        ),  
        value = `Demand for Funds`, 
        category = "Demand for Funds", 
        alpha_var = source_demand, 
        stack_var = source_demand)
  ) |>
    dplyr::filter(
      project_type == type
    )

# Pre-compute stacked positions within each category
  stacked_positioned_type <- stacked_bars_type |>
    dplyr::group_by(state_fiscal_year, category) |>
    dplyr::mutate(
      ymax = cumsum(value),
      ymin = ymax - value,
      stack_var = dplyr::case_when(
        stack_var == "requested_amount" ~ "Requested Amount",
        stack_var == "project_cost" ~ "Project Cost",
        stack_var == "funding_amount" ~ "Funding Amount", 
        stack_var == "requested_amount_ef" ~ "Requested Amount", 
        stack_var == "total_project_cost_ef" ~ "Project Cost",
        .default = "check"
      )
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      state_fiscal_year = as.factor(
        state_fiscal_year
      ),
      plt_str = paste0(category, ": ", scales::label_currency()(value), " (", stack_var, ")")
    ) 

  plot_tile_demand_type = paste0('Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List: ', type)

  avail_funds_layer <- if (nrow(financial_by_project_type|> dplyr::ungroup() |> dplyr::filter(project_type == type)) > 0) {
    ggplot2::geom_rect(
      data = financial_by_project_type |>
          dplyr::ungroup() |>
          dplyr::filter(project_type == type) |>
          dplyr::mutate(
            state_fiscal_year = factor(state_fiscal_year, levels = levels(stacked_positioned_type$state_fiscal_year)),  
            category = "Available Funds",
            plt_str = paste0(category, ": ", scales::label_currency()(total_funding_available))
          ),
      ggplot2::aes(
          xmin = as.numeric(state_fiscal_year) - 0.275 - 0.125,
          xmax = as.numeric(state_fiscal_year) - 0.275 + 0.125,
          ymin = 0,
          ymax = total_funding_available,
          fill = category,
          text = plt_str
        )
    )
  } else {
    ggplot2::geom_blank()
  }




  demand_type_p <- ggplot2::ggplot() +
    # Available Funds — left
    # ggplot2::geom_rect(
    #   data = financial_by_project_type |>
    #     dplyr::ungroup() |>
    #     dplyr::filter(project_type == type) |>
    #     dplyr::mutate(
    #       state_fiscal_year = factor(state_fiscal_year, levels = levels(stacked_positioned_type$state_fiscal_year)),  
    #       category = "Available Funds",
    #       plt_str = paste0(category, ": ", scales::label_currency()(total_funding_available))
    #     ),
    #   ggplot2::aes(
    #     xmin = as.numeric(state_fiscal_year) - 0.275 - 0.125,
    #     xmax = as.numeric(state_fiscal_year) - 0.275 + 0.125,
    #     ymin = 0,
    #     ymax = total_funding_available,
    #     fill = category,
    #     text = plt_str
    #   )
    # )  +
    avail_funds_layer +
    # On Fundable List — center
    ggplot2::geom_rect(
      data = dplyr::filter(stacked_positioned_type, category == "On Fundable List" & project_type == type),
      ggplot2::aes(
        xmin = as.numeric(state_fiscal_year) - 0.125,
        xmax = as.numeric(state_fiscal_year) + 0.125,
        ymin = ymin, ymax = ymax,
        fill = category, alpha = stack_var, text = plt_str
      )
    ) +
    # Demand for Funds — right
    ggplot2::geom_rect(
      data = dplyr::filter(stacked_positioned_type, category == "Demand for Funds" & project_type == type),
      ggplot2::aes(
        xmin = as.numeric(state_fiscal_year) + 0.275 - 0.125,
        xmax = as.numeric(state_fiscal_year) + 0.275 + 0.125,
        ymin = ymin, ymax = ymax,
        fill = category, alpha = stack_var, text = plt_str
      )
    ) +
    ggplot2::scale_fill_manual(values = c(
      "Available Funds"              = "#4ea324",
      "Demand for Funds"             = "#172f60",
      "On Fundable List"             = "#791a7b",
      "Demand for Funds (Lead and EC)" = "#B15712",
      "On Fundable List (Lead and EC)" = "#ff6f00ff"
    )) +
    ggplot2::scale_alpha_manual(values = c(
      "Funding Amount" = 1,
      "Project Cost" = 1,
      "Requested Amount" = 0.8
    ), ) +
    ggplot2::scale_x_continuous(
      breaks = seq_along(levels(stacked_positioned_type$state_fiscal_year)),
      labels = levels(stacked_positioned_type$state_fiscal_year)
      ) +  
    ggplot2::scale_y_continuous(labels = label_currency()) + 
    ggplot2::labs(
      x = "",
      y = "",
      fill = "",
      title = plot_tile_demand_type,
      subtitle = get_subtitle_str(stacked_positioned_type$state_fiscal_year, state_name)
      ) +
        guides(alpha = "none") +
    epic_chart_theme

  demand_type_gp <- ggplotly(demand_type_p, tooltip="text") |>
      layout(title = list(text = paste0(plot_tile_demand_type,
                                        '<br>',
                                        get_subtitle_str(stacked_positioned_type$state_fiscal_year, state_name)
                                        )))
    
  if (save_plots) {
    run_save_plots(
      gg_plot=demand_type_p,
      gp_object=demand_type_gp,
      name = paste0("prelim-",
        gsub("_", "-", janitor::make_clean_names(type)),
        "-compare-unmet-demand-available-funds-expected-funds-yoy"
        )
    )
  }
}

```

# Unmet demand: requested amount or project cost compared to available funds (UPDATED)

```{r}

 unmet_demand_df <- demand_by_source |>
  dplyr::group_by(state_fiscal_year) |>
  dplyr::summarise(
    total_demand = sum(Demand),
    plt_str = paste0(
      paste0(
        dplyr::recode(source_demand,
          "requested_amount" = "Demand (Requested Amount)",
          "project_cost"     = "Demand (Project Cost)"
        ),
        ": ",
        scales::label_currency()(Demand),
        collapse = "\n"
      )),
        .groups = "drop"
  ) |>
  dplyr::left_join(all_funds, by = "state_fiscal_year" )|>
  dplyr::mutate(
    `Unmet Demand` = dplyr::if_else(
      is.na(total_funding_available),
      total_demand,
      total_demand - total_funding_available
    )
    ) |>  
  dplyr::rename(`Available Funds` = total_funding_available) |>
  tidyr::pivot_longer(
    cols = c(`Unmet Demand` ,`Available Funds` )
  ) |>
  dplyr::mutate(
    plt_str = dplyr::case_when(
      name == "Available Funds" ~ paste0("Available Funds: ", scales::label_currency()(value)),
      name == "Unmet Demand" ~ paste0("Unmet Demand: ", scales::label_currency()(value), " \n", plt_str), 
      .default = "check"
    ),
    name = forcats::fct(name),
    name = forcats::fct_relevel(name, c( "Unmet Demand", "Available Funds"))
    )
    

if(sum(unmet_demand_df$value[unmet_demand_df$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  unmet_demand_p <- ggplot(unmet_demand_df, aes(x=state_fiscal_year, y=value, fill=name, text=plt_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds", 
         subtitle=get_subtitle_str(unmet_demand_df$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  unmet_demand_gp <- ggplotly(unmet_demand_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds',
                                      '<br>',
                                      get_subtitle_str(unmet_demand_df$state_fiscal_year, state_name))))
  
  if (save_plots) {
    run_save_plots(gg_plot = unmet_demand_p,
                   gp_object = unmet_demand_gp,
                   name = "prelim-compare-requested-amount-or-project-cost-available-funds-all-yoy") 
    }
  
  unmet_demand_gp
}
```

# Unmet demand: requested amount or project cost compared to available funds (by type)
```{r}
unmet_demand_by_type_df <- demand_df_summary |>
  dplyr::group_by(state_fiscal_year, project_type) |>
  dplyr::summarise(
    total_demand = sum(demand),
    plt_str = paste0(
      paste0(
        dplyr::recode(source_demand,
          "requested_amount" = "Demand (Requested Amount)",
          "project_cost"     = "Demand (Project Cost)"
        ),
        ": ",
        scales::label_currency()(demand),
        collapse = "\n"
      )),
        .groups = "drop"
  ) |>
    dplyr::mutate(
      project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type)
    ) |>
    dplyr::left_join(financial_by_project_type, by = c("state_fiscal_year", "project_type")) |>
  dplyr::mutate(
    `Unmet Demand` = dplyr::if_else(
      is.na(total_funding_available),
      total_demand,
      total_demand - total_funding_available
    )
    )|>
  dplyr::rename(`Available Funds` = total_funding_available) |>
  tidyr::pivot_longer(
    cols = c(`Unmet Demand` ,`Available Funds` )
  ) |>
  dplyr::mutate(
    plt_str = dplyr::case_when(
      name == "Available Funds" ~ paste0("Available Funds: ", scales::label_currency()(value)),
      name == "Unmet Demand" ~ paste0("Unmet Demand: ", scales::label_currency()(value), " \n", plt_str), 
      .default = "check"
    )) |>
  dplyr::mutate(
    project_type = forcats::fct(project_type)
    )

for (type in levels(unmet_demand_by_type_df$project_type)) {

  # Base filter: include Both Lead and EC rows for Lead / EC types
  filtered_data <- unmet_demand_by_type_df |>
    dplyr::filter(
      project_type == type
      # |
      # (type %in% c("Lead", "Emerging Contaminants (EC)") & project_type == "Both Lead and EC")
    )

  # # Drop 2022 for Lead / EC
  # if (type %in% c("Lead", "Emerging Contaminants (EC)")) {
  #   filtered_data <- filtered_data |>
  #     dplyr::filter(state_fiscal_year != "2022") |>
  #     dplyr::mutate(
  #       state_fiscal_year = as.character(state_fiscal_year),
  #       state_fiscal_year = as.factor(state_fiscal_year)
  #     )
  # }

  if (nrow(filtered_data) == 0) next

  filtered_data <- filtered_data |>
    dplyr::mutate(
    #   fill_group = dplyr::case_when(
    #     name == "Available Funds"  ~ "Available Funds",
    #     # name == "Unmet Demand" & project_type == "Both Lead and EC" ~ "Unmet Demand (Lead and EC)",
    #     name == "Unmet Demand" ~ "Unmet Demand"
    #   ),
      # Legend order
      fill_group = name,
        # factor(name, levels = c("Unmet Demand (Lead and EC)", "Unmet Demand", "Available Funds")),
      # Separate stack order column
      stack_order = dplyr::case_when(
        fill_group == "Available Funds"            ~ 1,
        fill_group == "Unmet Demand"               ~ 2
        # ,
        # fill_group == "Unmet Demand (Lead and EC)" ~ 3
      )
    ) |>
    dplyr::arrange(state_fiscal_year, stack_order) |>
    dplyr::group_by(state_fiscal_year) |>
    dplyr::mutate(
      ymax = cumsum(tidyr::replace_na(value, 0)),
      ymin = ymax - tidyr::replace_na(value, 0)
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      state_fiscal_year = as_factor(state_fiscal_year)
    )

  comp_unmet_avail_fund_type_p <- ggplot2::ggplot(filtered_data) +
    ggplot2::geom_rect(ggplot2::aes(
      xmin  = as.numeric(state_fiscal_year) - 0.4,
      xmax  = as.numeric(state_fiscal_year) + 0.4,
      ymin  = ymin,
      ymax  = ymax,
      fill  = fill_group,
      text  = plt_str,
      group = fill_group
    )) +
    ggplot2::scale_y_continuous(labels = label_currency()) +
    ggplot2::scale_x_continuous(
      breaks = seq_along(levels(filtered_data$state_fiscal_year)),
      labels = levels(filtered_data$state_fiscal_year)
    ) +
    ggplot2::scale_fill_manual(values = comp_colors, name = "") +
    labs(
      x        = "",
      y        = "",
      title    = paste0("Comparing Demand to Available Funds: ", type),
      subtitle = get_subtitle_str(filtered_data$state_fiscal_year, state_name)
    ) +
    epic_chart_theme +
    theme(legend.position = "bottom")

  comp_unmet_avail_fund_type_gp <- plotly::ggplotly(comp_unmet_avail_fund_type_p, tooltip = "text") |>
    plotly::layout(title = list(text = paste0(
      'Comparing Demand to Available Funds: ', type,
      '<br>',
      get_subtitle_str(filtered_data$state_fiscal_year, state_name)
    )))

  if (save_plots) {
    run_save_plots(
      gg_plot   = comp_unmet_avail_fund_type_p,
      gp_object = comp_unmet_avail_fund_type_gp,
      name = paste0("prelim-",
        gsub("_", "-", janitor::make_clean_names(type)),
        "-compare-unmet-demand-available-funds-yoy"
      )
    )
  }
  }



```

```{r}
```

