
# Combined Funding + Projects

This notebook combines data used in the `funding-data-yoy` and `project-data-yoy` notebooks and combines them for data viz and analysis where the two overlap. These include comparing the total funding available to requested amounts or project costs to consider oversubscription and other considerations.

# Setup

## Local Variables

```{r}
# filters data and changes plot names
state_name <- "Maryland"
state_abbr <- "MD"

# export plots to AWS and Drive?
save_plots <- FALSE

```

## Imports & Settings

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))

```

## Create Data Subsets

```{r, warning=FALSE}

financial <- get_financial(state_name)

base <- financial %>%
  filter(fed_cap_grant == "Base")

gen_supp <- financial %>%
  filter(fed_cap_grant == "IIJA Gen Supp")

base_gen_supp <- financial %>%
  filter(fed_cap_grant == "Base & IIJA Gen Supp")

base_gen_supp <- bind_rows(base, gen_supp, base_gen_supp) %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE))

lead <- financial %>%
  filter(fed_cap_grant %in%  c("LSLR", "LSLR_FFY22","LSLR_FFY23"))

ec <- financial %>%
  filter(fed_cap_grant %in% c("EC","EC_FFY22", "EC_FFY23"))

```


```{r}

project_type_df <- aws.s3::s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") |>
  janitor::clean_names() |>
  dplyr::filter(state == state_name) |>
  dplyr::mutate(
    state_fiscal_year = as.factor(state_fiscal_year),
    project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type),
    project_type = factor(project_type, levels=c("Both Lead and EC","Emerging Contaminants (EC)", "Lead", "General")),
    )

include_y0 <- ifelse("2022" %in% unique(project_type_df$state_fiscal_year), TRUE, FALSE)
```

```{r}
# combine all years included in both sets, using characters to preserve years rather than factor levels
year_table <- as.vector(rbind(as.character(financial$state_fiscal_year), 
                               as.character(project_type_df$state_fiscal_year)))
# get rid of all duplicates
year_table <- unique(year_table)
# reset as factor
year_table <- factor(year_table)
year_table <- data.frame(state_fiscal_year=year_table)
```

# All Sources & Projects

# Comparing Available, Costs, and Funding

```{r}  

## Comparing Project Costs & Funds Available
all_funds <- financial |>
  dplyr::select(state, state_fiscal_year, total_funding_available) |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE))

all_funds_to_compare <- project_type_df |>
  dplyr::select(state, state_fiscal_year, total_project_cost, total_project_cost_ef, total_requested_amount, total_requested_amount_ef, total_funding_amount) |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarize(
    total_project_cost = sum(total_project_cost, na.rm = TRUE),
    total_project_cost_ef = sum(total_project_cost_ef, na.rm = TRUE), #2025/11/03, per Lauren's request, if the total funding amount value is 0, use this value
    total_requested_amount = sum(total_requested_amount, na.rm = TRUE),
    total_requested_amount_ef = sum(total_requested_amount_ef, na.rm = TRUE), #2025/11/03 check in with Danielle
    total_funding_amount = sum(total_funding_amount, na.rm=TRUE)) |>
  dplyr::left_join(all_funds, by=c("state","state_fiscal_year")) 
  
compare_demand_available_funding <- all_funds_to_compare |>
  dplyr::ungroup() |>
  dplyr::mutate(
    # Depicted as On Fundable List
    total_funding_amount = dplyr::case_when(
      total_funding_amount > 0 ~ total_funding_amount,
      total_requested_amount_ef > 0 ~ total_requested_amount_ef,
      total_project_cost_ef > 0 ~ total_project_cost_ef,
      .default = 0
    ),
    funding_rule = dplyr::case_when(
      total_funding_amount > 0 ~ "* Funding Amount",
      total_requested_amount_ef > 0 ~ "** Requested Amount on Fundable List",
      total_project_cost_ef > 0 ~ "*** Project Cost on Fundable List",
      TRUE ~ "Default"
    ),
    # Depicted as Demand for Funds
    total_demand = dplyr::case_when(
      total_requested_amount > 0 ~ total_requested_amount,
      total_project_cost > 0 ~ total_project_cost,
      .default = 0
    ),
    demand_rule = dplyr::case_when(
     total_requested_amount > 0 ~ "^ Requested Amount",
     total_project_cost > 0 ~ "^^ Project Cost",
     TRUE ~ "Default"
    )
    )|>
  dplyr::select(-total_project_cost, -total_project_cost_ef, -total_requested_amount, -total_requested_amount_ef) |>   
  tidyr::pivot_longer(cols=c(total_demand, total_funding_available, total_funding_amount)) |> 
  dplyr::mutate(
    name =case_when(
      name == "total_funding_available" ~ "Available Funds",
      name == "total_funding_amount" ~ "On Fundable List",
      name == "total_demand" ~ "Demand for Funds"
      ),
    name = factor(name, levels=c("Available Funds", "On Fundable List", "Demand for Funds")),
    rule_label = dplyr::case_when(
      name == "On Fundable List" ~ funding_rule,
      name == "Demand for Funds" ~ demand_rule,
      TRUE ~ NA_character_
    ),
    plot_str = paste0(
      name, " ", 
      ifelse(!is.na(rule_label), 
      paste0("(", stringr::str_remove(rule_label, "^[\\*\\^]+\\s*"), "): "), ": "),
      format_currency(value),
      "<br>", 
      state_fiscal_year    
    ),
    rule_determinant = ifelse(is.na(rule_label), "", stringr::str_extract(rule_label, "^[\\*\\^]+")
    )
    )

dodge_width <- 0.9
if (sum(compare_demand_available_funding$value, na.rm = TRUE) > 0) {  
  comp_dem_avail_fund_p <- ggplot(compare_demand_available_funding, aes(x=state_fiscal_year, y=value, fill=name, 
  text=plot_str)) + 
    geom_bar(stat = "identity", position = position_dodge(width = dodge_width)) + 
    scale_y_continuous(labels=label_currency()) + 
    ggplot2::scale_fill_manual(values = 
    c(
      "Available Funds" = "#4ea324",
      "Demand for Funds" = "#172f60",
      "On Fundable List" = "#791a7b"
    ), name = "") + 
    labs(
      x="", 
      y="",
      title="Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List", 
      subtitle=get_subtitle_str(compare_demand_available_funding$state_fiscal_year, state_name)
      ) + 
    epic_chart_theme + 
    theme(legend.position="bottom")
  
  comp_dem_avail_fund_gp <- ggplotly(comp_dem_avail_fund_p, tooltip="text") |>
    layout(title = list(text = paste0('Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List',
                                      '<br>',
                                      get_subtitle_str(compare_demand_available_funding$state_fiscal_year, state_name)
                                      )))
  
  
  if (save_plots) {
    run_save_plots(gg_plot=comp_dem_avail_fund_p,
                   gp_object=comp_dem_avail_fund_gp,
                   name="compare-unmet-demand-available-funds-expected-funds-yoy")
  }
  
  comp_dem_avail_fund_gp
}
```

### Further break down by project type
```{r}
# get total funding available for each fed cap grant, merge with project type
financial_by_project_type <- financial |>
  dplyr::mutate(project_type = case_when(
    fed_cap_grant %in% c("LSLR", "LSLR_FFY22","LSLR_FFY23") ~ "Lead",
    fed_cap_grant %in% c("EC", "EC_FFY22", "EC_FFY23") ~ "Emerging Contaminants (EC)",
    fed_cap_grant %in% c("Base", "IIJA Gen Supp", "Base & IIJA Gen Supp") ~ "General", 
    TRUE ~ paste0("Combined: ", fed_cap_grant)
  )) |>
  dplyr::group_by(state, state_fiscal_year, project_type) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE),
            .groups = "drop") |>
  dplyr::filter(total_funding_available > 0)            

all_funds_to_compare_type <- project_type_df |>
  dplyr::select(state, state_fiscal_year, project_type, total_project_cost, total_project_cost_ef, total_requested_amount, total_requested_amount_ef, total_funding_amount) |>
  dplyr::group_by(state, state_fiscal_year, project_type) |>
  dplyr::summarize(
    total_project_cost = sum(total_project_cost, na.rm = TRUE),
    total_project_cost_ef = sum(total_project_cost_ef, na.rm = TRUE), #2025/11/03, per Lauren's request, if the total funding amount value is 0, use this value
    total_requested_amount = sum(total_requested_amount, na.rm = TRUE),
    total_requested_amount_ef = sum(total_requested_amount_ef, na.rm = TRUE), #2025/11/03 check in with Danielle
    total_funding_amount = sum(total_funding_amount, na.rm=TRUE)) |>
  dplyr::left_join(financial_by_project_type, by=c("state","state_fiscal_year", "project_type")) 
  
compare_demand_available_funding_type <- all_funds_to_compare_type |>
  dplyr::ungroup() |>
  dplyr::mutate(
    # Depicted as On Fundable List
    total_funding_amount = dplyr::case_when(
      total_funding_amount > 0 ~ total_funding_amount,
      total_requested_amount_ef > 0 ~ total_requested_amount_ef,
      total_project_cost_ef > 0 ~ total_project_cost_ef,
      .default = 0
    ),
    funding_rule = dplyr::case_when(
      total_funding_amount > 0 ~ "* Funding Amount",
      total_requested_amount_ef > 0 ~ "** Requested Amount on Fundable List",
      total_project_cost_ef > 0 ~ "*** Project Cost on Fundable List",
      TRUE ~ "Default"
    ),
    # Depicted as Demand for Funds
    total_demand = dplyr::case_when(
      total_requested_amount > 0 ~ total_requested_amount,
      total_project_cost > 0 ~ total_project_cost,
      .default = 0
    ),
    demand_rule = dplyr::case_when(
     total_requested_amount > 0 ~ "^ Requested Amount",
     total_project_cost > 0 ~ "^^ Project Cost",
     TRUE ~ "Default"
    )
    )|>
  dplyr::select(-total_project_cost, -total_project_cost_ef, -total_requested_amount, -total_requested_amount_ef) |>   
  tidyr::pivot_longer(cols=c(total_demand, total_funding_available, total_funding_amount)) |> 
  dplyr::mutate(
    name =case_when(
      name == "total_funding_available" ~ "Available Funds",
      name == "total_funding_amount" ~ "On Fundable List",
      name == "total_demand" ~ "Demand for Funds"
      ),
    name = factor(name, levels=c("Available Funds", "On Fundable List", "Demand for Funds")),
    rule_label = dplyr::case_when(
      name == "On Fundable List" ~ funding_rule,
      name == "Demand for Funds" ~ demand_rule,
      TRUE ~ NA_character_
    ),
    plot_str = paste0(
      name, " ", 
      ifelse(!is.na(rule_label), 
      paste0("(", stringr::str_remove(rule_label, "^[\\*\\^]+\\s*"), "): "), ": "),
      format_currency(value),
      "<br>", 
      state_fiscal_year    
    ),
    rule_determinant = ifelse(is.na(rule_label), "", stringr::str_extract(rule_label, "^[\\*\\^]+")),
    project_type = as.factor(project_type)
    )


for (type in levels(compare_demand_available_funding_type$project_type)){

# Define bar width and dodge width constants
bar_width  <- 0.25
dodge_width <- 0.8

# Assign each bar_group a dodge x-offset
dodge_offsets <- c(
    "Available Funds"  = -dodge_width / 3,
    "On Fundable List" =  0,
    "Demand for Funds" =  dodge_width / 3
)

filtered_data <- compare_demand_available_funding_type |>
    dplyr::filter(
        project_type == type |
        (type %in% c("Lead", "Emerging Contaminants (EC)") & project_type == "Both Lead and EC")
    ) |>
    dplyr::mutate(
        bar_group = as.character(name),
        fill_group = dplyr::case_when(
            bar_group == "Available Funds"  ~ "Available Funds",
            bar_group == "Demand for Funds" & project_type == "Both Lead and EC" ~ "Demand for Funds (Lead and EC)",
            bar_group == "Demand for Funds" ~ "Demand for Funds",
            bar_group == "On Fundable List" & project_type == "Both Lead and EC" ~ "On Fundable List (Lead and EC)",
            bar_group == "On Fundable List" ~ "On Fundable List"
        ),
        fill_group = factor(fill_group, levels=c("Available Funds", "On Fundable List", "On Fundable List (Lead and EC)", "Demand for Funds", "Demand for Funds (Lead and EC)")),
        x_num = as.numeric(state_fiscal_year) + dodge_offsets[bar_group]
    ) |>
    # stack within each year + bar_group: Primary first, Lead+EC on top
    dplyr::arrange(state_fiscal_year, bar_group, project_type == "Both Lead and EC") |>
    dplyr::group_by(state_fiscal_year, bar_group) |>
    dplyr::mutate(
        ymax = cumsum(tidyr::replace_na(value, 0)),
        ymin = ymax - tidyr::replace_na(value, 0)
    ) |>
    dplyr::ungroup()


  # Skip if no data for this type
  if(nrow(filtered_data) == 0) {
    next
  }

  comp_dem_avail_fund_type_p <- ggplot2::ggplot(filtered_data) +
    ggplot2::geom_rect(ggplot2::aes(
        xmin  = x_num - bar_width / 2,
        xmax  = x_num + bar_width / 2,
        ymin  = ymin,
        ymax  = ymax,
        fill  = fill_group,
        text  = plot_str,
        group = fill_group
    )) +
    ggplot2::scale_y_continuous(labels = label_currency()) +
    ggplot2::scale_x_continuous(
        breaks = seq_along(levels(filtered_data$state_fiscal_year)),
        labels = levels(filtered_data$state_fiscal_year)
    ) +
    ggplot2::scale_fill_manual(values = c(
        "Available Funds"              = "#4ea324",
        "Demand for Funds"             = "#172f60",
        "On Fundable List"             = "#791a7b",
        "Demand for Funds (Lead and EC)" = "#B15712",
        "On Fundable List (Lead and EC)" = "#ff6f00ff"
    ), name = "") +
    labs(
        x        = "",
        y        = "",
        title    = paste0("Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List: ", type),
        subtitle = get_subtitle_str(compare_demand_available_funding_type$state_fiscal_year, state_name)
    ) +
    epic_chart_theme +
    theme(legend.position = "bottom")
  
  comp_dem_avail_fund_type_gp <- plotly::ggplotly(comp_dem_avail_fund_type_p, tooltip="text") |>
    plotly::layout(title = list(text = paste0('Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List: ', type,
                                      '<br>',
                                      get_subtitle_str(compare_demand_available_funding_type$state_fiscal_year, state_name)
                                      ))
           )

  if (save_plots) {
    run_save_plots(
      gg_plot=comp_dem_avail_fund_type_p,
      gp_object=comp_dem_avail_fund_type_gp,
      name = paste0(
        gsub("_", "-", janitor::make_clean_names(type)),
        "-compare-unmet-demand-available-funds-expected-funds-yoy"
        )
        )
  }

  comp_dem_avail_fund_type_gp

}
```

# Unmet demand: requested amount or project cost compared to available funds

```{r}

unmet_demand <- all_funds_to_compare |>
  dplyr::ungroup() |>
  dplyr::mutate(
    # Depicted as Demand for Funds
    total_demand = dplyr::case_when(
      total_requested_amount > 0 ~ total_requested_amount,
      total_project_cost > 0 ~ total_project_cost,
      .default = 0
    ),
    demand_rule = dplyr::case_when(
     total_requested_amount > 0 ~ "^ Requested Amount",
     total_project_cost > 0 ~ "^^ Project Cost",
     TRUE ~ "Default"
    ),
    unmet_demand = total_demand - total_funding_available
    )|>
  dplyr::select(-total_project_cost, -total_project_cost_ef, -total_requested_amount, -total_requested_amount_ef, -total_funding_amount, -total_demand) |>   
  tidyr::pivot_longer(cols=c(unmet_demand, total_funding_available)) |> 
  dplyr::mutate(
    name =case_when(
      name == "total_funding_available" ~ "Available Funds",
      name == "unmet_demand" ~ "Unmet Demand"
      ),
    name = factor(name, levels=c("Unmet Demand", "Available Funds")),
    rule_label = dplyr::case_when(
      name == "Unmet Demand" ~ demand_rule,
      TRUE ~ NA_character_
    ),
    plot_str = paste0(
      name, " ", 
      ifelse(!is.na(rule_label), 
      paste0("(", stringr::str_remove(rule_label, "^[\\*\\^]+\\s*"), "): "), ": "),
      format_currency(value),
      "<br>", 
      state_fiscal_year    
    ),
    rule_determinant = ifelse(is.na(rule_label), "", stringr::str_extract(rule_label, "^[\\*\\^]+")
    )
    )


if(sum(unmet_demand$value[unmet_demand$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  all_pc_p <- ggplot(unmet_demand, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds", 
         subtitle=get_subtitle_str(unmet_demand$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  all_pc_gp <- ggplotly(all_pc_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds',
                                      '<br>',
                                      get_subtitle_str(unmet_demand$state_fiscal_year, state_name))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = all_pc_p,
                   gp_object = all_pc_gp,
                   name = "compare-requested-amount-or-project-cost-available-funds-all-yoy") 
    }
  
  all_pc_gp
}
```

# Unmet demand: requested amount or project cost compared to available funds by type (General, Lead, EC)

```{r}
unmet_demand_type <- all_funds_to_compare_type |>
  dplyr::ungroup() |>
  dplyr::mutate(
    total_demand = dplyr::case_when(
      total_requested_amount > 0 ~ total_requested_amount,
      total_project_cost > 0 ~ total_project_cost,
      .default = 0
    ),
    demand_rule = dplyr::case_when(
     total_requested_amount > 0 ~ "^ Requested Amount",
     total_project_cost > 0 ~ "^^ Project Cost",
     TRUE ~ "Default"
    ),
    unmet_demand = total_demand - tidyr::replace_na(total_funding_available, 0)
  ) |>
  dplyr::select(-total_project_cost, -total_project_cost_ef, -total_requested_amount, -total_requested_amount_ef, -total_funding_amount, -total_demand) |>   
  tidyr::pivot_longer(cols = c(unmet_demand, total_funding_available)) |> 
  dplyr::mutate(
    name = case_when(
      name == "total_funding_available" ~ "Available Funds",
      name == "unmet_demand"            ~ "Unmet Demand"
    ),
    name = factor(name, levels = c("Available Funds", "Unmet Demand")),
    rule_label = dplyr::case_when(
      name == "Unmet Demand" ~ demand_rule,
      TRUE ~ NA_character_
    ),
    plot_str = paste0(
      name, " ", 
      ifelse(!is.na(rule_label), 
             paste0("(", stringr::str_remove(rule_label, "^[\\*\\^]+\\s*"), "): "), ": "),
      format_currency(value),
      "<br>", 
      state_fiscal_year    
    ),
    rule_determinant = ifelse(is.na(rule_label), "", stringr::str_extract(rule_label, "^[\\*\\^]+")),
    project_type = as.factor(project_type)
  )


for (type in levels(unmet_demand_type$project_type)) {

  # Base filter: include Both Lead and EC rows for Lead / EC types
  filtered_data <- unmet_demand_type |>
    dplyr::filter(
      project_type == type |
      (type %in% c("Lead", "Emerging Contaminants (EC)") & project_type == "Both Lead and EC")
    )

  # Drop 2022 for Lead / EC
  if (type %in% c("Lead", "Emerging Contaminants (EC)")) {
    filtered_data <- filtered_data |>
      dplyr::filter(state_fiscal_year != "2022") |>
      dplyr::mutate(
        state_fiscal_year = as.character(state_fiscal_year),
        state_fiscal_year = as.factor(state_fiscal_year)
      )
  }

  if (nrow(filtered_data) == 0) next

  filtered_data <- filtered_data |>
    dplyr::mutate(
      fill_group = dplyr::case_when(
        name == "Available Funds"  ~ "Available Funds",
        name == "Unmet Demand" & project_type == "Both Lead and EC" ~ "Unmet Demand (Lead and EC)",
        name == "Unmet Demand" ~ "Unmet Demand"
      ),
      # Legend order
      fill_group = factor(fill_group, levels = c("Unmet Demand (Lead and EC)", "Unmet Demand", "Available Funds")),
      # Separate stack order column
      stack_order = dplyr::case_when(
        fill_group == "Available Funds"            ~ 1,
        fill_group == "Unmet Demand"               ~ 2,
        fill_group == "Unmet Demand (Lead and EC)" ~ 3
      )
    ) |>
    dplyr::arrange(state_fiscal_year, stack_order) |>
    dplyr::group_by(state_fiscal_year) |>
    dplyr::mutate(
      ymax = cumsum(tidyr::replace_na(value, 0)),
      ymin = ymax - tidyr::replace_na(value, 0)
    ) |>
    dplyr::ungroup()

  comp_unmet_avail_fund_type_p <- ggplot2::ggplot(filtered_data) +
    ggplot2::geom_rect(ggplot2::aes(
      xmin  = as.numeric(state_fiscal_year) - 0.4,
      xmax  = as.numeric(state_fiscal_year) + 0.4,
      ymin  = ymin,
      ymax  = ymax,
      fill  = fill_group,
      text  = plot_str,
      group = fill_group
    )) +
    ggplot2::scale_y_continuous(labels = label_currency()) +
    ggplot2::scale_x_continuous(
      breaks = seq_along(levels(filtered_data$state_fiscal_year)),
      labels = levels(filtered_data$state_fiscal_year)
    ) +
    ggplot2::scale_fill_manual(values = comp_colors, name = "") +
    labs(
      x        = "",
      y        = "",
      title    = paste0("Comparing Demand to Available Funds: ", type),
      subtitle = get_subtitle_str(compare_demand_available_funding_type$state_fiscal_year, state_name)
    ) +
    epic_chart_theme +
    theme(legend.position = "bottom")

  comp_unmet_avail_fund_type_gp <- plotly::ggplotly(comp_unmet_avail_fund_type_p, tooltip = "text") |>
    plotly::layout(title = list(text = paste0(
      'Comparing Demand to Available Funds: ', type,
      '<br>',
      get_subtitle_str(filtered_data$state_fiscal_year, state_name)
    )))

  if (save_plots) {
    run_save_plots(
      gg_plot   = comp_unmet_avail_fund_type_p,
      gp_object = comp_unmet_avail_fund_type_gp,
      name = paste0(
        gsub("_", "-", janitor::make_clean_names(type)),
        "-compare-unmet-demand-available-funds-yoy"
      )
    )
  }

  comp_unmet_avail_fund_type_gp

}
```


# DAC
## Comparing Project Cost and Funding Available

```{r}
pc_vs_fa_dac <- all_funds |>
  dplyr::left_join(
    project_type_df |>
      dplyr::select(state, state_fiscal_year, project_cost_dac) |>
      dplyr::group_by(state, state_fiscal_year) |>
      dplyr::summarise(total_project_cost_dac = sum(project_cost_dac), .groups = "drop") 
  ) |>
  dplyr::mutate(
    ratio = total_project_cost_dac/total_funding_available,
    plot_str = paste0("Project Cost: ", format_currency(total_project_cost_dac),"<br>", "Funding Available: ",format_currency(total_funding_available))
  ) 
  
pc_vs_fa_dac_p <- pc_vs_fa_dac |>
  ggplot2::ggplot(aes(state_fiscal_year, ratio, text=plot_str)) +
  ggplot2::geom_point(color = "#4ea324", size = 3, show.legend = FALSE) +
  ggplot2::geom_hline(yintercept = 1) +
  ggplot2::labs(x="State Fiscal Year",
      y="",
      title="What's the ratio of project cost to total available funding for DAC projects?",
      ) +
  ggplot2::scale_y_continuous(breaks = function(y) seq(0, ceiling(max(y)), by = 1)) +
  epic_chart_theme

pc_vs_fa_dac_gp <- ggplotly(pc_vs_fa_dac_p, tooltip="plot_str") |>
    layout(title = list(text = paste0("What's the ratio of project cost to total available funding for DAC projects?",
                                      '<br>',
                                      get_subtitle_str(pc_vs_fa_dac$state_fiscal_year, state_name))))

if (save_plots) {
    run_save_plots(gg_plot = pc_vs_fa_dac_p,
                   gp_object = pc_vs_fa_dac_gp, 
                   name = "dac-project-cost-vs-available-funding-yoy")
    }
  
pc_vs_fa_dac_gp
```

```{r}
```