
# Combined Funding + Projects

This notebook combines data used in the `funding-data-yoy` and `project-data-yoy` notebooks and combines them for data viz and analysis where the two overlap. These include comparing the total funding available to requested amounts or project costs to consider oversubscription and other considerations.

# Setup

## Local Variables

```{r}
# filters data and changes plot names
state_name <- "Texas"
state_abbr <- "TX"

# export plots to AWS and Drive?
save_plots <- FALSE

```

## Imports & Settings

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))

```

## Create Data Subsets

```{r, warning=FALSE}

financial <- get_financial(state_name)

base <- financial %>%
  filter(fed_cap_grant == "Base")

gen_supp <- financial %>%
  filter(fed_cap_grant == "IIJA Gen Supp")

base_gen_supp <- financial %>%
  filter(fed_cap_grant == "Base & IIJA Gen Supp")

base_gen_supp <- bind_rows(base, gen_supp, base_gen_supp) %>%
  group_by(state, state_fiscal_year) %>%
  summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE))

lead <- financial %>%
  filter(fed_cap_grant == "LSLR")

ec <- financial %>%
  filter(fed_cap_grant == "EC")

```


```{r}

project_type_df <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(state_fiscal_year = as.factor(state_fiscal_year))
```

```{r}
# combine all years included in both sets, using characters to preserve years rather than factor levels
year_table <- as.vector(rbind(as.character(financial$state_fiscal_year), 
                               as.character(project_type_df$state_fiscal_year)))
# get rid of all duplicates
year_table <- unique(year_table)
# reset as factor
year_table <- factor(year_table)
year_table <- data.frame(state_fiscal_year=year_table)
```

# Base & IIJA General Supplemental

## Comparing Project Costs & Funds Available

```{r}

compare_general_funds <- project_type_df %>%
  select(state, state_fiscal_year, project_type, total_project_cost, total_requested_amount) %>%
  filter(project_type == "General") %>%
  left_join(base_gen_supp %>% select(state, state_fiscal_year, total_funding_available)) %>%   
# create calculated columns that will be "what's left" so that the total and the proportions are correct in plots
  mutate(cost_less_available = total_project_cost - replace_na(total_funding_available, 0),
         cost_less_available = ifelse(cost_less_available < 0, 0, cost_less_available),
         requested_less_available = total_requested_amount - replace_na(total_funding_available, 0),
         requested_less_available = ifelse(requested_less_available < 0, 0, requested_less_available))

totals_by_year <- compare_general_funds %>%
  group_by(state_fiscal_year) %>%
  summarize(cost_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(cost_less_available, na.rm=TRUE),
            requested_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(requested_less_available, na.rm=TRUE))

# make sure all years are present
compare_general_funds <- year_table %>%
  left_join(compare_general_funds, by="state_fiscal_year") %>%
  left_join(totals_by_year, by="state_fiscal_year")


# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_project_cost_gen <- compare_general_funds %>%
  select(state, state_fiscal_year, total_funding_available, cost_less_available, cost_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, cost_less_available)) %>%
  mutate(pct_str = as.character(round((value/cost_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "cost_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)")
    )



# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_requested_amount_gen <- compare_general_funds %>%
  select(state, state_fiscal_year, total_funding_available, requested_less_available, requested_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, requested_less_available)) %>%
  mutate(pct_str = as.character(round((value/requested_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "requested_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)"))

# make sure all years are present
compare_requested_amount_gen <- year_table %>%
  left_join(compare_requested_amount_gen, by="state_fiscal_year")

```

```{r}

if(sum(compare_project_cost_gen$value[compare_project_cost_gen$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  gen_pc_p <- ggplot(compare_project_cost_gen, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds, General", 
         subtitle=get_subtitle_str(compare_project_cost_gen$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  gen_pc_gp <- ggplotly(gen_pc_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, General',
                                      '<br>',
                                      get_subtitle_str(compare_project_cost_gen$state_fiscal_year, state_name))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = gen_pc_p,
                   gp_object = gen_pc_gp,
                   name = "compare-project-cost-available-funds-gen-yoy") 
    }
  
  gen_pc_gp
}
```


## Comparing Requested Amount & Funds Available

```{r}

if(sum(compare_requested_amount_gen$value[compare_requested_amount_gen$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  gen_ra_p <- ggplot(compare_requested_amount_gen, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds, Base & IIJA Gen Supp",
         subtitle=get_subtitle_str(compare_requested_amount_gen$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  gen_ra_gp <- ggplotly(gen_ra_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, Base & IIJA Gen Supp',
                                      '<br>',
                                      get_subtitle_str(compare_requested_amount_gen$state_fiscal_year, state_name))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = gen_ra_p,
                   gp_object = gen_ra_gp,
                   name = "compare-requested-amount-available-funds-gen-yoy")
    }
  
  gen_ra_gp
}
```

# EC

## Data Prep

```{r}

compare_ec_funds <- project_type_df %>%
  select(state, state_fiscal_year, project_type, total_project_cost, total_requested_amount) %>%
  filter(project_type == "Emerging Contaminants") %>%
  left_join(ec %>% select(state, state_fiscal_year, total_funding_available)) %>%   
# create calculated columns that will be "what's left" so that the total and the proportions are correct in plots
  mutate(cost_less_available = total_project_cost - replace_na(total_funding_available, 0),
         cost_less_available = ifelse(cost_less_available < 0, 0, cost_less_available),
         requested_less_available = total_requested_amount - replace_na(total_funding_available, 0),
         requested_less_available = ifelse(requested_less_available < 0, 0, requested_less_available))


totals_by_year_ec <- compare_ec_funds %>%
  group_by(state_fiscal_year) %>%
  summarize(cost_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(cost_less_available, na.rm=TRUE),
            requested_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(requested_less_available, na.rm=TRUE))

# make sure all years are present for plotting, remove 2022 for lead specifically
compare_ec_funds <- year_table %>%
  left_join(compare_ec_funds, by="state_fiscal_year") %>%
  left_join(totals_by_year_ec, by="state_fiscal_year") %>%
  filter(state_fiscal_year !="2022")

# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_project_cost_ec <- compare_ec_funds %>%
select(state, state_fiscal_year, total_funding_available, cost_less_available, cost_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, cost_less_available)) %>%
  mutate(pct_str = as.character(round((value/cost_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "cost_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)")
  )


# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_requested_amount_ec <- compare_ec_funds %>%
  select(state, state_fiscal_year, total_funding_available, requested_less_available, requested_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, requested_less_available)) %>%
  mutate(pct_str = as.character(round((value/requested_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "requested_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)"))


```

## Project Costs

```{r}

if(sum(compare_project_cost_ec$value[compare_project_cost_ec$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  
  ec_pc_p <- ggplot(compare_project_cost_ec, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds, EC", 
         subtitle=get_subtitle_str(compare_project_cost_ec$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  ec_pc_gp <- ggplotly(ec_pc_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, EC',
                                      '<br>',
                                      get_subtitle_str(compare_project_cost_ec$state_fiscal_year, state_name))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = ec_pc_p,
                   gp_object = ec_pc_gp, # change
                   name = "compare-project-cost-available-funds-ec-yoy")
                   }
  
  ec_pc_gp
}
```

## Requested Amount

```{r}

if(sum(compare_requested_amount_ec$value[compare_requested_amount_ec$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  
  ec_ra_p <- ggplot(compare_requested_amount_ec, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds, EC",
         subtitle=get_subtitle_str(compare_requested_amount_ec$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  ec_ra_gp <- ggplotly(ec_ra_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, EC',
                                      '<br>',
                                      get_subtitle_str(compare_requested_amount_ec$state_fiscal_year, state_name))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = ec_ra_p,
                   gp_object = ec_ra_gp,
                   name = "compare-requested-amount-available-funds-ec-yoy")
    }
  
  ec_ra_gp
}
```

# Lead

## Data Prep

```{r}

compare_lead_funds <- project_type_df %>%
  select(state, state_fiscal_year, project_type, total_project_cost, total_requested_amount) %>%
  filter(project_type == "Lead") %>%
  left_join(lead %>% select(state, state_fiscal_year, total_funding_available)) %>%   
# create calculated columns that will be "what's left" so that the total and the proportions are correct in plots
  mutate(cost_less_available = total_project_cost - replace_na(total_funding_available, 0),
         cost_less_available = ifelse(cost_less_available < 0, 0, cost_less_available),
         requested_less_available = total_requested_amount - replace_na(total_funding_available, 0),
         requested_less_available = ifelse(requested_less_available < 0, 0, requested_less_available))

totals_by_year_lead <- compare_lead_funds %>%
  group_by(state_fiscal_year) %>%
  summarize(cost_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(cost_less_available, na.rm=TRUE),
            requested_available_combined = sum(total_funding_available, na.rm=TRUE) + sum(requested_less_available, na.rm=TRUE))

# make sure all years are present for plotting, remove 2022 for lead specifically
compare_lead_funds <- year_table %>%
  left_join(compare_lead_funds, by="state_fiscal_year") %>%
  left_join(totals_by_year_lead, by="state_fiscal_year") %>%
  filter(state_fiscal_year !="2022")

# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_project_cost_lead <- compare_lead_funds %>%
select(state, state_fiscal_year, total_funding_available, cost_less_available, cost_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, cost_less_available)) %>%
  mutate(pct_str = as.character(round((value/cost_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "cost_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)")
    )

# from the simplified dataset above, generate the longer dataframe for plotting stacked
compare_requested_amount_lead <- compare_lead_funds %>%
  select(state, state_fiscal_year, total_funding_available, requested_less_available, requested_available_combined) %>%
  pivot_longer(cols=c(total_funding_available, requested_less_available)) %>%
  mutate(pct_str = as.character(round((value/requested_available_combined),3)*100),
    name = case_when(
    name == "total_funding_available" ~ "Total Available",
    name == "requested_less_available" ~ "Unmet Demand"),
    name = factor(name, levels=c("Unmet Demand", "Total Available")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value),
                      " (", pct_str, "%)"))
```

### Project Cost

```{r}
if(sum(compare_project_cost_lead$value[compare_project_cost_lead$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  lead_pc_p <- ggplot(compare_project_cost_lead, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", title="Comparing Demand to Available Funds, Lead",
         subtitle=get_subtitle_str(compare_project_cost_lead$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  lead_pc_gp <- ggplotly(lead_pc_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, Lead',
                                      '<br>',
                                      get_subtitle_str(compare_project_cost_lead$state_fiscal_year, state_name))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = lead_pc_p,
                   gp_object = lead_pc_gp,
                   name = "compare-project-cost-available-funds-lead-yoy")
    }
  
  lead_pc_gp
}
```

### Requested Amount

```{r}

if(sum(compare_requested_amount_lead$value[compare_requested_amount_lead$name=="Unmet Demand"], na.rm=TRUE) > 0) {
  
  lead_ra_p <- ggplot(compare_requested_amount_lead, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = comp_colors, 
                      name = "") + 
    labs(x="", y="", 
         title="Comparing Demand to Available Funds, Lead",
         subtitle=get_subtitle_str(compare_requested_amount_lead$state_fiscal_year, state_name)) + 
    epic_chart_theme
  
  lead_ra_gp <- ggplotly(lead_ra_p, tooltip="text") %>%
    layout(title = list(text = paste0('Comparing Demand to Available Funds, Lead',
                                      '<br>',
                                      get_subtitle_str(compare_requested_amount_lead$state_fiscal_year, state_name))))
  
  
  
  # if (save_plots) {
  #save_to_aws(lead_ra_gp, "compare-demand-available-funds-lead", state_abbr) }
  
  if (save_plots) {
    run_save_plots(gg_plot = lead_ra_p,
                   gp_object = lead_ra_gp,
                   name = "compare-requested-amount-available-funds-lead-yoy")
    }
  
  lead_ra_gp
}
```