---
title: "Finalized Agreements Analysis"
author: "Phil Cork"
date: "2024-07-08"
---

This notebook is a draft of analysis that could potentially feed into the SRF Funding Tracker for the "Finalized Agreement" sections of a particular state, using Texas as an example.

In the future, this notebook will need to be modified to either handle multiple years of data or be modified to work on only one year of data and then develop a second year-over-year notebook, similar to how the IUP data is handled. This update will be dependent on future updates to the EPA's Public SRF Portal and then repulling that data into AWS.

```{r}
state_name <- "Texas"
years <- "SFY23"
subtitle_str <- paste0(state_name, ", ", years)
```



```{r}
library(tidyverse)
library(janitor)
library(aws.s3)
library(scales)
library(lubridate)
library(sysfonts)
library(showtext)

#TODO: bringing cleaning functions into shared functions and update references
source("../../dw-dashboard/resources.R")

source("../../ds-resources/shared-functions.R")
run_code_from_file("../../local/aws.txt")

options(scipen=999)

# fonts: 
sysfonts::font_add_google("Lato")
showtext::showtext_auto()

# theme: 
# legend position is right, text size is at least 10, and text is Lato
epic_chart_theme <- theme_minimal() + 
  theme(legend.position = "right", 
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), 
        axis.text.x = element_text(margin = margin(t = 10, r = 0, 
                                                   b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, 
                                                    b = 0, l = 0)), 
        axis.text.y = element_text(margin = margin(t = 0, r = 10, 
                                                   b = 0, l = 0)), 
        axis.title.y = element_text(margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0))) 

cont_palette <- colorRampPalette(c("#172f60","#4ea324"))
```

# Funding Decisions

```{r}
all_agreements <-  s3read_using(read.csv, object="s3://water-team-data/raw_data/dwsrf_awards/epa-portal-agreements/epa-portal-sfy23.csv") %>%
  clean_names()
```

```{r}

state_finalized_agreements <- all_agreements %>%
  filter(state == state_name) %>%
  filter(latest_agreement_action == "Initial Agreement") %>%
  mutate(initial_agreement_date = mdy(initial_agreement_date)) %>%
  #TODO: Make this more modular so it'll work for different years, controlled at the top, once new years are added
  filter(initial_agreement_date >= "2022-10-01" & initial_agreement_date <= "2023-09-30") %>%
  mutate(initial_agreement_amount = convert_to_numeric(initial_agreement_amount),
         additional_subsidy_amount = convert_to_numeric(additional_subsidy_amount),
         grant_assigned_subsidy = convert_to_numeric(grant_assigned_subsidy)
         )

```


## How much money was initially agreed upon?
```{r}
print(sum(state_finalized_agreements$initial_agreement_amount))
```

## How much of the agreement amounts are additional subsidy?
```{r}
print(sum(state_finalized_agreements$additional_subsidy_amount))
```

## How many agreements are going to disadvantaged communities?
```{r}
print(nrow(state_finalized_agreements %>% filter(disadvantaged_assistance == "Yes")))
```

## How much of the initial agreement amount is going to DACs?
```{r}
state_finalized_agreements %>% 
  filter(disadvantaged_assistance == "Yes") %>% 
  summarize(total_assistance = sum(initial_agreement_amount))
```
How much of the additional subsidy amount is going to DACs?
```{r}
state_finalized_agreements %>% 
  filter(disadvantaged_assistance == "Yes") %>% 
  summarize(total_assistance = sum(additional_subsidy_amount))
```


# Subset to only BIL
Note: This initial code is borrowed from srf-sprints/epa-portal-analysis

```{r}
only_bil <- state_finalized_agreements %>%
  filter(supplemental_appropriation %in% c("2022 BIL Emerging Contaminants", "2022 BIL General Supplemental",
                                           "2022 BIL Lead Service Line Replacement" ,"2023 BIL General Supplemental")) %>%
  mutate(supplemental_appropriation = case_when(grepl("General Supplemental", supplemental_appropriation) ~ "General Supplemental",
                                                grepl("Emerging Contaminants", supplemental_appropriation) ~ "Emerging Contaminants",
                                                grepl("Lead Service", supplemental_appropriation) ~ "Lead Service Line Replacement")
    )

bil_by_source <- only_bil %>%
  group_by(supplemental_appropriation) %>%
  summarize(agreements = n(),
            total_i_agreement_amount = sum(initial_agreement_amount),
            total_additional_subsidy_amount = sum(additional_subsidy_amount))

bil_by_source
```

With no BIL funds, it can be difficult to otherwise categorize projects into General, Lead, or EC, but there are a few other columns we can check

```{r}

# are any lead lines estimated to be replaced? NOTE most states don't provide this, but worth a shot as the dataset continues to evolve
sum(state_finalized_agreements$est_lead_service_lines_replaced)


tibble(state_finalized_agreements %>%
  group_by(state) %>%
  summarize(ec_projects = sum(grepl("PFAS|Emerging Contaminants", project_description, ignore.case=TRUE)),
            lead_projects = sum(grepl("Lead|lsl|grr", project_description, ignore.case=TRUE)))
)


```

# Data Viz

```{r}

by_dac <- state_finalized_agreements %>%
  group_by(disadvantaged_assistance) %>%
  summarize(agreements = n(),
            total_init_agreement = sum(initial_agreement_amount),
            total_add_subsidy = sum(additional_subsidy_amount),
            non_subsidy = total_init_agreement - total_add_subsidy) %>%
  pivot_longer(cols=c(non_subsidy, total_add_subsidy)) %>%
  select(disadvantaged_assistance, name, value) %>%
  mutate(name = ifelse(name=="non_subsidy", "Non-Subsidy", "Subsidy"))

ggplot(by_dac, aes(x=value, y=disadvantaged_assistance, fill=name)) + 
  geom_bar(stat="identity", position="stack") + 
  labs(x="", y="Disadvantaged", fill="",
       title="How much of the finalized awards are subsidies?", subtitle = subtitle_str) + 
  scale_fill_manual(values=cont_palette(2)) +  
  #NOTE: for this code to be more modular, may consider how to automatically handle scaling and adding a suffix to numbers
  # ie when it makes more sense in 100k or 1b scale
  scale_x_continuous(labels=label_currency(scale=.000001, suffix="M")) + 
  epic_chart_theme

```

