---
title: "Funding Data Year Over Year"
author: "Phil Cork"
date: "2024-11-08"
---

This notebook presents data viz and statistics for the SRF Funding Tracker state overview pages based on Funding & Finance data gathered by the Water policy team as they put together the data dictionaries and content for the website. The outputs focus on the "Funding" sections of the state overview pages.

# Setup

## Imports & Settings

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))
```


## Local Variables

```{r}
state_name <- "Tennessee"
state_abbr <- "TN"


# include 2022?
include_y0 <- TRUE

# export plots to AWS and Drive?
save_plots <- FALSE

gs_tab <- "Overview"
sub_folders <- paste0(state_abbr, "/", tolower(gs_tab), "/")


if (include_y0) {
  years <- "SFY22-25"
  # used in setting years as factor to display correctly in plots
  years_list <- c("2022", "2023", "2024", "2025") 
} else {
  years <- "SFY23-25"
  years_list <- c("2023", "2024", "2025")
}

subtitle_str <- paste0(state_name, ", ", years)
# in some cases, 22 shouldn't be shown, even if included, ie for funding sources that didn't exist prior to 23
subtitle_str_only_23 <- str_replace(subtitle_str, "22", "23")

```

# Prepare Data

```{r, warning=FALSE}

financial <- get_financial(state_name, years_list)

set_asides <- get_set_asides(state_name, years_list) %>%
  mutate(total_sa_amt_str = format_currency(total_sa_amt),
         total_sa_pct_str = format_percent(total_sa_pct)) %>%
  filter(fed_cap_grant != "Base & IIJA Gen Sup")

pf <- get_pf(state_name, years_list)

```


## Create Data Subsets

```{r, warning=FALSE}

financial <- financial %>%
  mutate(total_fcg_str = format_currency(total_fcg))

base <- financial %>%
  filter(fed_cap_grant == "Base")

gen_supp <- financial %>%
  filter(fed_cap_grant == "IIJA Gen Supp")

base_gen_supp <- financial %>%
  filter(fed_cap_grant == "Base & IIJA Gen Supp")

lead <- financial %>%
  filter(fed_cap_grant == "Lead Service Line Replacement")

ec <- financial %>%
  filter(fed_cap_grant == "Emerging Contaminants")

```


# Funding Data

## Combined Funds

```{r}

all_grants <- financial %>%
  filter(total_fcg > 0) %>%
  filter(fed_cap_grant != "Total") %>%
  mutate(plot_str = paste0(fed_cap_grant, ", ", state_fiscal_year, ':<br>',
                           total_fcg_str))

if(nrow(all_grants) > 0 &
   nrow(all_grants %>% filter(!is.na(total_fcg))) > 0) {
  
  all_grants_p <- ggplot(all_grants, aes(x=state_fiscal_year, y=total_fcg, 
                                         fill=fed_cap_grant, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    labs(x="", y="", title="Total Funds from Fed Cap Grants", fill="", subtitle=subtitle_str) + 
    scale_fill_manual(values=fed_cap_grant_colors) + 
    epic_chart_theme
  
  all_grants_gp <- ggplotly(all_grants_p, tooltip="text") %>%
    layout(title = list(text = paste0('Total Funds from Fed Cap Grants',
                                      '<br>',
                                      ifelse("2022" %in% all_grants$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = all_grants_p,
                   gp_object = all_grants_gp, # change
                   sub_folders, 
                   name = "total-funds-fed-cap-grant", # change
                   gs_tab) }
  
  all_grants_gp
  
}
```

```{r}
all_grants_pf <- pf %>%
  mutate(total_pf_str = format_currency(total_pf_amt),
         plot_str = paste0(pf_category, ", ", state_fiscal_year, ':<br>',
                           total_pf_str))

if(nrow(all_grants_pf) > 0 &
   nrow(all_grants_pf %>% filter(!is.na(total_pf_amt))) > 0) {
  
  all_grants_pf_p <- ggplot(all_grants_pf, aes(x=state_fiscal_year, 
                                               y=total_pf_amt, fill=pf_category,
                                               text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    labs(x="", y="",
         fill="",
         title="Principal Forgiveness from Fed Cap Grants", 
         subtitle=subtitle_str) + 
    scale_fill_manual(values=pf_colors) + 
    epic_chart_theme
  
  all_grants_pf_gp <- ggplotly(all_grants_pf_p, tooltip="text") %>%
    layout(title = list(text = paste0('Principal Forgiveness from Fed Cap Grants',
                                      '<br>',
                                      ifelse("2022" %in% all_grants_pf$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = all_grants_pf_p,
                   gp_object = all_grants_pf_gp, # change
                   sub_folders, 
                   name = "pf-from-fed-cap-grants", # change
                   gs_tab) }
  
  all_grants_pf_gp
  
}
```

### Stats

These tables inform the Year by Year section on each state's overview page.

```{r}
tibble(all_grants_pf %>% 
         group_by(state_fiscal_year) %>% 
         summarize(total_all_pf = format_currency(sum(total_pf_amt, na.rm=TRUE))))
```

```{r}
tibble(all_grants_pf %>% 
         group_by(state_fiscal_year) %>%
         summarize(total_all_fcg = sum(total_fcg, na.rm=TRUE)))

```


## Base 

### Fed Cap Grants


```{r}

if(sum(base$total_fcg, na.rm=TRUE) > 0) {
  
  base_fed_cap_grants_p <- ggplot(base, aes(x=state_fiscal_year, y=total_fcg, 
                                            fill=fed_cap_grant, text=total_fcg_str)) + 
    geom_bar(stat="identity") + 
    scale_fill_manual(values=fed_cap_grant_colors) + 
    guides(fill="none") + 
    scale_y_continuous(labels=label_currency()) + 
    labs(x="", y="", title="Base DWSRF Federal Capitlization Grants", subtitle=subtitle_str) + 
    epic_chart_theme
  
  base_fed_cap_grants_gp <- ggplotly(base_fed_cap_grants_p, tooltip="text") %>%
    layout(title = list(text = paste0('Base DWSRF Federal Cap Grants',
                                      '<br>',
                                      ifelse("2022" %in% base$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = base_fed_cap_grants_p,
                   gp_object = base_fed_cap_grants_gp, # change
                   sub_folders, 
                   name = "base-fed-cap-grant", # change
                   gs_tab) }
  
  base_fed_cap_grants_gp
}
```

### Leveraged Funds

```{r}


if(sum(financial$leveraged_funds, na.rm=TRUE) > 0) {
  
  leveraged <- financial %>%
    select(fed_cap_grant, state_fiscal_year, leveraged_funds) %>%
    filter(!is.na(leveraged_funds) & leveraged_funds > 0) %>%
    mutate(leveraged_funds_str = format_currency(leveraged_funds))
  
  leveraged_p <- ggplot(leveraged, aes(x=state_fiscal_year, y=leveraged_funds, 
                                       fill=fed_cap_grant, text=leveraged_funds_str)) + 
    geom_bar(stat="identity") + 
    scale_y_continuous(labels=label_currency()) + 
    labs(x="", y="", title="Leveraged Funds", fill="", subtitle=subtitle_str) + 
    scale_fill_manual(values=fed_cap_grant_colors) + 
    epic_chart_theme
  
  leveraged_gp <- ggplotly(leveraged_p, tooltip="text") %>%
    layout(title = list(text = paste0('Leveraged Funds',
                                      '<br>',
                                      ifelse("2022" %in% leveraged$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = leveraged_p,
                   gp_object = leveraged_gp, # change
                   sub_folders, 
                   name = "leveraged-funds", # change
                   gs_tab) }
  
  
  leveraged_gp
}
```

### Unutilized Funds

```{r}
unutilized <- financial %>%
  filter(unutilized_fcg > 0 & !is.na(unutilized_fcg)) %>%
  filter(fed_cap_grant != "Total") %>%
  mutate(unutilized_fcg_str = format_currency(unutilized_fcg))

if (nrow(unutilized) > 0) {
  
  unutilized_p <- ggplot(unutilized, aes(x=state_fiscal_year, y=unutilized_fcg, 
                                         fill=fed_cap_grant, text=unutilized_fcg_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values=fed_cap_grant_colors) + 
    labs(x="", y="", title="Unutilized Funds (Carryover)", subtitle=subtitle_str, fill="") + 
    epic_chart_theme
  
  unutilized_gp <- ggplotly(unutilized_p, tooltip="text") %>%
    layout(title = list(text = paste0('Unutilized Funds (Carryover)',
                                      '<br>',
                                      ifelse("2022" %in% unutilized$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = unutilized_p,
                   gp_object = unutilized_gp, # change
                   sub_folders, 
                   name = "unutilized-funds", # change
                   gs_tab) }
  
  unutilized_gp
}
```


### Base PF Percent

```{r}
base_pf_both_pct <- pf %>%
  filter(fed_cap_grant == "Base") %>%
  select(state_fiscal_year, total_pf_pct, pf_category) %>%
  mutate(max = case_when(
    pf_category == "DAC" ~ "35%",
    pf_category == "Discretionary" ~ "14%",
    TRUE ~ ""
  )) %>%
  mutate(plot_str = paste0(pf_category, ", ", state_fiscal_year, "<br>", 
                           format_percent(total_pf_pct),
                           " (Possible Max: ", max, ")"))

if(sum(base_pf_both_pct$total_pf_pct, na.rm=TRUE) > 0) {
  
  base_pf_both_pct_p <- ggplot(base_pf_both_pct, 
                               aes(x=state_fiscal_year, y=total_pf_pct, 
                                   color=pf_category, text=plot_str)) + 
    geom_line(aes(group=pf_category), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0,.4)) + 
    scale_color_manual(values = pf_colors) + 
    labs(x="", y="", color="", 
         title="Percent of Base Fed Cap Grant reserved for DAC & Discretionary PF",
         subtitle=subtitle_str) + 
    epic_chart_theme
  
  base_pf_both_pct_gp <- ggplotly(base_pf_both_pct_p, tooltip="text") %>%
    layout(title = list(text = paste0('Percent of Base Fed Cap Grant reserved for DAC & Discretionary PF',
                                      '<br>',
                                      ifelse("2022" %in% base_pf_both_pct$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = base_pf_both_pct_p,
                   gp_object = base_pf_both_pct_gp, # change
                   sub_folders, 
                   name = "base-pf-percent", # change
                   gs_tab) }
  
  base_pf_both_pct_gp
}
```

### Base PF Amount

```{r}
base_pf_combined <- pf %>%
  filter(fed_cap_grant == "Base") %>%
  select(fed_cap_grant, state_fiscal_year, pf_category, total_fcg, total_pf_amt)

# combine the PF categories to get the correct difference between PF and grant, 
# then combine them together to plot as a single variable
base_pf_calc <- base_pf_combined %>%
  select(fed_cap_grant, state_fiscal_year, total_pf_amt, total_fcg) %>%
  group_by(fed_cap_grant, state_fiscal_year) %>%
  summarize(combined_pf_amt = sum(total_pf_amt, na.rm=TRUE),
            total_fcg = max(total_fcg)) %>%
  mutate(pf_category = "Remaining Loans",
         total_pf_amt = total_fcg - combined_pf_amt) 

base_pf_combined <- bind_rows(base_pf_combined, base_pf_calc) %>%
  select(-combined_pf_amt) %>%
  mutate(
    pf_category = factor(pf_category, levels=c("Remaining Loans", "Discretionary", "DAC")),
    amt_value_str = format_currency(total_pf_amt),
    pct_value = total_pf_amt / total_fcg,
    pct_value_str = format_percent(pct_value),
    plot_str = paste0(pf_category, ", ", state_fiscal_year, '<br>', amt_value_str, " (", pct_value_str, ")")
  )

if(sum(base_pf_combined$total_pf_amt) > 0) {
  
  base_pf_combined_p <- ggplot(base_pf_combined, aes(x=state_fiscal_year, 
                                                     y=total_pf_amt, 
                                                     fill=pf_category, text=plot_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_dollar()) +
    scale_fill_manual(values = pf_colors) + 
    labs(x="", y="", fill="", title="Base Grant Principal Forgiveness", subtitle=subtitle_str) + 
    epic_chart_theme
  
  base_pf_combined_gp <- ggplotly(base_pf_combined_p, tooltip="text") %>%
    layout(title = list(text = paste0('Base Grant Principal Forgiveness',
                                      '<br>',
                                      ifelse("2022" %in% base_pf_combined$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = base_pf_combined_p,
                   gp_object = base_pf_combined_gp, # change
                   sub_folders, 
                   name = "base-pf-amount", # change
                   gs_tab) }
  
  base_pf_combined_gp
}
```

#IIJA Gen/Supp

```{r}

if (sum(gen_supp$total_fcg > 0)) {
  
  gensupp_fed_cap_grants_p <- ggplot(gen_supp, aes(x=state_fiscal_year, 
                                                   y=total_fcg, fill=fed_cap_grant, 
                                                   text=format_currency(total_fcg))) + 
    geom_bar(stat="identity") +
    scale_fill_manual(values=fed_cap_grant_colors) + 
    guides(fill="none") + 
    scale_y_continuous(labels=label_currency()) + 
    labs(x="", y="", 
         title="IIJA Gen. Supp. Federal Cap Grants", subtitle=subtitle_str) + 
    epic_chart_theme
  
  gensupp_fed_cap_grants_gp <- ggplotly(gensupp_fed_cap_grants_p, tooltip="text") %>%
    layout(title = list(text = paste0('IIJA Gen. Supp. Federal Cap Grants',
                                      '<br>',
                                      ifelse("2022" %in% gen_supp$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = gensupp_fed_cap_grants_p,
                   gp_object = gensupp_fed_cap_grants_gp, # change
                   sub_folders, 
                   name = "gensupp-fed-cap-grant", # change
                   gs_tab) }
  
  gensupp_fed_cap_grants_gp
}
```


# Set Aside Data

## By Fed Cap Grant

```{r}

set_asides_fcg_group <- set_asides %>%
  group_by(fed_cap_grant, state_fiscal_year) %>%
  summarize(total_sa_amt = sum(total_sa_amt),
            total_fcg = max(total_fcg),
            total_sa_pct = total_sa_amt / total_fcg) %>%
  mutate(total_sa_amt_str = format_currency(total_sa_amt),
         total_sa_pct_str = format_percent(total_sa_pct))

if(sum(set_asides_fcg_group$total_sa_amt, na.rm=TRUE) > 0) {
  
  set_asides_p <- ggplot(set_asides_fcg_group, aes(x=state_fiscal_year, 
                                                   y=total_sa_amt, 
                                                   fill=fed_cap_grant, text=total_sa_amt_str)) + 
    geom_bar(stat="identity", position="stack") + 
    scale_y_continuous(labels=label_currency()) + 
    scale_fill_manual(values = fed_cap_grant_colors, 
                      name = "") + 
    labs(x="", y="", title="Set Asides, Fed Cap Grant", subtitle=subtitle_str) + 
    epic_chart_theme
  
  set_asides_gp <- ggplotly(set_asides_p, tooltip="text") %>%
    layout(title = list(text = paste0('Total amount of set-asides taken per fed cap grant',
                                      '<br>',
                                      ifelse("2022" %in% set_asides_fcg_group$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = set_asides_p,
                   gp_object = set_asides_gp, # change
                   sub_folders, 
                   name = "set-asides-fed-cap-grant-amount", # change
                   gs_tab) }
  
  
  set_asides_gp
}
```

```{r}
set_asides_percent <- set_asides_fcg_group %>%
  filter(!is.na(total_sa_pct))

if(sum(set_asides_percent$total_sa_pct, na.rm=TRUE) > 0) {
  
  set_asides_pct_p <- ggplot(set_asides_percent, aes(x=state_fiscal_year, y=total_sa_pct, color=fed_cap_grant, text=total_sa_pct_str)) + 
    geom_line(aes(group=fed_cap_grant), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0, max(set_asides_percent$total_sa_pct)+.02 )) + 
    labs(x="", y="", title="Set Asides, Percent Utilized", subtitle=subtitle_str) + 
    scale_color_manual(values = fed_cap_grant_colors, 
                       guide="none") + 
    epic_chart_theme
  
  set_asides_pct_gp <- ggplotly(set_asides_pct_p, tooltip="text") %>%
    layout(title = list(text = paste0('Percent of set-asides taken per fed cap grant',
                                      '<br>',
                                      ifelse("2022" %in% set_asides_percent$state_fiscal_year, subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = set_asides_pct_p,
                   gp_object = set_asides_pct_gp, # change
                   sub_folders, 
                   name = "set-asides-percent-utilized", # change
                   gs_tab) }
  
  set_asides_pct_gp
}
```

## By Allowances

```{r}

sa_pct_base <- set_asides %>%
  filter(fed_cap_grant == "Base" & !is.na(total_sa_pct)) %>%
  mutate(plot_str = paste0(allowance, ", ", state_fiscal_year, ":<br>",
                           total_sa_pct_str, " (", max_set_aside, ")"))

if (sum(sa_pct_base$total_sa_pct, na.rm=TRUE) >0 ) {
  
  sa_pct_base_p <- ggplot(sa_pct_base, aes(x=state_fiscal_year, y=total_sa_pct, 
                                           color=allowance, text=plot_str)) + 
    geom_line(aes(group=allowance), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0, max(sa_pct_base$total_sa_pct)+.02)) +
    labs(x="", y="", 
         title="Set aside percentage per category under base program",
         subtitle=subtitle_str, color="") + 
    scale_color_manual(values = allowance_colors) + 
    epic_chart_theme
  
  sa_pct_base_gp <- ggplotly(sa_pct_base_p, tooltip="text") %>%
    layout(title = list(text = paste0('Set aside percentage per category under base program',
                                      '<br>',
                                      ifelse("2022" %in% sa_pct_base$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = sa_pct_base_p,
                   gp_object = sa_pct_base_gp, # change
                   sub_folders, 
                   name = "set-asides-allowance-percent-base", # change
                   gs_tab) }
  
  sa_pct_base_gp
}

```

```{r}
sa_pct_iija <- set_asides %>%
  filter(fed_cap_grant == "IIJA Gen Supp" & !is.na(total_sa_pct)) %>%
  mutate(plot_str = paste0(allowance, ", ", state_fiscal_year, ":<br>",
                           total_sa_pct_str, " (", max_set_aside, ")"))

if (sum(sa_pct_iija$total_sa_pct, na.rm=TRUE) >0 ) {
  
  sa_pct_iija_p <- ggplot(sa_pct_iija, aes(x=state_fiscal_year, y=total_sa_pct, 
                                           color=allowance, text=plot_str)) + 
    geom_line(aes(group=allowance), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0, max(sa_pct_iija$total_sa_pct)+.02)) + 
    labs(x="", y="", 
         color="",
         title="Set aside percentage per category under Gen Supp program", 
         subtitle=subtitle_str) + 
    scale_color_manual(values = allowance_colors) + 
    epic_chart_theme
  
  sa_pct_iija_gp <- ggplotly(sa_pct_iija_p, tooltip="text") %>%
    layout(title = list(text = paste0('Set aside percentage per category under Gen Supp program',
                                      '<br>',
                                      ifelse("2022" %in% sa_pct_iija$state_fiscal_year,
                                             subtitle_str, subtitle_str_only_23))))
  
  
  
  if (save_plots) {
    run_save_plots(gg_plot = sa_pct_iija_p,
                   gp_object = sa_pct_iija_gp, # change
                   sub_folders, 
                   name = "set-asides-allowance-percent-iija", # change
                   gs_tab) }
  
  sa_pct_iija_gp
}
```

```{r}

sa_pct_lead <- set_asides %>%
  filter(fed_cap_grant == "LSLR" & !is.na(total_sa_pct)) %>%
  mutate(plot_str = paste0(allowance, ", ", state_fiscal_year, ":<br>",
                           total_sa_pct_str, " (", max_set_aside, ")"))

if (sum(sa_pct_lead$total_sa_pct, na.rm=TRUE) >0) {
  
  
  sa_pct_lead_p <- ggplot(sa_pct_lead, aes(x=state_fiscal_year, 
                                           y=total_sa_pct, color=allowance, 
                                           text=plot_str)) + 
    geom_line(aes(group=allowance), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0, max(sa_pct_lead$total_sa_pct)+.02)) + 
    labs(x="", y="", title="Set Asides by Allowance, Percent", subtitle=subtitle_str, color="") + 
    scale_color_manual(values = allowance_colors) + 
    epic_chart_theme
  
  sa_pct_lead_gp <- ggplotly(sa_pct_lead_p, tooltip="text") %>%
    layout(title = list(text = paste0('Set aside percentage per category under Lead program',
                                      '<br>',
                                      ifelse("2022" %in% sa_pct_lead$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = sa_pct_lead_p,
                   gp_object = sa_pct_lead_gp, # change
                   sub_folders, 
                   name = "set-asides-allowance-percent-lead", # change
                   gs_tab) }
  
  sa_pct_lead_gp
}
```

```{r}

sa_pct_ec <- set_asides %>%
  filter(fed_cap_grant == "EC" & !is.na(total_sa_pct)) %>%
  mutate(plot_str = paste0(allowance, ", ", state_fiscal_year, ":<br>",
                           total_sa_pct_str, " (", max_set_aside, ")"))

if (sum(sa_pct_ec$total_sa_pct, na.rm=TRUE) >0 ) {
  
  sa_pct_ec_p <- ggplot(sa_pct_ec, aes(x=state_fiscal_year, y=total_sa_pct, 
                                       color=allowance, text=plot_str)) + 
    geom_line(aes(group=allowance), linewidth=1.5) + 
    geom_point(size=4) + 
    scale_y_continuous(labels=label_percent(), limits=c(0, max(sa_pct_ec$total_sa_pct)+.02)) + 
    labs(x="", y="", title="Set Asides by Allowance, Percent", subtitle=subtitle_str, color="") + 
    scale_color_manual(values = allowance_colors) + 
    epic_chart_theme
  
  sa_pct_ec_gp <- ggplotly(sa_pct_ec_p, tooltip="text") %>%
    layout(title = list(text = paste0('Set aside percentage per category under EC program',
                                      '<br>',
                                      ifelse("2022" %in% sa_pct_ec$state_fiscal_year, 
                                             subtitle_str, subtitle_str_only_23))))
  
  
  if (save_plots) {
    run_save_plots(gg_plot = sa_pct_ec_p,
                   gp_object = sa_pct_ec_gp, # change
                   sub_folders, 
                   name = "set-asides-allowance-percent-ec", # change
                   gs_tab) }
  
  sa_pct_ec_gp
}

```

