---
title: "State Funding Pages"
author: "Phil Cork"
date: "2024-07-08"
---

This notebook is a draft of analysis for creating statistics and data viz for the SRF Funding Tracker's individual state funding pages (those which cover a single state fiscal year for a particular state.) It pulls data from both the project-level data and the pre-generated state-level aggregate data.

# Setup

## Set Variables

```{r, warning=FALSE}
source("../resources.R")
run_code_from_file("../aws.txt")

state_name <- "Tennessee"
state_abbr <- "TN"

sfy <- "2025"
subtitle_str <- paste0(state_name, ", SFY", str_sub(sfy,-2,-1))

years_list <- c(sfy)

save_plots <- FALSE

gs_tab <- paste0("SFY", str_sub(sfy,-2,-1))
sub_folders <- paste0(state_abbr, "/", tolower(gs_tab), "/")
```

## Imports and Styling

```{r}
# when only comparing % of a single total
donut_colors_2 <- c("#4ea324","lightgrey")
donut_colors_2r <- c("lightgrey", "#4ea324")
# when comparing 2 types of PF versus remaining grants
donut_colors_3 <- c("#4ea324",  "#172f60","lightgrey")
# when comparing 2 types of PF, 4 types of set asides, and remaining grant
donut_colors_7 <- c("#4ea324", "#82AB6E",  "#172f60", "#1054a8", "#526489", "#527CAF", "lightgrey")
donut_colors_6 <- c("#4ea324", "#172f60", "#1054a8", "#526489", "#527CAF", "lightgrey")

```


### Import Projects & State Data

```{r}
# import project-level data
state_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_counts <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-counts.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_dollars <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-dollars.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

project_types <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

```


```{r}
financial <- get_financial(state_name, years_list) %>%
  filter(state_fiscal_year == sfy)

set_asides <- get_set_asides(state_name, years_list) %>%
  filter(state_fiscal_year == sfy)

pf <- get_pf(state_name, years_list) %>%
  filter(state_fiscal_year == sfy)
```


# Funding Data

## Grants Allocated

```{r}
grants <- financial %>%
  select(fed_cap_grant, ffy21_fcg, ffy22_fcg, ffy23_fcg, ffy24_fcg, ffy25_fcg, ffy26_fcg, reallotments) %>%
  pivot_longer(cols=c(ffy21_fcg, ffy22_fcg, ffy23_fcg, ffy24_fcg, ffy25_fcg, ffy26_fcg, reallotments)) %>%
  filter(!is.na(value) & value > 0) %>%
  mutate(name = str_to_sentence(str_remove(name, "_fcg")),
         name = str_replace(name, "Ffy", "FFY"),
         name = paste0(fed_cap_grant, ", ", name)
         )
```


```{r}

# display fed cap grants by source
if (sum(grants$value, na.rm=TRUE)>0) {

grants_p <- ggplot(grants, aes(x=value, y=reorder(name, value), text=format_currency(value))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", 
       title=" ", 
       subtitle=subtitle_str) + 
  scale_x_continuous(labels=label_dollar()) + 
  epic_chart_theme

grants_gp <-  ggplotly(grants_p, tooltip="text") %>%
  layout(title = list(text = paste0('Federal Cap Grants',
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {
  run_save_plots(gg_plot = grants_p,
                 gp_object = grants_gp, # change
                 sub_folders, 
                 name = paste0("fed-cap-grant-amount-",sfy), # change
                 gs_tab) }

grants_gp
}
```



## Principal Forgiveness

```{r}
total_pf_sfy <- pf %>%
    filter(fed_cap_grant != "Base & IIJA General Supplemental") %>%
    select(fed_cap_grant, ffy21_amt, ffy22_amt, ffy23_amt, ffy24_amt, ffy25_amt, ffy26_amt) %>%
  pivot_longer(cols=c(ffy21_amt, ffy22_amt, ffy23_amt, ffy24_amt, ffy25_amt, ffy26_amt)) %>%
  filter(!is.na(value) & value > 0) %>%
  mutate(name = str_to_upper(str_remove(name, "_amt")),
         name = paste0(fed_cap_grant, ", ", name)
         ) %>%
  group_by(name) %>%
  summarize(value = sum(value, na.rm=TRUE))

```


```{r}

if(sum(pf$total_pf_amt, na.rm=TRUE) > 0) {

all_pf_p <- ggplot(pf, aes(x=total_pf_amt, y=reorder(pf_category, total_pf_amt), text=format_currency(total_pf_amt))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title=" ", 
       subtitle=subtitle_str) + 
  scale_x_continuous(labels=label_dollar()) + 
  epic_chart_theme

all_pf_gp <-  ggplotly(all_pf_p, tooltip="text") %>%
  layout(title = list(text = paste0("How much of the fed cap grants are expected to be provided as principal forgiveness?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {
  run_save_plots(gg_plot = all_pf_p,
                 gp_object = all_pf_gp, # change
                 sub_folders, 
                 name = paste0("fed-cap-grant-pf-amount-",sfy), # change
                 gs_tab) }

all_pf_gp
}
```

## Base, Set Asides, PF Percents

```{r}

base_distribution <-data.frame(name=c(""), value=c(0))

base_distribution <- base_distribution %>%
  add_row(name="DAC PF", 
          value=pf$total_pf_amt[pf$pf_category=="DAC"]) %>%
  add_row(name="Discretionary PF",
          value=pf$total_pf_amt[pf$pf_category=="Discretionary"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="Base" & set_asides$allowance == "Local Assistance & Other"])

base_distribution <- base_distribution %>%
  add_row(name="Remaining Base Grant, for Loans",
          value = financial$total_fcg[financial$fed_cap_grant=="Base"] - sum(base_distribution$value)) %>%
  filter(name!="")

```

```{r}
if (nrow(base_distribution %>% filter(!is.na(value))) > 0) {
  
  base_all_fig <- base_distribution %>% plot_ly(labels = ~name, values = ~value,
                                       hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_7)) %>% 
  layout(title = paste0("How are the Base grants allocated?", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = base_all_fig, # change
                 sub_folders, 
                 name = paste0("base-pf-sa-pct-",sfy), # change
                 gs_tab) }

base_all_fig
}
```


## Gen Supp, Set Asides, PF Percents

```{r}

gensupp_distribution <-data.frame(name=c(""), value=c(0))

gensupp_distribution <- gensupp_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA Gen Supp"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="IIJA Gen Supp" & set_asides$allowance == "Local Assistance & Other"])

gensupp_distribution <- gensupp_distribution %>%
  add_row(name="Remaining Gen Supp grant, for loans",
          value = financial$total_fcg[financial$fed_cap_grant=="IIJA Gen Supp"] - sum(gensupp_distribution$value, na.rm=TRUE)) %>%
  filter(name!="")

```

```{r}
if (nrow(gensupp_distribution %>% filter(!is.na(value))) > 0) {
  
  gensupp_all_fig <- gensupp_distribution %>% plot_ly(labels = ~name, values = ~value,
                                       hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_6)) %>% 
  layout(title = paste0("How is the IIJA Gen Supp grant allocated?", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = gensupp_all_fig, # change
                 sub_folders, 
                 name = paste0("iija-gen-supp-pf-sa-pct-",sfy), # change
                 gs_tab) }

gensupp_all_fig
}
```

## Lead, Set Asides, PF Percents

```{r}

lead_distribution <-data.frame(name=c(""), value=c(0))

lead_distribution <- lead_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA - LSLR"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="LSLR" & set_asides$allowance == "Local Assistance & Other"])

lead_distribution <- lead_distribution %>%
  add_row(name="Remaining LSLR grants, for loans",
          value = financial$total_fcg[financial$fed_cap_grant=="LSLR"] - sum(lead_distribution$value, na.rm=TRUE)) %>%
  filter(name!="")

```

```{r}
if (nrow(lead_distribution %>% filter(!is.na(value))) > 0) {
  
  lead_all_fig <- lead_distribution %>% plot_ly(labels = ~name, values = ~value,
                                       hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_6)) %>% 
  layout(title = paste0("How are the LSLR grants allocated?", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = lead_all_fig, # change
                 sub_folders, 
                 name = paste0("lead-pf-sa-pct-",sfy), # change
                 gs_tab) }

lead_all_fig
}
```


## EC, Set Asides, PF Percents

```{r}

ec_distribution <-data.frame(name=c(""), value=c(0))

ec_distribution <- ec_distribution %>%
  add_row(name="Principal Forgiveness", 
          value=pf$total_pf_amt[pf$pf_category=="IIJA - EC"]) %>%
  add_row(name="Admin & TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Admin & TA"]) %>%
  add_row(name="Small System TA Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Small System TA"]) %>%
  add_row(name="State Program Management Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "State Program Management"]) %>%
  add_row(name="Local Assistance & Other Set Asides", 
          value=set_asides$total_sa_amt[set_asides$fed_cap_grant=="EC" & set_asides$allowance == "Local Assistance & Other"])

ec_distribution <- ec_distribution %>%
  add_row(name="Remaining EC grant, for Loans",
          value = financial$total_fcg[financial$fed_cap_grant=="EC"] - sum(ec_distribution$value, na.rm=TRUE)) %>%
  filter(name!="")

```

```{r}
if (nrow(ec_distribution %>% filter(!is.na(value))) > 0) {
  
  ec_all_fig <- ec_distribution %>% plot_ly(labels = ~name, values = ~value,
                                       hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_6)) %>% 
  layout(title = paste0("How are the EC grants allocated?", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = ec_all_fig, # change
                 sub_folders, 
                 name = paste0("ec-pf-sa-pct-",sfy), # change
                 gs_tab) }

ec_all_fig
}
```

# Comparing Funding and Project Data

## Funds by Process

```{r}

if (!is.na(project_dollars$total_project_cost)) {

funds_process_comp <- data.frame(name=c("Available to Fund Projects", "Project Demand", "Expected Funding"),
                                 value=c(sum(financial$total_funding_available, na.rm = TRUE),
                                         project_dollars$total_project_cost,
                                         project_dollars$total_funding_amount)) %>%
  mutate(category = factor(name, levels=c("Expected Funding", "Available to Fund Projects", "Project Demand")))

} else {
 
  funds_process_comp <- data.frame(name=c("Available to Fund Projects", "Project Demand", "Expected Funding"),
                                 value=c(sum(financial$total_funding_available, na.rm = TRUE),
                                         project_dollars$total_requested_amount,
                                         project_dollars$total_funding_amount)) %>%
  mutate(category = factor(name, levels=c("Expected Funding", "Available to Fund Projects", "Project Demand"))) 
  
}
```


```{r}

if(nrow(funds_process_comp %>% filter(!is.na(value))) == nrow(funds_process_comp)) {

funds_comp_p <- ggplot(funds_process_comp, aes(y=category, x=value, text=format_currency(value))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", 
       title="How do available funds compare to demand and funding expectations?", 
       subtitle=subtitle_str) + 
  scale_x_continuous(labels=label_dollar()) + 
  epic_chart_theme

funds_comp_gp <- ggplotly(funds_comp_p, tooltip="text") %>%
  layout(title = list(text = paste0("How do available funds compare to demand and funding expectations?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = funds_comp_gp, # change
                 sub_folders, 
                 name = paste0("compare-available-demand-expecting-",sfy), # change
                 gs_tab) }

funds_comp_gp
}
```

## Project Costs & Available Funds

### General

```{r}
# even though gen funds in 2022, not making plots
if (sfy != "2022") {

# check to see if the funding available is EITHER the sum of base and iija in separate rows OR a single row of Base & IIJA Gen Supp, taking the max
  #TODO: Validate with water policy team if there is a way to denote whether it is ever double counted, sometimes double counted, or never double counted to prevent these calculations from becoming misleading.
  
base_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="Base"]),
                               0,financial$total_funding_available[financial$fed_cap_grant=="Base"])

iija_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="IIJA Gen Supp"]),
                               0,financial$total_funding_available[financial$fed_cap_grant=="IIJA Gen Supp"])

base_iija_funds_available <- ifelse(is.na(financial$total_funding_available[financial$fed_cap_grant=="Base & IIJA Gen Supp"]),
                               0,financial$total_funding_available[financial$fed_cap_grant=="Base & IIJA Gen Supp"])
  

gen_funds_available <- max(base_iija_funds_available, base_funds_available+iija_funds_available)

# if genearl project cost is missing, switch to requested amount
# NOTE may manually need to use Requested Amount instead when both are present but the latter is preferred

general <- project_types %>%
  filter(project_type == "General")

if (!is.na(general$total_project_cost)) {
gen_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(gen_funds_available,
                                 max(general$total_project_cost - gen_funds_available,0)))
} else {
  gen_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(gen_funds_available,
                                 max(general$total_requested_amount - gen_funds_available,0)))
}

gen_costs <- gen_costs %>%
  mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
         value_str = format_currency(value),
         pct_value = round(value / sum(gen_costs$value),2),
         pct_value_str = format_percent(pct_value))

# if we have gen data for given sfy, make plot
if ( nrow(gen_costs %>% filter(!is.na(value))) == nrow(gen_costs))  {
  
gen_fig <- gen_costs %>% plot_ly(labels = ~name, values = ~value,
                                   hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_2)) %>% 
  layout(title = paste0("Unmet demand compared to available Base & IIJA Gen Supp funds", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = gen_fig, # change
                 sub_folders, 
                 name = paste0("compare-available-demand-general-",sfy), # change
                 gs_tab) }


gen_fig

} # if gen_costs > 0, make plot

} # if sfy!=2022


```

### Lead

```{r}
# no lead funds before year 0, skip data and plot
if (sfy != "2022") {

lead_funds_available <- financial$total_funding_available[financial$fed_cap_grant=="LSLR"]

# if project costs is NA, use Requested Amount instead
# NOTE: may need to manually change to requested amount if both are valid but the latter is more worth highlighting

lead <- project_types %>%
  filter(project_type == "Lead")

if (!is.na(lead$total_project_cost)) {

lead_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(lead_funds_available,
                                 max(lead$total_project_cost - lead_funds_available, 0)))
} else {
  lead_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(lead_funds_available,
                                 max(lead$total_requested_amount - lead_funds_available, 0)))
}

lead_costs <- lead_costs %>%
  mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
         value_str = format_currency(value),
         pct_value = round(value / sum(lead_costs$value),2),
         pct_value_str = format_percent(pct_value))

# if we have lead data for given sfy, make plot
if ( nrow(lead_costs %>% filter(!is.na(value))) == nrow(lead_costs))  {
  
lead_fig <- lead_costs %>% plot_ly(labels = ~name, values = ~value,
                                   hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_2)) %>% 
  layout(title = paste0("Unmet demand compared to available LSLR funds", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = lead_fig, # change
                 sub_folders, 
                 name = paste0("compare-available-demand-lead-",sfy), # change
                 gs_tab) }


lead_fig

} # if lead_costs > 0, make plot

} # if sfy!=2022, check for lead


```

### EC
```{r}

# no ec funds in y0
if (sfy != "2022") {

ec_funds_available <- financial$total_funding_available[financial$fed_cap_grant=="EC"]

ec <- project_types %>%
  filter(project_type == "Emerging Contaminants")

if (!is.na(ec$total_project_cost)) {
  ec_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(ec_funds_available,
                                 max(ec$total_project_cost - ec_funds_available,0)))
} else {
  ec_costs <- data.frame(name=c("Available Funds", "Unmet Demand"),
                         value=c(ec_funds_available,
                                 max(ec$total_requested_amount - ec_funds_available,0)))
}
  
  ec_costs <- ec_costs %>%
  mutate(name = factor(name, levels=c("Unmet Demand", "Available Funds")),
         value_str = format_currency(value),
         pct_value = round(value / sum(ec_costs$value),2),
         pct_value_str = format_percent(pct_value))

# if we have ec data for sfy, make plot
if ( nrow(ec_costs %>% filter(!is.na(value))) == nrow(ec_costs))  {

  
ec_fig <- ec_costs %>% plot_ly(labels = ~name, values = ~value,
                                hovertemplate="%{label}<br>$%{value}<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_2)) %>% 
  layout(title = paste0("Unmet demand compared to available EC funds", "<br>", subtitle_str),  
         margin = list(t = 100),  
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = ec_fig, # change
                 sub_folders, 
                 name = paste0("compare-available-demand-ec-",sfy), # change
                 gs_tab) }

ec_fig
} # end if we have ec data, make plot
} # end if sfy !=2022
```

```{r}
dacs_df <- data.frame(name=c("DACs Expecting Funding", "DACs Not Expecting Funding", "DACs, No Info"),
                         value=c(project_counts$dac_ef_projects, 
                                 project_counts$dac_nef_projects, 
                                 project_counts$dac_ef_ni_projects),
                         amount_str=c(as.character(project_counts$dac_ef_projects),
                                      as.character(project_counts$dac_nef_projects), 
                                      as.character(project_counts$dac_ef_ni_projects)),
                         amount_pct=c(format_percent(project_counts$dac_ef_projects/project_counts$dac_projects),
                                      format_percent(project_counts$dac_nef_projects/project_counts$dac_projects),
                                      format_percent(project_counts$dac_ef_ni_projects/project_counts$dac_projects)
                                      ))
dacs_df$name <- factor(dacs_df$name, levels=c("DACs, No Info", "DACs Not Expecting Funding", "DACs Expecting Funding"))


if ( nrow(dacs_df %>% filter(!is.na(name))) == nrow(dacs_df))  {

dacs_ef_fig <- dacs_df %>% plot_ly(labels = ~name, values = ~value,
                                    hovertemplate="%{label}<br>%{value} projects<extra></extra>") %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_3)) %>% 
  layout(title = paste0("How many DACs could expect funding?", "<br>", subtitle_str),  
         margin = list(t = 100),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = dacs_ef_fig, # change
                 sub_folders, 
                 name = paste0("dacs-expecting-funding-",sfy), # change
                 gs_tab) }

dacs_ef_fig
}
```


```{r}

ef_df <- data.frame(category=c("Expecting Funding, DAC", 
                               "Expecting Funding, Not DAC", 
                               "Expecting Funding, No Info"),
                         amount=c(project_counts$dac_ef_projects,
                                  project_counts$ndac_ef_projects,
                                  project_counts$dac_ni_ef_projects),
                         amount_str=c(as.character(project_counts$dac_ef_projects),
                                      as.character(project_counts$ndac_ef_projects),
                                      as.character(project_counts$dac_ni_ef_projects)),
                         amount_pct=c(format_percent(project_counts$dac_ef_projects / project_counts$ef_projects),
                                      format_percent(project_counts$ndac_ef_projects / project_counts$ef_projects),
                                      format_percent(project_counts$dac_ni_ef_projects / project_counts$ef_projects)
                                      ))


if ( nrow(ef_df %>% filter(!is.na(amount))) == nrow(ef_df))  {

ef_df_fig <- ef_df %>% plot_ly(labels = ~category, values = ~amount,
                               hovertemplate = '%{label}<br>%{value} projects<extra></extra>') %>%
  add_pie(hole = 0.6, direction="clockwise", sort=FALSE, marker=list(colors=donut_colors_3)) %>% 
  layout(title = paste0("How many projects expected to receive funding are for DACs?", "<br>", subtitle_str),
         margin = list(t = 100),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

if (save_plots) {
  run_save_plots(gg_plot = NULL,
                 gp_object = ef_df_fig, # change
                 sub_folders, 
                 name = paste0("expecting-funding-projects-dacs-",sfy), # change
                 gs_tab) }

  ef_df_fig
}
```


# Project-Level Data





## Project Score Range

```{r}
# range of project scores for projects expecting funding
tibble(
state_projects %>%
  filter(expecting_funding == "Yes") %>%
  mutate(score_numeric = as.numeric(project_score)) %>%
  group_by(state, state_fiscal_year) %>%
  summarize(min_score = min(score_numeric),
            max_score = max(score_numeric))
)
```