---
title: "SRF Funding Tracker Project Data YOY Analysis"
author: "Phil Cork"
date: "2024-11-08"
---

This notebook presents data viz and statistics for the SRF Funding Tracker state overview pages based on project- and state-level data gathered through the PPL/IUP scraping and cleaning process. The outputs focus on the "Project" sections of the state overview pages.

Simply changing the `state_name` will subset the data and update the plot's subtitles along with the `years` variable. Setting `include_y0` to TRUE will include state fiscal year 2022 in the data and subsequent plots. Note that 2022 is not available for most states at this time, but may be retroactively added to compare pre-IIJA to post-IIJA results.

# State Setup

```{r}
state_name <- "Tennessee"
state_abbr <- "TN"


# include 2022?
include_y0 <- TRUE

# export plots to AWS and Drive?
save_plots <- FALSE

gs_tab <- "Overview"
sub_folders <- paste0(state_abbr, "/", tolower(gs_tab), "/")


if (include_y0) {
  years <- "SFY22-25"
  # used in setting years as factor to display correctly in plots
  years_list <- c("2022", "2023", "2024", "2025") 
 } else {
  years <- "SFY23-25"
  years_list <- c("2023", "2024", "2025")
}

subtitle_str <- paste0(state_name, ", ", years)

```

# Imports & Settings 

```{r}
source("../resources.R")
run_code_from_file("../aws.txt")
```

# State-Level Data

```{r}
project_counts <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-counts.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(state_fiscal_year = as.factor(state_fiscal_year))

project_type_df <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-types.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(state_fiscal_year = as.factor(state_fiscal_year),
         project_type = factor(project_type, levels=c("Emerging Contaminants", "Lead", "General")))

small_community_df <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-small-community.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(state_fiscal_year = as.factor(state_fiscal_year))

feature_coverage <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-feature-coverage.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(state_fiscal_year = as.factor(state_fiscal_year))

project_dollars <- s3read_using(read.csv, object="s3://funding-tracker/state-stats/dwsrf-funding-tracker-project-dollars.csv") %>%
  clean_names() %>%
  filter(state == state_name) %>%
  mutate(total_project_cost_str = format_currency(total_project_cost),
         total_requested_amount_str = format_currency(total_requested_amount),
         total_principal_forgiveness_str = format_currency(total_principal_forgiveness),
         total_funding_amount_str = format_currency(total_funding_amount))

```


```{r}

if (include_y0 == FALSE) {
# select single state and drop any Y0 data to only evaluate IIJA years
  project_counts <- project_counts %>%
    filter(state_fiscal_year != "2022")
  
  project_type_df <- project_type_df %>%
    filter(state_fiscal_year != "2022")
  
  small_community_df <- small_community_df %>%
    filter(state_fiscal_year != "2022")
  
  feature_coverage <- feature_coverage %>%
    filter(state_fiscal_year != "2022")
  
  project_dollars <- project_dollars %>%
    filter(state_fiscal_year != "2022")
 }

```

# Data Viz

## Single Variable Bar Plots

### Applicants

```{r}
projects_p <- ggplot(project_counts, aes(x=state_fiscal_year, y=projects, text=projects)) + 
  geom_bar(stat="identity", fill="#526489") +
  labs(title="How many applicant projects per year?",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Projects") + 
  epic_chart_theme

projects_gp <- ggplotly(projects_p, tooltip="text") %>%
  layout(title = list(text = paste0('How many applicant projects per year?',
                                    '<br>',
                                    subtitle_str)))

if (save_plots) {

  run_save_plots(gg_plot = projects_p, #change
                 gp_object = projects_gp, #change
                 sub_folders, 
                 name = "applicant-projects-yoy", # change
                 gs_tab)
}

projects_gp
```

### Requested Amount

```{r}

if ( nrow(project_dollars %>% filter(!is.na(total_requested_amount)))  > 0 ) {


tra_p <- ggplot(project_dollars, aes(x=state_fiscal_year, y=total_requested_amount, text=total_requested_amount_str)) + 
  geom_bar(stat="identity", fill="#82AB6E") +
  scale_y_continuous(labels=label_dollar()) + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Requested Funds") + 
  epic_chart_theme

tra_gp <- ggplotly(tra_p, tooltip="text") %>%
  layout(title = list(text = paste0('Requested Funds',
                                    '<br>',
                                    subtitle_str)))

tra_gp
 if (save_plots) {

     run_save_plots(gg_plot = tra_p, #change
                    gp_object = tra_gp, #change
                    sub_folders, 
                    name = "total-requested-amount-yoy", # change
                    gs_tab)
   }
}

```

### Project Cost

```{r}
if ( nrow(project_dollars %>% filter(!is.na(total_project_cost)))  > 0 ) {
  
tpc_p <- ggplot(project_dollars, aes(x=state_fiscal_year, y=total_project_cost, text=total_project_cost_str)) + 
  geom_bar(stat="identity", fill="#82AB6E") +
  scale_y_continuous(labels=label_dollar()) + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Project Cost") + 
  epic_chart_theme

tpc_gp <- ggplotly(tpc_p, tooltip="text") %>%
  layout(title = list(text = paste0("What are the total project costs of submitted applications?",
                                    '<br>',
                                    subtitle_str)))

if (save_plots) {
  
          run_save_plots(gg_plot = tpc_p, #change
                         gp_object = tpc_gp, #change
                         sub_folders, 
                         name = "total-project-cost-yoy", # change
                         gs_tab)
  }

tpc_gp
}

```

### Funding Amount

```{r}
if ( nrow(project_dollars %>% filter(!is.na(total_funding_amount)))  > 0 ) {

tfa_p <- ggplot(project_dollars, aes(x=state_fiscal_year, y=total_funding_amount, text=total_funding_amount_str)) + 
  geom_bar(stat="identity", fill="#82AB6E") +
  scale_y_continuous(labels=label_dollar()) + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Expected Funding Amount") + 
  epic_chart_theme

tfa_gp <- ggplotly(tfa_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the total expected funding per year?",
                                    '<br>',
                                    subtitle_str)))

 if (save_plots) {
   
             run_save_plots(gg_plot = tfa_p, #change
                            gp_object = tpfa_gp, #change
                            sub_folders, 
                            name = "total-funding-amount-yoy", # change
                            gs_tab) 
   }

tfa_gp
}
```

```{r}
if ( nrow(project_dollars %>% filter(!is.na(total_principal_forgiveness)))  > 0 ) {


tpf_p <- ggplot(project_dollars, aes(x=state_fiscal_year, y=total_principal_forgiveness, text=total_principal_forgiveness_str)) + 
  geom_bar(stat="identity", fill="#82AB6E") +
  scale_y_continuous(labels=label_dollar()) + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Expecting Principal Forgiveness") + 
  epic_chart_theme

tpf_gp <- ggplotly(tpf_p, tooltip="text") %>%
  layout(title = list(text = paste0("How much PF is expected each year?",
                                    '<br>',
                                    subtitle_str)))


 if (save_plots) {
   
             run_save_plots(gg_plot = tpf_p, #change
                            gp_object = tpf_gp, #change
                            sub_folders, 
                            name = "total-principal-forgiveness-yoy", # change
                            gs_tab)
                 }

tpf_gp
}
```

### Expecting Funding Comps

```{r}

if (nrow(project_counts %>% filter(!is.na(ef_projects))) > 0) {

  ef_project_comp <- project_counts %>%
    select(state, state_fiscal_year, ef_projects, nef_projects, ef_ni_projects, projects) %>%
    pivot_longer(cols=c("ef_projects", "nef_projects", "ef_ni_projects")) %>%
    mutate(expecting_funding = case_when(
      name == "ef_projects" ~ "Yes",
      name == "nef_projects" ~ "No",
      name == "ef_ni_projects" ~ "No Info"),
      expecting_funding = factor(expecting_funding, levels=c("No Info", "No", "Yes")),
      project_percent = value / projects,    project_percent_str = format_percent(project_percent),
    plot_str = paste0(expecting_funding, ": ", as.character(value), " (", project_percent_str, ")")
    )
  

efp_p <- ggplot(ef_project_comp, aes(x=state_fiscal_year, y=value, fill=expecting_funding, text=plot_str)) + 
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Expecting Funding") + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Projects") + 
  epic_chart_theme

efp_gp <- ggplotly(efp_p, tooltip="text") %>%
  layout(title = list(text = paste0("What proportion of projects expect funding?",
                                    '<br>',
                                    subtitle_str)))

if (save_plots) {

    run_save_plots(gg_plot = efp_p,
                   gp_object = efp_gp, # change
                   sub_folders, 
                   name = "expecting-funding-applicant-comparison-yoy", # change
                   gs_tab) }
   

efp_gp
}

```


### Disadvantaged Comps

```{r}
if (nrow(project_counts %>% filter(!is.na(dac_projects))) > 0) {
  
    dac_project_comp <- project_counts %>%
    select(state, state_fiscal_year, dac_projects, ndac_projects, dac_ni_projects, projects) %>%
    pivot_longer(cols=c("dac_projects", "ndac_projects", "dac_ni_projects")) %>%
    mutate(disadvantaged = case_when(
      name == "dac_projects" ~ "Yes",
      name == "ndac_projects" ~ "No",
      name == "dac_ni_projects" ~ "No Info"),
      disadvantaged = factor(disadvantaged, levels=c("No Info", "No", "Yes")),
      project_percent = value / projects,    project_percent_str = format_percent(project_percent),
    plot_str = paste0(disadvantaged, ": ", as.character(value), " (", project_percent_str, ")")
    )


dcp_p <- ggplot(dac_project_comp, aes(x=state_fiscal_year, y=value, fill=disadvantaged, text=plot_str)) + 
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Disadvantaged") + 
  labs(title=" ",
       subtitle=subtitle_str,
       x="State Fiscal Year",
       y="Projects") + 
  epic_chart_theme

dcp_gp <- ggplotly(dcp_p, tooltip="text") %>%
  layout(title = list(text = paste0("What proportion of projects served DACs?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = dcp_p,
                   gp_object = dcp_gp, # change
                   sub_folders, 
                   name = "dac-applicant-comparison-yoy", # change
                   gs_tab) }

dcp_gp
}
```

```{r}

if (nrow(project_type_df %>% filter(!is.na(dac_projects))) > 0) {

dac_project_types <- project_type_df %>%
  select(state_fiscal_year, project_type, dac_projects) %>%
  mutate(
    project_type = factor(project_type, levels=c("Emerging Contaminants", "Lead", "General")),
    plot_str = paste0(project_type, ", ", state_fiscal_year,": <br>", dac_projects, " projects")
  )

dac_pt_p <- ggplot(dac_project_types, aes(x=state_fiscal_year, y=dac_projects, fill=project_type, text=plot_str)) + 
  geom_bar(stat="identity", position="stack") + 
    scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) +
  epic_chart_theme

dac_pt_gp <- ggplotly(dac_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("What type of projects serve DACs?",
                                    '<br>',
                                    subtitle_str)))

if (save_plots) {

    run_save_plots(gg_plot = dac_pt_p,
                   gp_object = dac_pt_gp, # change
                   sub_folders, 
                   name = "dac-project-types-yoy", # change
                   gs_tab) }

dac_pt_gp

}
```

## Expecting Funding & DAC

### How many DACs are EF?
```{r}

if(nrow(project_counts %>% filter(!is.na(dac_ef_projects))) > 0 &
   nrow(project_counts %>% filter(!is.na(dac_nef_projects))) > 0) {
  
  ef_dac_comps <- project_counts %>%
    select(state_fiscal_year, dac_ef_projects, dac_nef_projects, dac_ef_ni_projects) %>%
    pivot_longer(cols=c("dac_ef_projects", "dac_nef_projects", "dac_ef_ni_projects")) %>%
    rename(expecting_funding = name) %>%
    mutate(expecting_funding = case_when(
      expecting_funding == "dac_ef_projects" ~ "Yes",
      expecting_funding == "dac_nef_projects" ~ "No",
      TRUE ~ "No Info"),
      expecting_funding = factor(expecting_funding, levels=c("No Info", "No", "Yes")),
      plot_str = paste0(expecting_funding, ": ", value, " projects")
    )


ef_dac_p <- ggplot(ef_dac_comps, aes(x=state_fiscal_year, y=value, fill=expecting_funding, text=plot_str)) + 
  geom_bar(stat="identity", position="stack") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) + 
    scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Expecting Funding") + 
  epic_chart_theme

ef_dac_gp <- ggplotly(ef_dac_p, tooltip="text") %>%
  layout(title = list(text = paste0("How many DAC projects expect funding?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = ef_dac_p,
                   gp_object = ef_dac_gp, # change
                   sub_folders, 
                   name = "dacs-expecting-funding-yoy", # change
                   gs_tab) }

ef_dac_gp
}
```


### How many EFs are DACs?
```{r}

if(nrow(project_counts %>% filter(!is.na(dac_ef_projects))) > 0) {

    dac_ef_comps <- project_counts %>%
    select(state_fiscal_year, dac_ef_projects, ndac_ef_projects, dac_ni_ef_projects) %>%
    pivot_longer(cols=c("dac_ef_projects", "ndac_ef_projects", "dac_ni_ef_projects")) %>%
    rename(disadvantaged = name) %>%
    mutate(disadvantaged = case_when(
      disadvantaged == "dac_ef_projects" ~ "Yes",
      disadvantaged == "ndac_ef_projects" ~ "No",
      TRUE ~ "No Info"),
      disadvantaged = factor(disadvantaged, levels=c("No Info", "No", "Yes")),
      plot_str = paste0(disadvantaged, ": ", value, " projects")
    )
  
dac_ef_p <- ggplot(dac_ef_comps, aes(x=state_fiscal_year, y=value, fill=disadvantaged,
                                      text = plot_str)) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_continuous(labels=label_number(suffix="%"), expand = c(0, 0), limits = c(0, 100)) + 
  labs(x="State Fiscal Year",
       y="",
       title="How many projects expecting funding serve DACs?",
       subtitle=subtitle_str) +
  scale_fill_manual(values=ef_dac_sc_colors) + 
  epic_chart_theme

dac_ef_gp <- ggplotly(dac_ef_p, tooltip="text") %>%
  layout(title = list(text = paste0("How many projects expecting funding serve DACs?",
                                    '<br>',
                                    subtitle_str)))


dac_ef_gp
if (save_plots) {

    run_save_plots(gg_plot = dac_ef_p,
                   gp_object = dac_ef_gp, # change
                   sub_folders, 
                   name = "expecting-funding-serving-dacs-yoy", # change
                   gs_tab) }


}
```

### DAC/EF Heat Map
```{r}
dac_ef_grid <- project_counts %>%
  select(state_fiscal_year, 
         dac_ef_projects, dac_nef_projects, dac_ef_ni_projects,
         ndac_ef_projects, ndac_nef_projects, ndac_ef_ni_projects,
         dac_ni_ef_projects, dac_ni_nef_projects, dac_ni_ef_ni_projects
        ) %>%
  pivot_longer(cols=c("dac_ef_projects", "dac_nef_projects", "dac_ef_ni_projects",
         "ndac_ef_projects", "ndac_nef_projects", "ndac_ef_ni_projects",
         "dac_ni_ef_projects", "dac_ni_nef_projects", "dac_ni_ef_ni_projects")) %>%
  mutate(dac = case_when(
    name %in% c("dac_ef_projects", "dac_nef_projects", "dac_ef_ni_projects") ~ "Yes",
    name %in% c("ndac_ef_projects", "ndac_nef_projects", "ndac_ef_ni_projects") ~ "No",
    TRUE ~ "No Info"),
         expecting_funding = case_when(
    name %in% c("dac_ef_projects", "ndac_ef_projects", "dac_ni_ef_projects") ~ "Yes",
    name %in% c("dac_nef_projects", "ndac_nef_projects", "dac_ni_nef_projects") ~ "No",
    TRUE ~ "No Info") 
    ) %>%
  select(-name)
  

test <- ggplot(dac_ef_grid, aes(x=expecting_funding, y=dac, fill=value, text=value)) + 
  geom_tile(color="white") + 
  labs(title="Projects based on DAC Status and Expecting Funding", subtitle=subtitle_str,
       x="Expecting Funding", y="Disadvantaged", fill="") + 
  epic_chart_theme

test_facet <- test + facet_wrap(~state_fiscal_year)

ggplotly(test_facet, tooltip="text") %>%
  layout(title = list(text = paste0("Projects based on DAC Status and Expecting Funding",
                                    '<br>',
                                    subtitle_str))
  )
  
```

## Project Type

### Applicants
```{r}

if(sum(project_type_df$projects, na.rm=TRUE) > 0) {

all_pt_p <- ggplot(project_type_df, aes(x=state_fiscal_year, y=projects, fill=project_type, 
                                        text=paste0(project_type, ", ", state_fiscal_year, ":<br>", projects, " projects"))) + 
  geom_bar(stat="identity", position="stack") + 
    scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) +
  epic_chart_theme

all_pt_gp <- ggplotly(all_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("Applicant Projects by Type",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = all_pt_p,
                   gp_object = all_pt_gp, # change
                   sub_folders, 
                   name = "project-types-yoy", # change
                   gs_tab) }

all_pt_gp
}
```

### Project Cost
```{r}
if(sum(project_type_df$total_project_cost, na.rm=TRUE) > 0) {

tpc_pt_p <- ggplot(project_type_df, aes(x=state_fiscal_year, y=total_project_cost, fill=project_type, 
                                       text=format_currency(total_project_cost))) + 
  geom_bar(stat="identity", position="stack") + 
      scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  scale_y_continuous(labels=label_dollar()) + 
  labs(x="State Fiscal Year",
       y="",
       title=" ",
       subtitle=subtitle_str) +
  epic_chart_theme

tpc_pt_gp <- ggplotly(tpc_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("What are the total project costs for each type of project?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = tpc_pt_p,
                   gp_object = tpc_pt_gp, # change
                   sub_folders, 
                   name = "total-project-cost-project-type-yoy", # change
                   gs_tab) }

tpc_pt_gp
}
```

### Requested Amount
```{r}

if(sum(project_type_df$total_requested_amount, na.rm=TRUE) > 0)  {

tra_pt_p <- ggplot(project_type_df, aes(x=state_fiscal_year, y=total_requested_amount, fill=project_type, text=format_currency(total_requested_amount))) + 
  geom_bar(stat="identity", position="stack") + 
      scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  scale_y_continuous(labels=label_dollar()) + 
  labs(x="State Fiscal Year",
       y="",
       title="Requested Amount by Type",
       subtitle=subtitle_str) +
  epic_chart_theme


tra_pt_gp <- ggplotly(tra_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the total requested amount for each type of project?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = tra_pt_p,
                   gp_object = tra_pt_gp, # change
                   sub_folders, 
                   name = "total-requested-amount-project-type-yoy", # change
                   gs_tab) }
tra_pt_gp
}
```

### Funding Amount
```{r}
if(sum(project_type_df$total_funding_amount, na.rm=TRUE) > 0) {

tfa_pt_p <- ggplot(project_type_df, aes(x=state_fiscal_year, y=total_funding_amount, fill=project_type, text=total_funding_amount_str)) + 
  geom_bar(stat="identity", position="stack") + 
      scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  scale_y_continuous(labels=label_dollar()) + 
  labs(x="State Fiscal Year",
       y="",
       title="Funding Amount by Type",
       subtitle=subtitle_str) +
  epic_chart_theme

tfa_pt_gp <- ggplotly(tfa_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the total expected funding amount for each type of project?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = tfa_pt_p,
                   gp_object = tfa_pt_gp, # change
                   sub_folders, 
                   name = "total-funding-amount-project-type-yoy", # change
                   gs_tab) }
tfa_pt_gp
}
```

### Principal Forgiveness
```{r}

if(sum(project_type_df$total_principal_forgiveness, na.rm=TRUE) > 0) {


tpf_pt_p <- ggplot(project_type_df, aes(x=state_fiscal_year, y=total_principal_forgiveness, fill=project_type)) + 
  geom_bar(stat="identity", position="stack") + 
      scale_fill_manual(values = pt_colors, 
                      name = "Project Type") + 
  scale_y_continuous(labels=label_dollar()) + 
  labs(x="State Fiscal Year",
       y="",
       title="Principal Forgiveness by Type",
       subtitle=subtitle_str) +
  epic_chart_theme

tpf_pt_gp <- ggplotly(tpf_pt_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the total PF for each type of project?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = tpf_pt_p,
                   gp_object = tpf_pt_gp, # change
                   sub_folders, 
                   name = "total-principal-forgiveness-project-type-yoy", # change
                   gs_tab) }
tpf_pt_gp
}
```

## Small Community

### Applicants

```{r}



if(nrow(small_community_df %>% filter(!is.na(projects))) > 0) {

  
sc_pop_p <- ggplot(small_community_df, aes(x=state_fiscal_year, y=projects, fill=small_community, 
                              text=paste0(small_community, ", ", state_fiscal_year, ":<br>", projects, " projects"))) + 
  geom_bar(stat="identity", position="stack") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) + 
    scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Population < 10k") + 
  epic_chart_theme

sc_pop_gp <- ggplotly(sc_pop_p, tooltip="text") %>%
  layout(title = list(text = paste0("How many projects serve small communities?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = sc_pop_p,
                   gp_object = sc_pop_gp, # change
                   sub_folders, 
                   name = "small-community-projects-yoy", # change
                   gs_tab) }

sc_pop_gp
}

```

### Expecting Funding

```{r}
if(nrow(small_community_df %>% filter(!is.na(ef_projects))) > 0) {

sc_ef_p <- ggplot(small_community_df, aes(x=state_fiscal_year, y=ef_projects, fill=small_community, 
                                          text=paste0(small_community, ", ", state_fiscal_year, ":<br>", ef_projects, " projects"))) + 
  geom_bar(stat="identity", position="stack") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) + 
    scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Population < 10k") + 
  epic_chart_theme

sc_ef_gp <- ggplotly(sc_ef_p, tooltip="text") %>%
  layout(title = list(text = paste0("How many projects serving small communities expect funding?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = sc_ef_p,
                   gp_object = sc_ef_gp, # change
                   sub_folders, 
                   name = "small-community-expecting-funding-yoy", # change
                   gs_tab) }

sc_ef_gp
}
```


### Disadvantaged 
```{r}
if(nrow(small_community_df %>% filter(!is.na(dac_projects))) > 0) {

sc_dac_p <- ggplot(small_community_df, aes(x=state_fiscal_year, y=dac_projects, fill=small_community, 
                                  text=paste0(small_community, ", ", state_fiscal_year, ":<br>", ef_projects, " projects")))+ 
  geom_bar(stat="identity", position="stack") + 
  labs(x="State Fiscal Year",
       y="Projects",
       title=" ",
       subtitle=subtitle_str) + 
    scale_fill_manual(values = ef_dac_sc_colors, 
                      name = "Population < 10k") + 
  epic_chart_theme

sc_dac_gp <- ggplotly(sc_dac_p, tooltip="text") %>%
  layout(title = list(text = paste0("How many projects serving small communities serve DACs?",
                                    '<br>',
                                    subtitle_str)))


if (save_plots) {

    run_save_plots(gg_plot = sc_dac_p,
                   gp_object = sc_dac_gp, # change
                   sub_folders, 
                   name = "small-community-dacs-yoy", # change
                   gs_tab) }

sc_dac_gp
}
```

## Misc Plots

### Avg & Med Project Costs

```{r}

avg_med_project_cost <- project_dollars %>%
  select(state_fiscal_year, avg_project_cost, median_project_cost) %>%
  pivot_longer(cols=c("avg_project_cost", "median_project_cost")) %>%
  mutate(name = case_when(
    name == "avg_project_cost" ~ "Average",
    TRUE ~ "Median"),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value)))

if (sum(avg_med_project_cost$value) > 0) {

avg_demand_p <- ggplot(avg_med_project_cost, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_y_continuous(labels=label_dollar()) + 
  scale_fill_manual(values = c("#526489", "#82AB6E"), 
                      name = "") + 
  labs(x="State Fiscal Year",
       y="",
       title=" ",
       subtitle=subtitle_str) +
  epic_chart_theme

avg_demand_gp <- ggplotly(avg_demand_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the average demand for funds per project?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = avg_demand_p,
                   gp_object = avg_demand_gp, # change
                   sub_folders, 
                   name = "average-median-project-cost-yoy", # change
                   gs_tab) }
avg_demand_gp
}
```

### LSLI & LSLR

```{r}

lead_project_cost <- project_dollars %>%
  select(state_fiscal_year, project_cost_lsli, project_cost_lslr, project_cost_lsln) %>%
  pivot_longer(cols=c("project_cost_lsli", "project_cost_lslr", "project_cost_lsln")) %>%
  mutate(name = case_when(
    name == "project_cost_lsli" ~ "Inventory",
    name == "project_cost_lslr" ~ "Replacement",
    TRUE ~ "Neither"),
    name = factor(name, levels=c("Replacement", "Inventory", "Neither")),
    plot_str = paste0(name, ", ", state_fiscal_year, ":<br>", format_currency(value)))

if (sum(lead_project_cost$value, na.rm=TRUE) > 0) {

lead_pc_p <- ggplot(lead_project_cost, aes(x=state_fiscal_year, y=value, fill=name, text=plot_str)) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_continuous(labels=label_dollar()) + 
  scale_fill_manual(values = lead_colors, 
                      name = "") + 
  labs(x="State Fiscal Year",
       y="",
       title=" ",
       subtitle=subtitle_str) +
  epic_chart_theme

lead_pc_gp <- ggplotly(lead_pc_p, tooltip="text") %>%
  layout(title = list(text = paste0("What's the project cost for lead service line projects?",
                                    '<br>',
                                    subtitle_str)))



if (save_plots) {

    run_save_plots(gg_plot = lead_pc_p,
                   gp_object = lead_pc_gp, # change
                   sub_folders, 
                   name = "lead-project-cost-yoy", # change
                   gs_tab) }
lead_pc_gp

}
```


