---
title: "SRF Funding Tracker Other Federal and State Funds Data YOY Analysis"
author: "Maria Sevillano"
date: "2026-05-04"
---

# State Setup

## Imports & Settings

```{r}
state_name <- "Ohio"
state_abbr <- "OH"

# export plots to AWS and Drive?
save_plots <- FALSE

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(here)
source(here("resources", "inputs.R"))
source(here("resources", "styling.R"))
source(here("resources", "importing.R"))
source(here("resources", "exporting.R"))
run_code_from_file(here("aws.txt"))
```

# Read Data

```{r}
# Other Federal and State Funds Data
ofsf_all_projects <- aws.s3::s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/ofsf-funding-tracker-all-projects.csv") |>
  janitor::clean_names() |>
  dplyr::filter(state == state_name) 

# DWSRF Project Data
dwsrf_all_projects <- aws.s3::s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() |>
  dplyr::filter(state == state_name)

# Financial Data  
financial <- get_financial(state_name)
```

# Clean Data
```{r}
ofsf_all_projects_df <- ofsf_all_projects |>
  dplyr::mutate(
    dplyr::across(dplyr::everything(), as.character),
    small_community = case_when(
      population == "No Information" ~ "No Information",
      is.na(population) ~ "No Information", 
      convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 ~ "Yes",
      TRUE ~ "No"
    ),
    project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type)
  )  

dwsrf_all_projects_df <- dwsrf_all_projects |>
  dplyr::mutate(
    dplyr::across(dplyr::everything(), as.character),
    small_community = case_when(
      population == "No Information" ~ "No Information",
      is.na(population) ~ "No Information", 
      convert_to_numeric(population, TRUE) > 0 & convert_to_numeric(population, TRUE) <= 10000 ~ "Yes",
      TRUE ~ "No"
    ),
    project_type = ifelse(project_type == "Emerging Contaminants", "Emerging Contaminants (EC)", project_type)
  )  

 financial_df <-  financial |>
  dplyr::mutate(project_type = case_when(
    fed_cap_grant %in% c("LSLR", "LSLR_FFY22","LSLR_FFY23") ~ "Lead",
    fed_cap_grant %in% c("EC", "EC_FFY22", "EC_FFY23") ~ "Emerging Contaminants (EC)",
    fed_cap_grant %in% c("Base", "IIJA Gen Supp", "Base & IIJA Gen Supp", "Base and IIJA Gen Supp") ~ "General", 
    TRUE ~ paste0("Combined: ", fed_cap_grant)
  ))
```

# Prep Data

## Financial - Available Funds
```{r}
financial_df_summary_ofsf <- financial_df |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarize(total_funding_available_state_funds = sum(total_funding_available_state_funds, na.rm=TRUE),
            .groups = "drop") 

financial_df_summary_by_type_ofsf <- financial_df |>
  dplyr::group_by(state, state_fiscal_year, project_type) |>
  dplyr::summarize(total_funding_available_state_funds = sum(total_funding_available_state_funds, na.rm=TRUE),
            .groups = "drop") 

financial_df_summary_dwsrf <- financial_df |>
  dplyr::group_by(state, state_fiscal_year) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE),
            .groups = "drop") 

financial_df_summary_by_type_dwsrf <- financial_df |>
  dplyr::group_by(state, state_fiscal_year, project_type) |>
  dplyr::summarize(total_funding_available = sum(total_funding_available, na.rm=TRUE),
            .groups = "drop") 
            
```

## Funding Amount (requested amount, project cost) - On Fundable List
```{r}
# OFSF
funding_amount_ofsf_df <- ofsf_all_projects_df |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding_ofsf,
    funding_amount_ofsf,
    requested_amount_ofsf,
    project_cost_ofsf,
    disadvantaged,
    small_community
  ) |>
   dplyr::mutate(
    # convert $ data to numeric (No Information --> NA)
    dplyr::across(c(funding_amount_ofsf, requested_amount_ofsf, project_cost_ofsf), as.numeric),
    # Make any $0 NA for coalesce operation
    dplyr::across(c(funding_amount_ofsf, requested_amount_ofsf, project_cost_ofsf), \(x) ifelse(x == 0, NA, x)),
    requested_amount_ofsf_ef = requested_amount_ofsf * (expecting_funding_ofsf == "Yes"),
    total_project_cost_ofsf_ef = project_cost_ofsf * (expecting_funding_ofsf == "Yes"),
    # Choose first non NA value following hierarchy
    selected_funding_amount_ofsf = dplyr::coalesce(funding_amount_ofsf, requested_amount_ofsf_ef, total_project_cost_ofsf_ef),
    # Note funding source of selected funding amount (string check will flag any issue)
    source_funding_ofsf = dplyr::case_when(
      funding_amount_ofsf > 0 ~ "funding_amount_ofsf", 
      requested_amount_ofsf_ef > 0 ~ "requested_amount_ofsf_ef", 
      total_project_cost_ofsf_ef > 0 ~ "total_project_cost_ofsf_ef",
      .default = "check"
    )
  )

funding_amount_ofsf_df_summary <- funding_amount_ofsf_df |>
  dplyr::group_by(state_fiscal_year, source_funding_ofsf) |>
  dplyr::summarise(funding_amount_ofsf = sum(selected_funding_amount_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount_ofsf == 0) |>
  tidyr::pivot_longer(cols = "funding_amount_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

funding_amount_ofsf_df_summary_by_type <- funding_amount_ofsf_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_funding_ofsf) |>
  dplyr::summarise(funding_amount_ofsf = sum(selected_funding_amount_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount_ofsf == 0) |>
  tidyr::pivot_longer(cols = "funding_amount_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

funding_amount_ofsf_df_summary_by_sc <- funding_amount_ofsf_df |>
  dplyr::group_by(state_fiscal_year, small_community, source_funding_ofsf) |>
  dplyr::summarise(funding_amount_ofsf = sum(selected_funding_amount_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount_ofsf == 0) |>
  tidyr::pivot_longer(cols = "funding_amount_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  )   

funding_amount_ofsf_df_summary_by_dac <- funding_amount_ofsf_df |>
  dplyr::group_by(state_fiscal_year, disadvantaged, source_funding_ofsf) |>
  dplyr::summarise(funding_amount_ofsf = sum(selected_funding_amount_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount_ofsf == 0) |>
  tidyr::pivot_longer(cols = "funding_amount_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

#DWSRF

funding_amount_dwsrf_df <- dwsrf_all_projects_df |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding,
    funding_amount,
    requested_amount,
    project_cost,
    disadvantaged,
    small_community
  ) |>
   dplyr::mutate(
    # convert $ data to numeric (No Information --> NA)
    dplyr::across(c(funding_amount, requested_amount, project_cost), as.numeric),
    # Make any $0 NA for coalesce operation
    dplyr::across(c(funding_amount, requested_amount, project_cost), \(x) ifelse(x == 0, NA, x)),
    requested_amount_ef = requested_amount * (expecting_funding == "Yes"),
    total_project_cost_ef = project_cost * (expecting_funding == "Yes"),
    # Choose first non NA value following hierarchy
    selected_funding_amount = dplyr::coalesce(funding_amount, requested_amount_ef, total_project_cost_ef),
    # Note funding source of selected funding amount (string check will flag any issue)
    source_funding = dplyr::case_when(
      funding_amount > 0 ~ "funding_amount", 
      requested_amount_ef > 0 ~ "requested_amount_ef", 
      total_project_cost_ef > 0 ~ "total_project_cost_ef",
      .default = "check"
    )
  )

funding_amount_dwsrf_df_summary <- funding_amount_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, source_funding) |>
  dplyr::summarise(funding_amount = sum(selected_funding_amount, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount == 0) |>
  tidyr::pivot_longer(cols = "funding_amount") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

funding_amount_dwsrf_df_summary_by_type <- funding_amount_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_funding) |>
  dplyr::summarise(funding_amount = sum(selected_funding_amount, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount == 0) |>
  tidyr::pivot_longer(cols = "funding_amount") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

funding_amount_dwsrf_df_summary_by_sc <- funding_amount_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, small_community, source_funding) |>
  dplyr::summarise(funding_amount = sum(selected_funding_amount, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount == 0) |>
  tidyr::pivot_longer(cols = "funding_amount") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

funding_amount_dwsrf_df_summary_by_dac <- funding_amount_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, disadvantaged, source_funding) |>
  dplyr::summarise(funding_amount = sum(selected_funding_amount, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!funding_amount == 0) |>
  tidyr::pivot_longer(cols = "funding_amount") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

```

## Demand (requested amount, project cost) - Comprehensive
```{r}
# demand
demand_ofsf_df <- ofsf_all_projects_df |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding_ofsf,
    funding_amount_ofsf,
    requested_amount_ofsf,
    project_cost_ofsf,
    disadvantaged,
    small_community
  ) |>
   dplyr::mutate(
    dplyr::across(c(funding_amount_ofsf, requested_amount_ofsf, project_cost_ofsf), as.numeric),
    dplyr::across(c(funding_amount_ofsf, requested_amount_ofsf, project_cost_ofsf), \(x) ifelse(x == 0, NA, x)),
    selected_demand_ofsf = coalesce(requested_amount_ofsf, project_cost_ofsf),
    source_demand_ofsf = dplyr::case_when(
      requested_amount_ofsf > 0 ~ "requested_amount_ofsf", 
      project_cost_ofsf > 0 ~ "project_cost_ofsf", 
      .default = "check"
    )
  )

demand_ofsf_df_summary <- demand_ofsf_df |>
  dplyr::group_by(state_fiscal_year, source_demand_ofsf) |>
  dplyr::summarise(demand_ofsf = sum(selected_demand_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_ofsf == 0) |>
  tidyr::pivot_longer(cols = "demand_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_ofsf_df_summary_by_type <- demand_ofsf_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_demand_ofsf) |>
  dplyr::summarise(demand_ofsf = sum(selected_demand_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_ofsf == 0) |>
  tidyr::pivot_longer(cols = "demand_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_ofsf_df_summary_by_sc <- demand_ofsf_df |>
  dplyr::group_by(state_fiscal_year, small_community, source_demand_ofsf) |>
  dplyr::summarise(demand_ofsf = sum(selected_demand_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_ofsf == 0) |>
  tidyr::pivot_longer(cols = "demand_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_ofsf_df_summary_by_dac <- demand_ofsf_df |>
  dplyr::group_by(state_fiscal_year, disadvantaged, source_demand_ofsf) |>
  dplyr::summarise(demand_ofsf = sum(selected_demand_ofsf, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_ofsf == 0) |>
  tidyr::pivot_longer(cols = "demand_ofsf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_dwsrf_df <- dwsrf_all_projects_df |>
  dplyr::filter(state == state_name) |>
  dplyr::select(
    state_fiscal_year,
    project_type,
    expecting_funding,
    funding_amount,
    requested_amount,
    project_cost,
    disadvantaged,
    small_community
  ) |>
   dplyr::mutate(
    dplyr::across(c(funding_amount, requested_amount, project_cost), as.numeric),
    dplyr::across(c(funding_amount, requested_amount, project_cost), \(x) ifelse(x == 0, NA, x)),
    selected_demand = coalesce(requested_amount, project_cost),
    source_demand = dplyr::case_when(
      requested_amount > 0 ~ "requested_amount", 
      project_cost > 0 ~ "project_cost", 
      .default = "check"
    )
  )

demand_dwsrf_df_summary <- demand_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, source_demand) |>
  dplyr::summarise(demand_dwsrf = sum(selected_demand, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_dwsrf == 0) |>
  tidyr::pivot_longer(cols = "demand_dwsrf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_dwsrf_df_summary_by_type <- demand_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, project_type, source_demand) |>
  dplyr::summarise(demand_dwsrf = sum(selected_demand, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_dwsrf == 0) |>
  tidyr::pivot_longer(cols = "demand_dwsrf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_dwsrf_df_summary_by_sc <- demand_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, small_community, source_demand) |>
  dplyr::summarise(demand_dwsrf = sum(selected_demand, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_dwsrf == 0) |>
  tidyr::pivot_longer(cols = "demand_dwsrf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

demand_dwsrf_df_summary_by_dac <- demand_dwsrf_df |>
  dplyr::group_by(state_fiscal_year, disadvantaged, source_demand) |>
  dplyr::summarise(demand_dwsrf = sum(selected_demand, na.rm = T)) |>
  dplyr::ungroup() |>
  dplyr::filter(!demand_dwsrf == 0) |>
  tidyr::pivot_longer(cols = "demand_dwsrf") |>
  tidyr::pivot_wider(
    names_from = name,
    values_from = value
  ) 

```

## Combine all demand
```{r}
# concatenate summary dfs
available_funds_by_type <- financial_df_summary_by_type_ofsf |>
  dplyr::mutate(funding_source = "Other Funds Channeled Through SRFs") |>
  dplyr::bind_rows(financial_df_summary_by_type_dwsrf |> dplyr::mutate(funding_source = "DWSRF Funds")) |>
  dplyr::mutate(
    funding_amount = dplyr::coalesce(total_funding_available_state_funds, total_funding_available),
    bar = ifelse(grepl("Other", funding_source), "Available Other Funds Channeled Through SRFs","Available DWSRF Funds")
  ) |>
  dplyr::select(state_fiscal_year, project_type, funding_amount, funding_source, bar) |>
  dplyr::mutate(
    plt_str = paste0(bar, ": ", format_currency(funding_amount))
  )

funding_amount_by_type <- funding_amount_ofsf_df_summary_by_type |>
  dplyr::mutate(funding_source = "Other Funds Channeled Through SRFs") |>
  dplyr::bind_rows(funding_amount_dwsrf_df_summary_by_type |> dplyr::mutate(funding_source = "DWSRF Funds")) |>
  dplyr::mutate(
    source_funding = dplyr::coalesce(source_funding_ofsf, source_funding),
    funding_amount = dplyr::coalesce(funding_amount_ofsf, funding_amount),
    bar = ifelse(grepl("ofsf", source_funding), "On Fundable List Other Funds Channeled Through SRFs","On Fundable List DWSRF Funds")
  ) |>
  dplyr::select(state_fiscal_year, project_type, source_funding, funding_amount, funding_source, bar) |>
  dplyr::mutate(
    source_funding = dplyr::case_when(
      grepl("funding_amount", source_funding ) ~ "Funding Amount",
      grepl("requested_amount", source_funding) ~ "Requested Amount",
      grepl("project_cost", source_funding) ~ "Project Cost",
      .default = source_funding
    ),
    plt_str = paste0(bar, ": ", format_currency(funding_amount), " (", source_funding, ")")
  )

# Helper: ensures expected cols exist before binding
add_missing_cols <- function(df, cols, fill = NA) {
  missing <- setdiff(cols, names(df))
  df[missing] <- fill
  df
}

ofsf_cols <- c("source_demand_ofsf", "demand_ofsf")

demand_by_type <- add_missing_cols(
    demand_ofsf_df_summary_by_type, ofsf_cols
  ) |>
  dplyr::mutate(funding_source = "Other Funds Channeled Through SRFs") |>
  dplyr::bind_rows(demand_dwsrf_df_summary_by_type |> dplyr::mutate(funding_source = "DWSRF Funds")) |>
  dplyr::mutate(
    #this is demand, but keeping consistent colnames to joint dfs later
    source_funding = dplyr::coalesce(source_demand_ofsf, source_demand),
    funding_amount = dplyr::coalesce(demand_ofsf, demand_dwsrf), 
    bar = ifelse(grepl("ofsf", source_funding), "Demand Other Funds Channeled Through SRFs","Demand DWSRF Funds")
  ) |>
  dplyr::select(state_fiscal_year, project_type, source_funding, funding_amount, funding_source, bar) |>
  dplyr::mutate(
    source_funding = dplyr::case_when(
      grepl("requested_amount", source_funding) ~ "Requested Amount",
      grepl("project_cost", source_funding) ~ "Project Cost",
      .default = source_funding
    ),
    plt_str = paste0(bar, ": ", format_currency(funding_amount), " (", source_funding, ")")
  )


combined_bars_by_type <- dplyr::bind_rows(available_funds_by_type, funding_amount_by_type, demand_by_type) |>
  dplyr::filter(funding_amount >0)|>
  dplyr::mutate(
    bar = forcats::fct(bar, levels = c(
      "Available Other Funds Channeled Through SRFs",
      "Available DWSRF Funds",
  
      "On Fundable List Other Funds Channeled Through SRFs",
      "On Fundable List DWSRF Funds",

      "Demand Other Funds Channeled Through SRFs",
      "Demand DWSRF Funds"
    )),
    project_type = forcats::fct(project_type)
  )
```


# Data Viz -----
## Comparison of demand for funds, available funds, and funds on fundable list

```{r}

# 1. Define a group for dodging and a stack order within each group
bar_data <- combined_bars_by_type |>
  dplyr::filter(project_type == type) |>
  dplyr::mutate(
    dodge_group = dplyr::case_when(
      grepl("Available", plt_str)         ~ 1L,
      grepl("On Fundable List", plt_str) ~ 2L,
      grepl("Demand", plt_str)           ~ 3L
    ),
    # stack order within each dodge group (DWSRF on bottom, Other on top)
    stack_order = dplyr::case_when(
      funding_source == "DWSRF Funds"                          ~ 1L,
      funding_source == "Other Funds Channeled Through SRFs"   ~ 2L
    )
  )

# 2. Build a complete spine so every year × dodge_group × stack_order exists
spine <- tidyr::expand_grid(
  state_fiscal_year = unique(bar_data$state_fiscal_year),
  dodge_group       = 1:3,
  stack_order       = 1:2
)

bar_data <- spine |>
  dplyr::left_join(bar_data, by = c("state_fiscal_year", "dodge_group", "stack_order")) |>
  dplyr::mutate(funding_amount = tidyr::replace_na(funding_amount, 0))

# 3. Compute cumulative stack heights
bar_data <- bar_data |>
  dplyr::arrange(state_fiscal_year, dodge_group, stack_order) |>
  dplyr::group_by(state_fiscal_year, dodge_group) |>
  dplyr::mutate(
    ymax = cumsum(funding_amount),
    ymin = ymax - funding_amount
  ) |>
  dplyr::ungroup()

# 4. Compute xmin/xmax — anchor every year to the same integer regardless of missingness
bar_width <- 0.25
dodge_gap <- 0.05
group_gap <- 1.0

year_levels <- sort(unique(bar_data$state_fiscal_year))
n_groups    <- 3L

offsets <- seq(
  from       = -((n_groups - 1) / 2) * (bar_width + dodge_gap),
  by         = bar_width + dodge_gap,
  length.out = n_groups
)

bar_data <- bar_data |>
  dplyr::mutate(
    x_centre = match(state_fiscal_year, year_levels) * group_gap + offsets[dodge_group],
    xmin     = x_centre - bar_width / 2,
    xmax     = x_centre + bar_width / 2
  )

# 5. Drop zero-height rects (missing combos) so they don't affect the legend or tooltips
plot_data <- bar_data |>
  dplyr::filter(funding_amount > 0) 

# 6. Plot
  plot_tile_demand = "Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List"

  demand_p <- ggplot2::ggplot(plot_data) +
    ggplot2::geom_rect(
      ggplot2::aes(
        xmin = xmin, xmax = xmax,
        ymin = ymin, ymax = ymax,
        fill = bar,
        text = plt_str
      ),
      color = "black",
    ) +
    ggplot2::scale_x_continuous(
      breaks = seq_along(year_levels) * group_gap,
      labels = year_levels
    ) +
    ggplot2::scale_y_continuous(labels = label_currency()) +  
    ggplot2::scale_fill_manual(
      values = c(
        "Available Other Funds Channeled Through SRFs"        = "#3d7f1c",
        "Available DWSRF Funds"                               = "#6bc842",
        "On Fundable List Other Funds Channeled Through SRFs" = "#791a7b",
        "On Fundable List DWSRF Funds"                        = "#d927dcff",
        "Demand Other Funds Channeled Through SRFs"           = "#3861b2ff",
        "Demand DWSRF Funds"                                  = "#172f60"
      ),
      name = ""
    ) +
    ggplot2::labs(
      title = plot_tile_demand, 
      x = "State Fiscal Year", 
      y = "Funding Amount") +
    epic_chart_theme


demand_gp <- ggplotly(demand_p, tooltip = "text") |>
  plotly::layout(title = list(text = paste0(
    plot_tile_demand,
    '<br>',
    get_subtitle_str(bar_data$state_fiscal_year, state_name)
      )
    )
  )

if (save_plots) {
    run_save_plots(
      gg_plot   = demand_p,
      gp_object = demand_gp,
      name = "test-ofsf-demand-yoy"
      )
  }

```


## Comparison of demand for funds, available funds, and funds on fundable list by type ----

```{r}

for (type in levels(combined_bars_by_type$project_type)) {

  # 1. Define a group for dodging and a stack order within each group
bar_data_by_type <- combined_bars_by_type |>
  dplyr::filter(project_type == type) |>
  dplyr::mutate(
    dodge_group = dplyr::case_when(
      grepl("Available", plt_str)         ~ 1L,
      grepl("On Fundable List", plt_str) ~ 2L,
      grepl("Demand", plt_str)           ~ 3L
    ),
    # stack order within each dodge group (DWSRF on bottom, Other on top)
    stack_order = dplyr::case_when(
      funding_source == "DWSRF Funds"                          ~ 1L,
      funding_source == "Other Funds Channeled Through SRFs"   ~ 2L
    )
  )

  if (nrow(bar_data_by_type) == 0) next 

# 2. Build a complete spine so every year × dodge_group × stack_order exists
spine_by_type <- tidyr::expand_grid(
  state_fiscal_year = unique(bar_data_by_type$state_fiscal_year),
  dodge_group       = 1:3,
  stack_order       = 1:2
)

bar_data_by_type <- spine_by_type |>
  dplyr::left_join(bar_data_by_type, by = c("state_fiscal_year", "dodge_group", "stack_order")) |>
  dplyr::mutate(funding_amount = tidyr::replace_na(funding_amount, 0))

# 3. Compute cumulative stack heights
bar_data_by_type <- bar_data_by_type |>
  dplyr::arrange(state_fiscal_year, dodge_group, stack_order) |>
  dplyr::group_by(state_fiscal_year, dodge_group) |>
  dplyr::mutate(
    ymax = cumsum(funding_amount),
    ymin = ymax - funding_amount
  ) |>
  dplyr::ungroup()

# 4. Compute xmin/xmax — anchor every year to the same integer regardless of missingness
bar_width <- 0.25
dodge_gap <- 0.05
group_gap <- 1.0

year_levels_by_type <- sort(unique(bar_data_by_type$state_fiscal_year))
n_groups    <- 3L

offsets <- seq(
  from       = -((n_groups - 1) / 2) * (bar_width + dodge_gap),
  by         = bar_width + dodge_gap,
  length.out = n_groups
)

bar_data_by_type <- bar_data_by_type |>
  dplyr::mutate(
    x_centre = match(state_fiscal_year, year_levels_by_type) * group_gap + offsets[dodge_group],
    xmin     = x_centre - bar_width / 2,
    xmax     = x_centre + bar_width / 2
  )

# 5. Drop zero-height rects (missing combos) so they don't affect the legend or tooltips
plot_data_by_type <- bar_data_by_type |>
  dplyr::filter(funding_amount > 0) 

# 6. Plot
  plot_tile_demand_by_type = paste0('Comparing Demand for Funds to Available Funds and Funds for projects on Fundable List: ', type)

  plot_by_type <- ggplot2::ggplot(plot_data_by_type) +
    ggplot2::geom_rect(
      ggplot2::aes(
        xmin = xmin, xmax = xmax,
        ymin = ymin, ymax = ymax,
        fill = bar,
        text = plt_str
      ),
      color = "black",
    ) +
    ggplot2::scale_x_continuous(
      breaks = seq_along(year_levels_by_type) * group_gap,
      labels = year_levels_by_type
    ) +
    ggplot2::scale_y_continuous(labels = label_currency()) +  
    ggplot2::scale_fill_manual(
      values = c(
        "Available Other Funds Channeled Through SRFs"        = "#3d7f1c",
        "Available DWSRF Funds"                               = "#6bc842",
        "On Fundable List Other Funds Channeled Through SRFs" = "#791a7b",
        "On Fundable List DWSRF Funds"                        = "#d927dcff",
        "Demand Other Funds Channeled Through SRFs"           = "#3861b2ff",
        "Demand DWSRF Funds"                                  = "#172f60"
      ),
      name = ""
    ) +
    ggplot2::labs(
      title = plot_tile_demand_by_type, 
      x = "State Fiscal Year", 
      y = "Funding Amount") +
    epic_chart_theme


plot_by_type_gp <- ggplotly(plot_by_type, tooltip = "text") |>
  plotly::layout(title = list(text = paste0(
    plot_tile_demand_by_type,
    '<br>',
    get_subtitle_str(bar_data_by_type$state_fiscal_year, state_name)
      )
    )
  )




if (save_plots) {
    run_save_plots(
      gg_plot   = plot_by_type,
      gp_object = plot_by_type_gp,
      name = paste0("test-",
        gsub("_", "-", janitor::make_clean_names(type)),
        "-ofsf-demand-yoy"
      )
    )
  }

}

```


## Funding amount DACs using DWSRF funds vs state funds -----

State-funds-dac-dwsrf-state-funds-yoy
Improve title - reference fundable list; DACs on the fundable list, remove the rest of the title
Incorporate funding available, stack srf and other state funds
Other states will have this granularity?
Note: MI aligns state funds to fundable lists
Incorporate a small system 
This might be more more consistent across states, since there is generally more availability of population data
Fix gray colors (and keep plot)
Future: dac + system size layers

```{r}

```


```{r}

```