---
title: "State Funding Pages"
author: "Phil Cork"
date: "2024-07-08"
---

This notebook is a draft of analysis for creating statistics and data viz for the SRF Funding Tracker's individual state funding pages (those which cover a single state fiscal year for a particular state.) It pulls data from both the project-level data and the pre-generated state-level agregate data. It uses Texas SFY 23 as an example.

As of December 2024, some of the tables and data visualizations are made manually from stats pulled from state IUPs, but this information will eventually be made into a tabular format by the Water Policy team and then should be imported from that forthcoming Google Sheet. At that time, much of this code which is uniquely specific to Texas SFY23 can and should be converted to an abstracted version that can be run for all states/fiscal years by simply changing the top-level variables.

# Setup

## Set Variables

```{r}
state_name <- "Texas"
sfy <- "2023"
```

## Imports and Styling

```{r}
options(scipen=999999999)

library(tidyverse)
library(janitor)
library(aws.s3)
library(scales)
library(ggpubr)

source("../../ds-resources/shared-functions.R")
run_code_from_file("../../local/aws.txt")

```

```{r}
# fonts: 
sysfonts::font_add_google("Lato")
showtext::showtext_auto()

# theme: 
# legend position is right, text size is at least 10, and text is Lato
epic_chart_theme <- theme_minimal() + 
  theme(legend.position = "right", 
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), 
        axis.text.x = element_text(margin = margin(t = 10, r = 0, 
                                                   b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, 
                                                    b = 0, l = 0)), 
        axis.text.y = element_text(margin = margin(t = 0, r = 10, 
                                                   b = 0, l = 0)), 
        axis.title.y = element_text(margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0))) 
```

## Import Projects & State Data

```{r}
# import project-level data
all_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

state_data <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-state-stats.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)
```

# Create Plots from Manual Data Entry

## Grants Allocated

```{r}
grants_allocated <- data.frame(fed_cap_grants = c("FFY22 Base DWSRF", "FFY22 BIL General Supplemental",
                                                  "FFY22 BIL Lead", "FFY23 BIL Lead",
                                                  "FFY22 BIL Emerging Contaminants"),
                               amount = c(54911000, 140993000, 222155000, 146246000, 59202000))

```

```{r}
ggplot(grants_allocated, aes(x=amount, y=reorder(fed_cap_grants, amount))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Federal Cap Grants", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_x_continuous(labels=label_dollar(scale=.000001, suffix="M")) + 
  epic_chart_theme
```

## Principal Forgiveness

```{r}
principal_forgiveness <- data.frame(fed_cap_grants = c("FFY22 Base DWSRF, DACs", "FFY22 Base DWSRF, Any", "FFY22 BIL General Supplemental",
                                                       "FFY22 BIl Lead", "FFY23 BIL Lead", "FFY22 BIL Emerging Contaminants"),
                               amount = c(8913430, 9000000, 69086570, 108855950, 71660540, 57910960))

```

```{r}
ggplot(principal_forgiveness, aes(x=amount, y=reorder(fed_cap_grants, amount))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Principal Forgiveness", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_x_continuous(labels=label_dollar(scale=.000001, suffix="M")) + 
  epic_chart_theme
```

## Comparing Funds in Process Stages

```{r}
available_award_comp <- data.frame(category = c("Available", "Submitted", "Expecting Funding"),
                                   amount = c(342000000, 2500000000, 324000000)) %>%
  mutate(category = factor(category, levels=c("Submitted", "Available", "Expecting Funding")))

ggplot(available_award_comp, aes(x=category, y=amount)) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Comparing Available Funds to Submitted Needs and Expecting Funding", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_y_continuous(labels=label_dollar(scale=.000000001, suffix="B")) + 
  epic_chart_theme
```

## Comparing Project Costs & Available Funds

```{r}

lead_costs <- data.frame(category=c("Available Funds", "Remaining Submitted Lead Project Costs"),
                         amount=c(353975300, (4102947493.48-353975300)),
                         amount_str=c("$353M", "$3.7B"),
                         amount_pct=c("9%", "91%"))
#NOTE: Have to plot the bigger number first for it to appear intuitively
lead_costs$category <- factor(lead_costs$category, levels=c("Remaining Submitted Lead Project Costs", "Available Funds"))

plot_labs <- paste0(lead_costs$category, " (", lead_costs$amount_pct, ", ", lead_costs$amount_str, ")")
p <- ggdonutchart(lead_costs,
             "amount",
             fill="category",
             label=plot_labs,
             color="white",
             palette=c("lightgrey", "#4ea324"))
p + theme(legend.position="none",
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), ) + 
  labs(title="Comparing Lead Replacement Costs to Available Funds")

```

```{r}
ec_costs <- data.frame(category=c("Available Funds", "Remaining Emerging Contaminants Costs"),
                         amount=c(57910960, (175000000-57910960)),
                         amount_str=c("$57M", "$175M"),
                         amount_pct=c("33%", "67%"))
#NOTE: Have to plot the bigger number first for it to appear intuitively
ec_costs$category <- factor(ec_costs$category, levels=c("Remaining Emerging Contaminants Costs", "Available Funds"))

plot_labs <- paste0(ec_costs$category, " (", ec_costs$amount_pct, ", ", ec_costs$amount_str, ")")
p <- ggdonutchart(ec_costs,
             "amount",
             fill="category",
             label=plot_labs,
             color="white",
             palette=c("lightgrey", "#4ea324"))
p + theme(legend.position="none",
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), ) + 
  labs(title="Comparing Emerging Contaminants Costs to Available Funds")
```

# State-Level Data

```{r}

if (state_data$principal_forgiveness_status != "Missing") {
  
  pf_projects <- all_projects %>%
    filter(principal_forgiveness != "0" & principal_forgiveness != "No Information")
  
}

```

How many state-defined DACs are not expecting to receive funding? Out of how many?

```{r}
state_data$disadvantaged_not_expecting_funding_projects[1]
state_data$disadvantaged_projects[1]
```

```{r}
dacs_ef <- data.frame(category=c("DACs Expecting Funding", "DACs Not Expecting Funding"),
                         amount=c(27, 97),
                         amount_str=c("27", "97"),
                         amount_pct=c("22%", "78%"))
#NOTE: Have to plot the bigger number first for it to appear intuitively
dacs_ef$category <- factor(dacs_ef$category, levels=c("DACs Not Expecting Funding", "DACs Expecting Funding"))

plot_labs <- paste0(dacs_ef$category, " (", dacs_ef$amount_pct, ", ", dacs_ef$amount_str, ")")
p <- ggdonutchart(dacs_ef,
             "amount",
             fill="category",
             label=plot_labs,
             color="white",
             palette=c("lightgrey", "#4ea324"))
p + theme(legend.position="none",
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), ) + 
  labs(title="How many DACs are expected to receive funding?")
```


```{r}
# Add plot of expecting_funding Y/N with DAC Y/N
```

X percent of SFY23 funding cycle applicants qualify as DACs, and X percent of projects designated for DWSRF anticipated awards qualify as DACs"

```{r}
state_data$disadvantaged_percent[1]
state_data$disadvantaged_expecting_funding_percent_ef
```

```{r}
big_number_donut_plot(
  state_data$disadvantaged_expecting_funding_percent_ef/100,
  "Lato", "#4ea324",
  "Percent of Projects Expecting Funding that are DACs"
)

#38 projects EF, 27 are DAC

ef_dacs <- data.frame(category=c("Non-DACs Expecting Funding", "DACs Expecting Funding"),
                         amount=c(11, 27),
                         amount_str=c("11", "27"),
                         amount_pct=c("29%", "71%"))
#NOTE: Have to plot the bigger number first for it to appear intuitively
ef_dacs$category <- factor(ef_dacs$category, levels=c("DACs Expecting Funding", "Non-DACs Expecting Funding"))

plot_labs <- paste0(ef_dacs$category, " (", ef_dacs$amount_pct, ", ", ef_dacs$amount_str, ")")
p <- ggdonutchart(ef_dacs,
             "amount",
             fill="category",
             label=plot_labs,
             color="white",
             palette=c("#4ea324", "lightgrey"))
p + theme(legend.position="none",
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), ) + 
  labs(title="How many projects expected to receive funding are going to DACs?")
```


X percent of SFY23 funding cycle applicants qualify as DACs, and X percent of projects designated for DWSRF anticipated awards qualify as DACs -- same as above, but for Lead specifically
```{r}

if (state_data$disadvantaged_status != "Complete" | state_data$disadvantaged_projects == 0) {
  state_data$disadvantaged_status[1]
  print("Check data dictionary for more context to see how this feature should be discussed, because a 0 in disadvantaged_projects might indiciate that the defintion is missing or portially incomplete. Example: In Texas SFY23, 'Disadvantaged' is present for General projects, but not Lead projects, and EC projects are entirely missing. Thus, there is no problem reporting about DACs in general, but when it comes to Lead-specific questions, we are unable to address those questions due to insufficient data.")
} else {
  # percent of projects that are dacs and lead out of all projects
  state_data$dac_lead_projects / state_data$projects *100
  # percent of expecting_funding projects that are dac and lead
  state_data$dac_lead_ef_projects / state_data$expecting_funding_projects
  
}


```

# Project-Level Data

```{r}
# range of project scores for projects expecting funding
tibble(
all_projects %>%
  filter(expecting_funding == "Yes") %>%
  mutate(score_numeric = as.numeric(project_score)) %>%
  group_by(state, state_fiscal_year) %>%
  summarize(min_score = min(score_numeric),
            max_score = max(score_numeric))
)
```
