---
title: "State Funding Pages"
author: "Phil Cork"
date: "2024-07-08"
---

This notebook is a draft of analysis for creating statistics and data viz for the SRF Funding Tracker's individual state funding pages (those which cover a single state fiscal year for a particular state.) It pulls data from both the project-level data and the pre-generated state-level agregate data. It uses Texas SFY 23 as an example.

As of December 2024, some of the tables and data visualizations are made manually from stats pulled from state IUPs, but this information will eventually be made into a tabular format by the Water Policy team and then should be imported from that forthcoming Google Sheet. At that time, much of this code which is uniquely specific to Texas SFY23 can and should be converted to an abstracted version that can be run for all states/fiscal years by simply changing the top-level variabes.

# Setup

## Set Variables

```{r}
state_name <- "Texas"
sfy <- "2023"
```

## Imports and Styling

```{r}
options(scipen=999999999)

library(tidyverse)
library(janitor)
library(aws.s3)
library(scales)

source("../../ds-resources/shared-functions.R")
run_code_from_file("../../local/aws.txt")

```

```{r}
# fonts: 
sysfonts::font_add_google("Lato")
showtext::showtext_auto()

# theme: 
# legend position is right, text size is at least 10, and text is Lato
epic_chart_theme <- theme_minimal() + 
  theme(legend.position = "right", 
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), 
        axis.text.x = element_text(margin = margin(t = 10, r = 0, 
                                                   b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, 
                                                    b = 0, l = 0)), 
        axis.text.y = element_text(margin = margin(t = 0, r = 10, 
                                                   b = 0, l = 0)), 
        axis.title.y = element_text(margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0))) 
```

## Import Projects & State Data

```{r}
# import project-level data
all_projects <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-all-projects.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)

state_data <- s3read_using(read.csv, object="s3://water-team-data/clean_data/srf_project_priority_lists/dwsrf-funding-tracker-state-stats.csv") %>%
  clean_names() %>%
  filter(state == state_name & state_fiscal_year == sfy)
```

# Create Plots from Manual Data Entry

## Grants Allocated

```{r}
grants_allocated <- data.frame(fed_cap_grants = c("FFY22 Base DWSRF", "FFY22 BIL General Supplemental",
                                                  "FFY22 BIL Lead", "FFY23 BIL Lead",
                                                  "FFY22 BIL Emerging Contaminants"),
                               amount = c(54911000, 140993000, 222155000, 146246000, 59202000))

```

```{r}
ggplot(grants_allocated, aes(x=amount, y=reorder(fed_cap_grants, amount))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Federal Cap Grants", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_x_continuous(labels=label_dollar(scale=.000001, suffix="M")) + 
  epic_chart_theme
```

## Principal Forgiveness

```{r}
principal_forgiveness <- data.frame(fed_cap_grants = c("FFY22 Base DWSRF, DACs", "FFY22 Base DWSRF, Any", "FFY22 BIL General Supplemental",
                                                       "FFY22 BIl Lead", "FFY23 BIL Lead", "FFY22 BIL Emerging Contaminants"),
                               amount = c(8913430, 9000000, 69086570, 108855950, 71660540, 57910960))

```

```{r}
ggplot(principal_forgiveness, aes(x=amount, y=reorder(fed_cap_grants, amount))) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Principal Forgiveness", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_x_continuous(labels=label_dollar(scale=.000001, suffix="M")) + 
  epic_chart_theme
```

## Comparing Funds in Process Stages

```{r}
available_award_comp <- data.frame(category = c("Available", "Submitted", "Invited", "Awarded"),
                                   amount = c(342000000, 2500000000, 324000000, 88884918)) %>%
  mutate(category = factor(category, levels=c("Available", "Submitted", "Invited", "Awarded")))

ggplot(available_award_comp, aes(x=category, y=amount)) + 
  geom_bar(stat="identity", fill="#4ea324") +
  labs(x="", y="", title="Funds From Availability to Awarded", subtitle=paste0(state_name, ", ", sfy)) + 
  scale_y_continuous(labels=label_dollar(scale=.000000001, suffix="B")) + 
  epic_chart_theme
```

## Comparing Project Costs & Available Funds

```{r}
#TODO: Push this function into the EPIC plotting theme or function folders when tested
# from https://rfortherestofus.com/2022/09/how-to-make-a-donut-chart-in-ggplot
big_number_donut_plot <- function(value, font_family, highlight_color, title_str) {
  # Wrangle data to get a data frame in the format we need it in to make our donut chart
  df <- tibble(x = 1, y = value) %>% 
    mutate(y_negative = 1 - y) %>% 
    pivot_longer(cols = -x) 
  # Create a nicely formatted big number to go in the donut hole
  big_number_text_label <- percent(value, accuracy = 1)
  
  # Create our plot
  ggplot(df,
         aes(x = x,
             y = value,
             fill = name)) +
    # Add a bar, but don't add the legend
    geom_col(show.legend = FALSE) +
    # A pie/donut chart is a bar chart with polar coordinates
    # Add polar coordinates and set the direction to -1 
    # so the filled in part starts at the top and goes clockwise
    coord_polar(theta = "y",
                direction = -1) +
    # Set the limits, which is important for adding the hole
    xlim(c(-2, 2)) +
    # Set a color scale with the highlighted section in whatever color
    # is chosen with the highlight_color argument and the rest in a light gray
    scale_fill_manual(values = c(highlight_color, "grey90")) +
    # Set theme_void() to remove grid lines and everything else from the plot
    theme_void() +
    labs(title=title_str) + 
    # Add the big number in the center of the hole
    annotate("text",
             label = big_number_text_label,
             family = font_family,
             fontface = "bold",
             color = highlight_color,
             size = 12,
             x = -2,
             y = 0)
}
```

```{r}

big_number_donut_plot(353975300/4000000000, "Lato", "#4ea324", "Comparing Lead Replacement Project Costs to Available Funds")

```

```{r}
big_number_donut_plot(57910960/175000000, "Lato", "#4ea324", "Comparing Emerging Contaminant Project Costs to Available Funds")
```

# State-Level Data

```{r}

if (state_data$principal_forgiveness_status != "Missing") {
  
  pf_projects <- all_projects %>%
    filter(principal_forgiveness != "0" & principal_forgiveness != "No Information")
  
}

```

How many state-defined DACs are not expecting to receive funding? Out of how many?

```{r}
state_data$disadvantaged_not_expecting_funding_projects[1]
state_data$disadvantaged_projects[1]
```

```{r}
big_number_donut_plot(
  state_data$disadvantaged_not_expecting_funding_projects[1] / state_data$disadvantaged_projects[1],
  "Lato", "#4ea324",
  "Percent of DACs not expected to receive funding"
)
```


```{r}
# Add plot of expecting_funding Y/N with DAC Y/N
```

X percent of SFY23 funding cycle applicants qualify as DACs, and X percent of projects designated for DWSRF anticipated awards qualify as DACs"

```{r}
state_data$disadvantaged_percent[1]
state_data$disadvantaged_expecting_funding_percent_ef
```

```{r}
big_number_donut_plot(
  state_data$disadvantaged_expecting_funding_percent_ef/100,
  "Lato", "#4ea324",
  "Percent of Projects Expecting Funding that are DACs"
)
```


X percent of SFY23 funding cycle applicants qualify as DACs, and X percent of projects designated for DWSRF anticipated awards qualify as DACs -- same as above, but for Lead specifically
```{r}

if (state_data$disadvantaged_status != "Complete" | state_data$disadvantaged_projects == 0) {
  state_data$disadvantaged_status[1]
  print("Check data dictionary for more context to see how this feature should be discussed, because a 0 in disadvantaged_projects might indiciate that the defintion is missing or portially incomplete. Example: In Texas SFY23, 'Disadvantaged' is present for General projects, but not Lead projects, and EC projects are entirely missing. Thus, there is no problem reporting about DACs in general, but when it comes to Lead-specific questions, we are unable to address those questions due to insufficient data.")
} else {
  # percent of projects that are dacs and lead out of all projects
  state_data$dac_lead_projects / state_data$projects *100
  # percent of expecting_funding projects that are dac and lead
  state_data$dac_lead_ef_projects / state_data$expecting_funding_projects
  
}


```

# Project-Level Data

```{r}
# range of project scores for projects expecting funding
tibble(
all_projects %>%
  filter(expecting_funding == "Yes") %>%
  mutate(score_numeric = as.numeric(project_score)) %>%
  group_by(state, state_fiscal_year) %>%
  summarize(min_score = min(score_numeric),
            max_score = max(score_numeric))
)
```
