

This notebook generates plots from manually-gathered statistics found on the Texas State Overview page template.
In the future, these numbers will be entered into a database and this code will need to be updated to pull it in and recreate the plots in a more reproducible fashion.

# Imports & Settings 

```{r}
state_name <- "Texas"
```


```{r}
library(tidyverse)
library(janitor)
library(aws.s3)
library(scales)
library(lubridate)
library(sysfonts)
library(showtext)

source("../../ds-resources/shared-functions.R")
run_code_from_file("../../local/aws.txt")

options(scipen=999)

# fonts: 
sysfonts::font_add_google("Lato")
showtext::showtext_auto()

# theme: 
# legend position is right, text size is at least 10, and text is Lato
epic_chart_theme <- theme_minimal() + 
  theme(legend.position = "right", 
        text = element_text(size = 11, family = "Lato"), 
        legend.text = element_text(size = 10, family = "Lato"), 
        legend.title = element_text(size = 11, family = "Lato"), 
        axis.text.x = element_text(margin = margin(t = 10, r = 0, 
                                                   b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, 
                                                    b = 0, l = 0)), 
        axis.text.y = element_text(margin = margin(t = 0, r = 10, 
                                                   b = 0, l = 0)), 
        axis.title.y = element_text(margin = margin(t = 0, r = 10, 
                                                    b = 0, l = 0))) 

cont_palette <- colorRampPalette(c("#172f60","#4ea324"))

cat_palette <- colorRampPalette(c("#172f60","#1054a8",
                                  "#791a7b","#de9d29", 
                                  "#b15712","#4ea324"))

cat_palette_pastel <- colorRampPalette(c("#526489","#527CAF",
                                         "#B077B2","#E4BE7C",
                                         "#b15712","#82AB6E"))

```


# Base Federal Capitalization Grants
$54,911,000
$40,181,000
$37,157,000

```{r}
fed_cap_grants <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025"),
                             amount = c(54911000, 40181000, 37157000))

ggplot(fed_cap_grants, aes(x=state_fiscal_year, y=amount)) + 
  geom_bar(stat="identity", fill="#172f60") + 
  scale_y_continuous(labels=label_currency(scale=.000001, suffix="M")) + 
  labs(x="", y="", title="Base DWSRF Federal Capitlization Grants", subtitle=paste0(state_name, ", SFY23-25")) + 
  epic_chart_theme

```


# Base Principal Forgiveness for DACs
$8,913,430 (16%)
$7,712,000 (19.1%)
$8.500.000 (22.7%)

```{r}
pf_for_dacs <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025"),
                          amount = c(8913430, 7712000, 8500000),
                          # percents are 16, 19.1, 22.7, need to scale from left-side going 0-10m to right-side being 0-100
                          pct = c(1600000, 1910000, 2270000),
                          grouping = "TX", "TX", "TX")

ggplot(pf_for_dacs, aes(x=state_fiscal_year)) + 
  geom_bar(aes(y=amount), stat="identity", fill="#172f60") +
  
    # Add secondary y-axis and manually set the range of the right hand
  scale_y_continuous(name = "PF for DACs", labels=label_dollar()) +
  labs(x="", y="", title="Principal Forgiveness for DACs, Base DWSRF Program", subtitle=paste0(state_name, ", SFY23-25")) + 
  epic_chart_theme
```

```{r}
pf_for_dacs_pct <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025"),
                             amount = c(.16, .191, .227),
                             grouping = c("TX", "TX", "TX"))

ggplot(pf_for_dacs_pct, aes(x=state_fiscal_year, y=amount)) + 
  geom_line(aes(group=grouping), color="#172f60", size=1.5) + 
  geom_point(size=4) + 
  scale_y_continuous(labels=label_percent(), limits=c(0,1)) + 
  labs(x="", y="", title="Percent Principal Forgiveness for DACs, Base DWSRF Program", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme
```


# Total Funding Requested under Base & BIL General Supplemental Program PPL
$2492000652
$3255535627
$4643722053.80
```{r}

funding_requested <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025"),
                             amount = c(2492000652, 3255535627, 4643722053.80))

ggplot(funding_requested, aes(x=state_fiscal_year, y=amount)) + 
  geom_bar(stat="identity", fill="#172f60") + 
  scale_y_continuous(labels=label_currency(scale=.000000001, suffix="B")) + 
  labs(x="", y="", title="Total Funding Requested under Base & BIL General Supplemental Program PPL", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme


```

# Total Funding Available for Projects under Base & BIL General Supplemental Program
342000000
435066830
444395440

```{r}
funding_requested_available <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025",
                                                      "SFY2023", "SFY2024", "SFY2025"),
                             amount = c(2492000652-342000000, 3255535627-435066830, 4643722053.80-444395440,
                                        342000000, 435066830, 444395440),
                             category = c("Requested", "Requested", "Requested",
                                          "Total Available", "Total Available", "Total Available")
                             )

ggplot(funding_requested_available, aes(x=state_fiscal_year, y=amount, fill=category)) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_continuous(labels=label_currency(scale=.000000001, suffix="B")) + 
  scale_fill_manual(values = cat_palette(2), 
                      name = "") + 
  labs(x="", y="", title="Comparing Requested & Available Funds under Base & Bil Gen. Supp", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme

```

# Total Funding Available for Lead Projects
213455950
140519350
27528500


```{r}

lead_available <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025"),
                             amount = c(213455950+140519350, 0, 27528500))

ggplot(lead_available, aes(x=state_fiscal_year, y=amount)) + 
  geom_bar(stat="identity", fill="#172f60") + 
  scale_y_continuous(labels=label_currency(scale=.000001, suffix="M")) + 
  labs(x="", y="", title="Total Funding Available, BIL LSLR", subtitle=paste0(state_name, ", SFY23-25")) + 
  epic_chart_theme



```


EC Program

# Total Funding Requested, EC
175620496
552609451
# Total Funding Available for Projects
57910960
59372280

```{r}
funding_requested_available_ec <- data.frame(state_fiscal_year = c("SFY2023", "SFY2024", "SFY2025",
                                                      "SFY2023", "SFY2024", "SFY2025"),
                             amount = c(175620496-57910960, 552609451-59372280, 0,
                                        57910960, 59372280, 0),
                             category = c("Requested", "Requested", "Requested",
                                          "Total Available", "Total Available", "Total Available")
                             )

ggplot(funding_requested_available_ec, aes(x=state_fiscal_year, y=amount, fill=category)) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_continuous(labels=label_currency(scale=.000001, suffix="M")) + 
  scale_fill_manual(values = cat_palette(2), 
                      name = "") + 
  labs(x="", y="", title="Comparing Requested & Available Funds, BIL EC", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme
```

# Set Asides

State	Program	State Fiscal Year	Set Asides $	Set Asides %
Texas	Base	2023	12885760	23%
Texas	Base	2024	10528960	26%
Texas	Base	2025	10045120	27%
Texas	BIL General Supplemental	2023	5639720	4%
Texas	BIL General Supplemental	2024	18714680	11%
Texas	BIL General Supplemental	2025	19330240	11%
Texas	Base & BIL General Supplemental	2023	18525480	9%
Texas	Base & BIL General Supplemental	2024	29243640	14%
Texas	Base & BIL General Supplemental	2025	29375360	13%
Texas	Lead Service Line Replacement	2023	14425700	4%
Texas	Lead Service Line Replacement	2024	NA	NA	NA
Texas	Lead Service Line Replacement	2025	1121500	4%
Texas	Emerging Contaminants	2023	2368080	4%
Texas	Emerging Contaminants	2024	2436560	4%
Texas	Emerging Contaminants	2025

```{r}
set_asides <- data.frame(state_fiscal_year=c("SFY2023","SFY2024", "SFY2025",
                                             "SFY2023","SFY2024", "SFY2025",
                                             "SFY2023","SFY2024", "SFY2025",
                                             "SFY2023","SFY2024", "SFY2025",
                                             "SFY2023","SFY2024", "SFY2025"),
                         program=c("Base", "Base", "Base",
                                   "BIL General Supplemental", "BIL General Supplemental", "BIL General Supplemental",
                                   "Base & BIL General Supplemental", "Base & BIL General Supplemental", "Base & BIL General Supplemental",
                                   "Lead Service Line Replacement", "Lead Service Line Replacement", "Lead Service Line Replacement",
                                   "Emerging Contaminants", "Emerging Contaminants", "Emerging Contaminants"),
                         set_aside_dollar=c(12885760, 10528960, 10045120,
                                            5639720, 18714680, 19330240,
                                            18525480, 29243640, 29375360,
                                            14425700, NA, 1121500,
                                            2368080, 2436560, NA),
                         set_aside_pct=c(.23, .26, .27,
                                         .04, .11, .11,
                                         .09, .14, .13,
                                         .04, 0, .04,
                                         .04, .04, 0)
                         )
```

```{r}
ggplot(set_asides, aes(x=state_fiscal_year, y=set_aside_dollar, fill=program)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_y_continuous(labels=label_currency()) + 
  scale_fill_manual(values = cat_palette(5), 
                      name = "") + 
  labs(x="", y="", title="Set Asides", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme
  
```

```{r}
ggplot(set_asides, aes(x=state_fiscal_year, y=set_aside_dollar, fill=program)) + 
  geom_bar(stat="identity", position="stack") + 
  scale_y_continuous(labels=label_currency()) + 
  scale_fill_manual(values = cat_palette(5), 
                      name = "") + 
  labs(x="", y="", title="Set Asides", subtitle=paste0(state_name, " SFY23-25")) + 
  epic_chart_theme
```

```{r}
ggplot(set_asides, aes(x=state_fiscal_year, y=set_aside_pct, color=program)) + 
  geom_line(aes(group=program), size=1.5) + 
  geom_point(size=4) + 
  scale_y_continuous(labels=label_percent(), limits=c(0,.31)) + 
  labs(x="", y="", title="Set Asides, Percent Utilized", subtitle=paste0(state_name, " SFY23-25")) + 
  scale_color_manual(values = cat_palette(5), 
                      name = "") + 
  epic_chart_theme
```

